### 2017-03-09 paper.tex
First very rough pass, laying out basic ideas in mostly coherent sentences.

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\title{Efficient Reversible Arithmetic Circuits using a Dirty Bit}
\author{et al}

\begin{document}
\maketitle

\begin{abstract}
We describe classical reversible constructions, that require at most 1 ancilla in an unspecified state, for incrementing a register in $O(N)$ depth and size, for modular addition of one register into another in $O(N)$ depth and size, and for modular addition of a compile-time constant into a register in $O(N)$ depth and $O(N \lg N)$ size.
We note some applications of the constructions, such as performing Shor's algorithm on an $n$-bit product with only $2n+1$ qubits.
\end{abstract}

\section{Introduction}

When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
Traditionally [[[where? who?]]], an ancilla is a bit initialized into a {\em known} state, available for use by the circuit.
But "dirty" ancilla, that are in an unknown state that must be preserved, are also a useful resource [[[where? who? haner2016]]].
Although clean ancilla are better than dirty ancilla for making compact simple circuits [[[example?]]], dirty ancilla are more readily available  because any unused bit in a circuit can be temporarily borrowed as a dirty ancilla.

The ability to borrow dirty ancilla from other parts of the circuit makes circuit constructions that require only dirty ancilla flexible, especially when space-constrained or when proper ancilla may be far away [[[(e.g. surface code)]]].
For example, [haner2016] used dirty ancilla to cut the cost of performing Shor's algorithm on an $n$-bit number, using only $2n+2$ qubits, from ???? to $O(n^3)$ depth and $O(n^3 \lg n)$ size.
In this paper we improve on [haner2016]'s result by cutting the number of required qubits from $2n+2$ to $2n+1$.

The paper is structured as follows.
Sections ??? through ??? describe various circuit constructions.
Section ??? applies those circuit constructions to exist problems.
... That's it.

[[[What about underlying details, like reducing many-controlled many-nots? Nearby details like bootstrapping an ancilla with quantum operations? Parity issues that prevent some operations from being performed with no ancilla?]]]

All constructions assume a 2s-complement representation of integers.
All circuit diagrams with multi-wire integer operations put the least significant bit towards the top.

\section{$O(N)$ inline adder}

[van2004] describes a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla.
The ancilla must be off, or else the result is incremented into returning $a+b+1$ instead of $a+b$.

The need for the ancilla to be clean can be fixed by conditionally NOT-ing the source and target registers before and after the addition when the carry is on.
Because $\lnot x = -x-1$, temporarily inverting the target turns the addition into a subtraction.
Temporarily inverting the source turns the subtraction back into an addition, but also introduces an extra shift by -1.
The -1 cancels the carry's +1 effect.

On top of this improvement, the carry role can be merged into the most-significant-bit of the source register.
Although the MSB can't invert itself, the circuit still works because we don't need an overflow signal that distinguishes between the MSB toggling from on to off instead of from off to on.

The improved reversible adder is shown in figure \ref{fig:inlineadder}.
Note that, because the source register is used as workspace, it must be at least as large as the target register.

\begin{figure}
  \centering
  \includegraphics[totalheight=3cm]{inline-adder.png}
  \caption{\em Adder using no ancilla and $\Theta(N)$ CNOTs and CSWAPs. Based on [van2004].}
  \label{fig:inlineadder}
\end{figure}

To perform subtraction instead of addition: run the circuit in reverse, temporarily invert the target register by bordering the circuit with NOT gates on the target, or invert the CSWAP controls.

\section{$O(N)$ incrementer}

Note that a register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it for any $x$.
When $N$ dirty ancilla are available, this can be done directly as shown in figure \ref{fig:double-sub-increment}.

\begin{figure}
  \centering
  \includegraphics[totalheight=2cm]{double-sub-increment.png}
  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(N)$ depth, size, and $N$ dirty ancilla.}
  \label{fig:double-sub-increment}
\end{figure}

An increment gate can be decomposed into halves where the top half is incremented only if all of the bottom bits are on, and then the bottom half is incremented.
By borrowing the top half, the bottom half can be incremented with the double-subtraction trick.
But the trick doesn't work on the top half, because the top half can't be borrowed when being controlled by it.

A workaround for this is to change the second subtraction to an addition, and to conditionally toggle the target bits instead of the input bits.
When the condition isn't satisfied, the addition and subtraction cancel each other.
When the condition is satisfied, the addition inverts into a second subtraction and the input is subtracted from the target twice.
However, because we are conditioning on the input bits all being on, the input must be -1 and the target must have been incremented by 2.
To halve the +2 into a +1, left-shift the target and use a dirty bit as its LSB.

\begin{figure}
  \centering
  \includegraphics[totalheight=1.5cm]{compact-increment.png}
  \caption{\em Odd-sized $\Theta(N)$ increment with 1 dirty ancilla.}
  \label{fig:compact-increment}
\end{figure}

The described incrementer construction is shown in figure \ref{fig:compact-increment}.
However, because the adder/subtracter construction requires the target and source registers to be the same size, this incrementer construction only works on odd-sized registers.
For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
The construction is shown in \ref{fig:compact-increment-even}.
Note that it is easily modified to perform efficient increments with arbitrarily many controls.

\begin{figure}
  \centering
  \includegraphics[totalheight=2cm]{compact-increment-even.png}
  \caption{\em Even-sized $\Theta(N)$ increment with 1 dirty ancilla.}
  \label{fig:compact-increment-even}
\end{figure}

As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting all of the controls.

Our incrementer is classically optimal in asymptotic size and number of ancilla, but not depth.
The size is asymptotically optimal because any incrementer must touch all $n$ bits.
The number of ancilla is classically optimal, assuming we want to reduce to a fixed-size gate set, because the parity of the permutation performed by an increment operation is odd but the parity of the permutation performed by a classical gate that doesn't cover the entire circuit is even.
Therefore, without a dirty bit to use as workspace, no classical circuit can perform an increment gate.
Quantumly, the parity barrier can be bypassed by using partial rotations (see figure \ref{fig:bootstrap-ancilla}).

\begin{figure}
  \centering
  \includegraphics[totalheight=1.3cm]{ancilla-bootstrap.png}
  \caption{\em Bootstrapping dirty ancilla out of an increment gate using quantum operations.}
  \label{fig:bootstrap-ancilla}
\end{figure}

\section{$O(N \lg N)$ constant $\text{PivotFlip}$}

A pivot-flip operation reverses the order of states less than or equal to a given pivot value.
For example, a pivot-flip with the pivot equal to 3 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
We will use pivot-flips later, when implementing modular addition.

For concreteness, the exact permutation of a pivot-flip is:

$$\text{PivotFlip}_k = \sum_{i=0}^k |k-i\rangle \langle i| + \sum_{i=k+1}^{N-1} |i\rangle \langle i|$$

To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
However, instead of doing nothing to the elements past the pivot, it also flips those elements as shown in figure \ref{fig:double-flip}.

\begin{figure}
  \centering
  \includegraphics[totalheight=7cm]{double-flip.png}
  \caption{\em Subtraction followed by logical negation performs a flip both above and below the pivot.}
  \label{fig:double-flip}
\end{figure}

Note that $x \rightarrow \lnot(x - a)$ is its own inverse and maintains the invariant that a given value will cause an underflow when subtracting $a$.
So, by applying the operation twice, controlled by a dirty bit that is toggled by the underflow signal, the desired effect is achieved.
The part of the range below the pivot does trigger the underflow, so only one of the controlled flips happens.
The part of the range above the pivot does not cause an underflow, and so flips no times or two times, undoing itself.

The resulting pivot-flip circuit is shown in figure \ref{fig:const-pivot-flip}.

\begin{figure}
  \centering
  \includegraphics[totalheight=2cm]{pivot-flip.png}
  \caption{\em Pivot flip circuit.
  Note that the top wire bundle is a compile-time constant.
  The underflow toggle is done with an $n+1$-bit subtraction then an $n$-bit addition.
  [[[Requires an extra dirty bit for the n+1-bit subtraction?]]]
  The controlled subtractions are done by controlling every generated operation.}
  \label{fig:const-pivot-flip}
\end{figure}

Because the pivot is a compile-time constant, and so there is no source register workspace, the inline adder construction can't be used.
Instead, we use the constant-adder from [haner2016].
That adder requires $O(N)$ depth and $O(N \lg N)$ size, and therefore so does the constant-pivot-flip circuit.

\section{$O(N \lg N)$ Modular Constant-Adder}

A modular addition is three pivot flips.
To add $a$ modulo $n$, perform a pivot-flip at $n-a$, a pivot-flip at $n$, and then a pivot-flip at $a$.

[[[diagram]]]

Because the modular addition does a constant number of pivot-flips, its resource usage is identical: one [[[two?]]] dirty bits, $O(N)$ depth, and $O(N \lg N)$ size.

\section{$O(N)$ $\text{PivotFlip}$}

For completeness, we also discuss how to do a pivot-flip when the pivot is stored in a register instead of a compile-time constant.
The original construction doesn't work in this case because it uses an extra bit in the source register.

To work around the problem we propagate the increment into the operations on the target.
[[[But the addition sizes don't match up?]]]

[[[Maybe just ignore the increment size problem, then fixup with a conditioned all-flip?]]]

Note that, when storing in a register, we offset the pivot $k$ by one.
This is to avoid the redundancy $\text{PivotFlip}_0 = \text{PivotFlip}_1 = I$ and to move the flip-all operation $\text{PivotFlip}_{N}$ into range.
Concretely, we want to perform this permutation:

$$\text{RegisterPivotFlip} = \sum_{k=0}^{N-1} |k\rangle \langle k| \otimes \text{PivotFlip}_{k+1}$$

Basically we do this by just inlining the increments and decrements into the addition operations or by using that other thing I tried.

\section{$\Theta(N)$ Modular Addition}

Same high-level construction as the constant modular adder.
Do three pivot flips, but register-based instead of constant-based.
This reduces the size cost from $O(N \lg N)$ to $O(N)$.

\section{Applications}

Shor's algorithm in $2n+1$ qubits.

References:

haner2016 https://arxiv.org/abs/1611.07995
Factoring using 2n+2 qubits with Toffoli based modular multiplication
Thomas Häner, Martin Roetteler, Krysta M. Svore

van2004
Optimal Design of A Reversible Full Adder
YVAN VAN RENTERGEM AND ALEXIS DE VOS
2004

\end{document}

### 2017-03-10 paper.tex
Working out the rest of the constructions in detail

@@
 \begin{abstract}
-We describe classical reversible constructions, that require at most 1 ancilla in an unspecified state, for incrementing a register in $O(N)$ depth and size, for modular addition of one register into another in $O(N)$ depth and size, and for modular addition of a compile-time constant into a register in $O(N)$ depth and $O(N \lg N)$ size.
-We note some applications of the constructions, such as performing Shor's algorithm on an $n$-bit product with only $2n+1$ qubits.
+We describe classical and quantum circuits for performing reversible arithmetic using at most 1 ancilla in an unknown state.
+Our constructions can perform incrementation, comparisons, addition, and modular addition with classical circuits of linear size and depth.
+We also show how to perform modular offset, improving the best known space requirements for Shor's algorithm from $2n+2$ qubits to $2n+1$ qubits, in linear depth and $O(n \lg n)$ size.
 \end{abstract}
@@
 When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
-Traditionally [[[where? who?]]], an ancilla is a bit initialized into a {\em known} state, available for use by the circuit.
-But "dirty" ancilla, that are in an unknown state that must be preserved, are also a useful resource [[[where? who? haner2016]]].
-Although clean ancilla are better than dirty ancilla for making compact simple circuits [[[example?]]], dirty ancilla are more readily available  because any unused bit in a circuit can be temporarily borrowed as a dirty ancilla.
+An ancilla qubit is an extra qubit, on top of the qubits used for input and output, available for use as temporary workspace.
+Ancilla may be initialized to a known state ("clean bit"), or be given to the circuit in an unknown state ("dirty bit") that must be restored before the circuit finishes.
+Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, because any temporarily unused qubit is a borrowable dirty bit.
 
-The ability to borrow dirty ancilla from other parts of the circuit makes circuit constructions that require only dirty ancilla flexible, especially when space-constrained or when proper ancilla may be far away [[[(e.g.
-surface code)]]].
-For example, [haner2016] used dirty ancilla to cut the cost of performing Shor's algorithm on an $n$-bit number, using only $2n+2$ qubits, from ???? to $O(n^3)$ depth and $O(n^3 \lg n)$ size.
-In this paper we improve on [haner2016]'s result by cutting the number of required qubits from $2n+2$ to $2n+1$.
+The ability for part of a circuit to borrow dirty bits from another part of the same circuit makes circuit constructions that require only dirty bits very flexible.
+Especially in situations with space constraints, or on circuit topologies where clean ancilla are too far away to be acquired quickly.
+For example, \cite{haner2016} improves the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size by using an offset ("constant addition") construction that borrows dirty bits from one register while offsetting the other.
+We improve on \cite{haner2016} by cutting the number of required qubits from $2n+2$ to $2n+1$.
 
@@
 
-\section{$O(N)$ inline adder}
+\section{$O(N)$ inline adders}
 
-[van2004] describes a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla.
-The ancilla must be off, or else the result is incremented into returning $a+b+1$ instead of $a+b$.
+In \cite{van2004}, a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla, is described.
+If the ancilla is not initialized in the off state, the target register ends up storing $a+b+1$ instead of $a+b$.
 
-The need for the ancilla to be clean can be fixed by conditionally NOT-ing the source and target registers before and after the addition when the carry is on.
-Because $\lnot x = -x-1$, temporarily inverting the target turns the addition into a subtraction.
-Temporarily inverting the source turns the subtraction back into an addition, but also introduces an extra shift by -1.
-The -1 cancels the carry's +1 effect.
+The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the source and target registers before and after the addition when the carry is on.
+Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
+Temporarily inverting the source turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
 
-On top of this improvement, the carry role can be merged into the most-significant-bit of the source register.
-Although the MSB can't invert itself, the circuit still works because we don't need an overflow signal that distinguishes between the MSB toggling from on to off instead of from off to on.
-
-The improved reversible adder is shown in figure \ref{fig:inlineadder}.
-Note that, because the source register is used as workspace, it must be at least as large as the target register.
+On top of allowing the ancilla to be dirty, the ancilla's role can be merged into the most-significant-bit (MSB) of the source register.
+As shown in figure \ref{fig:inlineadder}, addition of same-sized registers can be performed {\em inline}, with no ancilla at all!
 
@@
   \includegraphics[totalheight=3cm]{inline-adder.png}
-  \caption{\em Adder using no ancilla and $\Theta(N)$ CNOTs and CSWAPs. Based on [van2004].}
+  \caption{\em Same-register-size adder using no ancilla, $6N-3$ CNOTs, and $2N-2$ CSWAPs. Based on \cite{van2004}.}
   \label{fig:inlineadder}
@@
 
+Because the source register is used as workspace, the construction in figure \ref{fig:inlineadder} only works when the source register is the same size as the target register.
+When the source register is larger, the issue is easily avoided by ignoring the high bits of the source register.
+When the target register is larger, the workaround is more difficult.
+Not only is there insufficient workspace, the register sizes being different breaks the logical-negation trick that allowed the carry signal bit to be dirty.
+
+The problems caused by the target register being larger can be overcome by using the increment and decrement constructions from the next section.
+(Note that they use only the same-size adder.)
+The carry signal can be forwarded into the high part of the target register with a controlled increment.
+The MSB of the source register can be free'd for use as a borrowed dirty bit by performing the MSB's controlled increment.
+Finally, the extra increment caused by the borrowed dirty bit being on can be cancelled with a controlled decrement.
+See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
+
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=3cm]{inline-adder-into-large.png}
+  \caption{\em
+      Adder with target larger than source, using no ancilla.
+      Uses increment constructions that use the same-size adder.}
+  \label{fig:inline-adder-into-large}
+\end{figure}
+
 \section{$O(N)$ incrementer}
 
-Note that a register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it for any $x$.
-When $N$ dirty ancilla are available, this can be done directly as shown in figure \ref{fig:double-sub-increment}.
+A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
+When $N$ dirty bits are available, $x$ comes from the register defined by those $N$ bits, as shown in figure \ref{fig:double-sub-increment}.
 
@@
 
-An increment gate can be decomposed into halves where the top half is incremented only if all of the bottom bits are on, and then the bottom half is incremented.
-By borrowing the top half, the bottom half can be incremented with the double-subtraction trick.
-But the trick doesn't work on the top half, because the top half can't be borrowed when being controlled by it.
+To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
+A top half that is incremented only if all of the bottom bits are on, and a bottom half that is unconditionally incremented.
+The bottom half can be incremented with the double-subtraction trick by borrowing the top half.
+But the trick doesn't work on the top half, because the bottom half can't be borrowed when it is being used as a control.
 
-A workaround for this is to change the second subtraction to an addition, and to conditionally toggle the target bits instead of the input bits.
-When the condition isn't satisfied, the addition and subtraction cancel each other.
-When the condition is satisfied, the addition inverts into a second subtraction and the input is subtracted from the target twice.
-However, because we are conditioning on the input bits all being on, the input must be -1 and the target must have been incremented by 2.
-To halve the +2 into a +1, left-shift the target and use a dirty bit as its LSB.
+To work around not being able to operate on the borrowed bits while them as a control, start with the double-subtraction trick but change the second subtraction to an addition and conditionally toggle the target bits instead of the input bits.
+When the condition isn't satisfied, the addition and subtraction will cancel each other.
+When the condition is satisfied, the addition is inverted into a subtraction.
+Both the subtraction and addition-turned-subtraction fire, subtracting the input register from the target register twice.
+However, because we are conditioning on the input bits all being on, the input must be -1.
+Therefore the target was incremented by 2.
+To halve the +2 into a +1, prepend a dirty LSB onto the target register as shown in \ref{fig:compact-increment}.
 
@@
   \includegraphics[totalheight=1.5cm]{compact-increment.png}
-  \caption{\em Odd-sized $\Theta(N)$ increment with 1 dirty ancilla.}
+  \caption{\em Odd-sized $O(N)$ increment with 1 dirty ancilla.}
   \label{fig:compact-increment}
@@
 
-The described incrementer construction is shown in figure \ref{fig:compact-increment}.
-However, because the adder/subtracter construction requires the target and source registers to be the same size, this incrementer construction only works on odd-sized registers.
+The incrementer construction shown in .
+Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the construction only works on odd-sized registers.
 For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
 The construction is shown in \ref{fig:compact-increment-even}.
-Note that it is easily modified to perform efficient increments with arbitrarily many controls.
 
@@
   \includegraphics[totalheight=2cm]{compact-increment-even.png}
-  \caption{\em Even-sized $\Theta(N)$ increment with 1 dirty ancilla.}
+  \caption{\em Even-sized $O(N)$ increment with 1 dirty ancilla.}
   \label{fig:compact-increment-even}
@@
 As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting all of the controls.
+Also, increment gates can be controlled by using another invert-input-and-output trick (see figures \ref{} and \ref{}).
 
-Our incrementer is classically optimal in asymptotic size and number of ancilla, but not depth.
-The size is asymptotically optimal because any incrementer must touch all $n$ bits.
-The number of ancilla is classically optimal, assuming we want to reduce to a fixed-size gate set, because the parity of the permutation performed by an increment operation is odd but the parity of the permutation performed by a classical gate that doesn't cover the entire circuit is even.
-Therefore, without a dirty bit to use as workspace, no classical circuit can perform an increment gate.
+Note that, unlike with addition, it is impossible to create a classical incrementer that uses no ancilla.
+The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
 Quantumly, the parity barrier can be bypassed by using partial rotations (see figure \ref{fig:bootstrap-ancilla}).
@@
   \centering
-  \includegraphics[totalheight=1.3cm]{ancilla-bootstrap.png}
-  \caption{\em Bootstrapping dirty ancilla out of an increment gate using quantum operations.}
+  \includegraphics[totalheight=1.4cm]{ancilla-bootstrap.png}
+  \caption{\em Bootstrapping dirty ancilla out of an increment gate using quantum operations.
+  The $i^{\text{frac}}$ operation is implemented by a column of $Z^{2^{-k}}$ gates.
+  The 'frac' refers to the fraction $v/2^d$, where $v$ is the computational-basis value in the register and $d$ is the size of the register.}
   \label{fig:bootstrap-ancilla}
@@
 
-\section{$O(N \lg N)$ constant $\text{PivotFlip}$}
+\section{$O(N)$ inline comparisons}
+
+Comparison operations toggle a target bit based on the relationship between two input registers.
+
+The is-B-less-than-A comparison is implemented by subtracting $A$ out of $B$ with the target bit appended to $B$, then adding $A$ into $B$ without the target bit appended.
+See figure \ref{fig:comparison-less}.
+
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=3cm]{comparison-less.png}
+  \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
+  \label{fig:comparison-less}
+\end{figure}
+
+The same idea is used for is-B-less-than-or-equal-to-A, except $A+1$ is subtracted and added instead of $A$.
+The extra $\pm 1$ is performed by increment and decrement gates, as shown in figure \ref{fig:comparison-leq}.
+When the two registers are the same size, the increment gate can be omitted by inverting the $A$ and $B$ registers during the addition.
+
+$>$ and $\geq$ comparisons are done by performing the opposite comparison ($\leq$ and $<$ respectively) and inverting the result by toggling the target.
+To perform a comparison when the $B$ register is smaller than the $A$ register, swap their roles and use the opposite comparison.
+
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=3cm]{comparison-leq.png}
+  \caption{\em Toggling a target bit if one register is less than or equal to another, without using ancilla.}
+  \label{fig:comparison-leq}
+\end{figure}
+
+\section{$\text{PivotFlip}$}
 
@@
 For example, a pivot-flip with the pivot equal to 3 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
-We will use pivot-flips later, when implementing modular addition.
+Pivot-flips are useful for implementing cyclic operations, such as modular addition.
 
-For concreteness, the exact permutation of a pivot-flip is:
+The exact permutation performed by a pivot-flip with pivot equal to $k$ is:
 
@@
 
+And the overall permutation, including the pivot input, is:
+
+$$\text{PivotFlip} = \sum_{k=0}^{N-1} |k\rangle \langle k| \otimes \text{PivotFlip}_{k}$$
+
 To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
-However, instead of doing nothing to the elements past the pivot, it also flips those elements as shown in figure \ref{fig:double-flip}.
+As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a)$ flips below the pivot but also above the pivot.
+However, because $x \rightarrow \lnot(x - a)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the input can be controlled by dirty comparisons against the pivot.
+By applying $x \rightarrow \lnot(x - a)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved (see figure \ref{fig:const-pivot-flip}).
+
+For states that are less than or equal to $A$, only one of the controlled flips happens.
+For other states, the controlled flip happens no times or two times, undoing itself.
 
@@
   \centering
-  \includegraphics[totalheight=7cm]{double-flip.png}
-  \caption{\em Subtraction followed by logical negation performs a flip both above and below the pivot.}
+  \includegraphics[totalheight=4cm]{double-flip.png}
+  \caption{\em Subtraction followed by logical negation performs flips both above and below $k$.}
   \label{fig:double-flip}
@@
 
-Note that $x \rightarrow \lnot(x - a)$ is its own inverse and maintains the invariant that a given value will cause an underflow when subtracting $a$.
-So, by applying the operation twice, controlled by a dirty bit that is toggled by the underflow signal, the desired effect is achieved.
-The part of the range below the pivot does trigger the underflow, so only one of the controlled flips happens.
-The part of the range above the pivot does not cause an underflow, and so flips no times or two times, undoing itself.
-
-The resulting pivot-flip circuit is shown in figure \ref{fig:const-pivot-flip}.
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=2.7cm]{pivot-flip.png}
+  \caption{\em Pivot flip circuit.
+  When the top wire bundle is a register, the subtractions are controlled by replacing the CNOTs in figure \ref{fig:inlineadder} with CCNOTs.
+  When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
+  \label{fig:const-pivot-flip}
+\end{figure}
 
@@
   \centering
-  \includegraphics[totalheight=2cm]{pivot-flip.png}
-  \caption{\em Pivot flip circuit.
-  Note that the top wire bundle is a compile-time constant.
-  The underflow toggle is done with an $n+1$-bit subtraction then an $n$-bit addition.
-  [[[Requires an extra dirty bit for the n+1-bit subtraction?]]]
-  The controlled subtractions are done by controlling every generated operation.}
-  \label{fig:const-pivot-flip}
+  \includegraphics[totalheight=3.1cm]{controlled-offset.png}
+  \caption{\em Offset circuit from \cite{haner2016}, modified to include a control.}
+  \label{fig:controlled-offset}
 \end{figure}
 
-Because the pivot is a compile-time constant, and so there is no source register workspace, the inline adder construction can't be used.
-Instead, we use the constant-adder from [haner2016].
-That adder requires $O(N)$ depth and $O(N \lg N)$ size, and therefore so does the constant-pivot-flip circuit.
+The cost of a $\text{PivotFlip}$ depends on if the pivot is a compile-time constant or not.
+When the pivot is input from another register, the $O(N)$ inline adder can be used.
+But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead (note: they refer to it as a constant addition gate).
 
-\section{$O(N \lg N)$ Modular Constant-Adder}
+\section{$O(N \lg N)$ Modular Addition and Offset}
 
 A modular addition is three pivot flips.
-To add $a$ modulo $n$, perform a pivot-flip at $n-a$, a pivot-flip at $n$, and then a pivot-flip at $a$.
+To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a$.
+See figure \ref{fig:modular-add}.
 
-[[[diagram]]]
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=1.7cm]{modular-addition.png}
+  \caption{\em Modular addition circuit.
+  The $+r$ offset gates require $O(N \lg N)$ size.
+  The pivot-flips require a dirty bit and that a constant.}
+  \label{fig:modular-add}
+\end{figure}
 
-Because the modular addition does a constant number of pivot-flips, its resource usage is identical: one [[[two?]]] dirty bits, $O(N)$ depth, and $O(N \lg N)$ size.
+Note that we have implemented modular addition in terms of normal addition.
+This fact is obfuscated by the number of constructions used but, as shown in figure \ref{fig:modular-dependencies}, addition and offset (constant addition) are the only operations not defined in terms of another operations.
 
-\section{$O(N)$ $\text{PivotFlip}$}
-
-For completeness, we also discuss how to do a pivot-flip when the pivot is stored in a register instead of a compile-time constant.
-The original construction doesn't work in this case because it uses an extra bit in the source register.
-
-To work around the problem we propagate the increment into the operations on the target.
-[[[But the addition sizes don't match up?]]]
-
-[[[Maybe just ignore the increment size problem, then fixup with a conditioned all-flip?]]]
-
-Note that, when storing in a register, we offset the pivot $k$ by one.
-This is to avoid the redundancy $\text{PivotFlip}_0 = \text{PivotFlip}_1 = I$ and to move the flip-all operation $\text{PivotFlip}_{N}$ into range.
-Concretely, we want to perform this permutation:
-
-$$\text{RegisterPivotFlip} = \sum_{k=0}^{N-1} |k\rangle \langle k| \otimes \text{PivotFlip}_{k+1}$$
-
-Basically we do this by just inlining the increments and decrements into the addition operations or by using that other thing I tried.
-
-\section{$\Theta(N)$ Modular Addition}
-
-Same high-level construction as the constant modular adder.
-Do three pivot flips, but register-based instead of constant-based.
-This reduces the size cost from $O(N \lg N)$ to $O(N)$.
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=18cm]{modular-add-graph.png}
+  \caption{\em
+      Dependency graph of constructions used to create a classical reversible modular adder / modular offset gates using only 1 dirty bit.}
+  \label{fig:modular-dependencies}
+\end{figure}
 
@@
 
-References:
-
-haner2016 https://arxiv.org/abs/1611.07995
-Factoring using 2n+2 qubits with Toffoli based modular multiplication
-Thomas Häner, Martin Roetteler, Krysta M. Svore
-
-van2004
-Optimal Design of A Reversible Full Adder
-YVAN VAN RENTERGEM AND ALEXIS DE VOS
-2004
+\bibliographystyle{plain}
+\bibliography{citations}
 

### 2017-03-10 paper.tex
Half draft.

@@
 \title{Efficient Reversible Arithmetic Circuits using a Dirty Bit}
-\author{et al}
+\author{Craig Gidney}
 
@@
 \begin{abstract}
-We describe classical and quantum circuits for performing reversible arithmetic using at most 1 ancilla in an unknown state.
-Our constructions can perform incrementation, comparisons, addition, and modular addition with classical circuits of linear size and depth.
-We also show how to perform modular offset, improving the best known space requirements for Shor's algorithm from $2n+2$ qubits to $2n+1$ qubits, in linear depth and $O(n \lg n)$ size.
+\em
+We present efficient quantum and reversible classical circuits for performing incrementation, addition, comparisons, modular addition, and modular offset using no ancilla or 1 ancilla in an unknown state.
+We use these circuits to improve the best known space requirements for Shor's algorithm from $2n+2$ qubits to $2n+1$ qubits, without increasing the depth or size complexity.
+[[[FIND OTHER APPLICATIONS]]]
 \end{abstract}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=4cm]{fourier-addition.png}
+  \caption{\em Controlled-increment from fourier addition. [[[NOT MENTIONED]]] [[[CITE THE PAPER THAT INTRODUCED THE IDEA]]]}
+  \label{fig:fourier-addition}
+\end{figure}
+
 To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
@@
 As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting all of the controls.
-Also, increment gates can be controlled by using another invert-input-and-output trick (see figures \ref{} and \ref{}).
+Also, increment gates can be controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).
+
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=2cm]{controlled-increment-odd.png}
+  \caption{\em Odd-sized $O(N)$ controlled increment with 1 dirty ancilla.}
+  \label{fig:controlled-increment-odd}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[totalheight=2.5cm]{controlled-increment-even.png}
+  \caption{\em Even-sized $O(N)$ controlled increment with 1 dirty ancilla.}
+  \label{fig:controlled-increment-even}
+\end{figure}
 
@@
 
-\section{$O(N \lg N)$ Modular Addition and Offset}
+\section{$O(N \lg N)$ Modular Addition and Modular Offset}
 
@@
   The $+r$ offset gates require $O(N \lg N)$ size.
-  The pivot-flips require a dirty bit and that a constant.}
+  The pivot-flips require a dirty bit.}
   \label{fig:modular-add}
@@
 
+[[[Surely there are more? Need to look around.]]]
+
+\section{Conclusion}
+
+[[[TODO]]]
+
 \bibliographystyle{plain}
@@
 
+[[[Surely there are more?]]]
+
+[[[Previous paper constructing each type of circuit shown here, and how they compare.]]]
+
 \end{document}

### 2017-03-11 paper.tex
I'd call it a first draft, but there's still holes.

@@
 \usepackage{graphicx}
-\title{Efficient Reversible Arithmetic Circuits using a Dirty Bit}
+\usepackage{hyperref}
+\title{Reversible Modular Addition using a Dirty Bit}
 \author{Craig Gidney}
@@
 \em
-We present efficient quantum and reversible classical circuits for performing incrementation, addition, comparisons, modular addition, and modular offset using no ancilla or 1 ancilla in an unknown state.
-We use these circuits to improve the best known space requirements for Shor's algorithm from $2n+2$ qubits to $2n+1$ qubits, without increasing the depth or size complexity.
-[[[FIND OTHER APPLICATIONS]]]
+We present reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using no ancilla or 1 ancilla in an unknown state.
+We improve the best known space requirements for factoring an $n$-bit number from $2n+2$ qubits to $2n+1$ qubits, without increasing the asymptotic depth or size of the circuit.
 \end{abstract}
@@
 When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
-An ancilla qubit is an extra qubit, on top of the qubits used for input and output, available for use as temporary workspace.
-Ancilla may be initialized to a known state ("clean bit"), or be given to the circuit in an unknown state ("dirty bit") that must be restored before the circuit finishes.
+An ancilla is an extra bit or qubit, on top of those used for input and output, available for use as temporary workspace.
+Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
 Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, because any temporarily unused qubit is a borrowable dirty bit.
@@
 Especially in situations with space constraints, or on circuit topologies where clean ancilla are too far away to be acquired quickly.
-For example, \cite{haner2016} improves the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size by using an offset ("constant addition") construction that borrows dirty bits from one register while offsetting the other.
+For example, \cite{haner2016} improves the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size by introducing an offset gate that borrows dirty bits from one register while offsetting the other.
 We improve on \cite{haner2016} by cutting the number of required qubits from $2n+2$ to $2n+1$.
@@
 The paper is structured as follows.
-Sections ??? through ??? describe various circuit constructions.
-Section ??? applies those circuit constructions to exist problems.
-... That's it.
-
-[[[What about underlying details, like reducing many-controlled many-nots?
-Nearby details like bootstrapping an ancilla with quantum operations?
-Parity issues that prevent some operations from being performed with no ancilla?]]]
+The introduction describes the various sections and also itself for no discernible reason.
+Section \ref{sec:review} discusses existing circuit constructions.
+In sections \ref{sec:first-circuit} through \ref{sec:last-circuit} we describe our various arithmetic circuit constructions.
+Section \ref{sec:applications} discusses how to use our constructions to factor an $n$ bit number using $2n+1$ qubits.
+Section \ref{sec:conclusion} concludes.
 
@@
 
-\section{$O(N)$ inline adders}
+\section{Existing Constructions} \label{sec:review}
+
+\subsection{Addition}
+
+\cite{draper2000} describes an addition circuit that uses the quantum Fourier transform to convert phasing into shifting.
+Their naive construction uses $O(N)$ depth and $O(N^2)$ controlled rotations, but this can be improved to $O(\lg N)$ depth and $O(N \lg N)$ size by using approximation techniques \cite{barenco1996} \cite{cleve2000}.
+
+\cite{cuccaro2004} and \cite{van2004} describe classical reversible adders that use a single clean ancilla, $O(N)$ depth, and $O(N)$ size.
+
+\cite{takahashi2005} and \cite{takahashi2009} describe classical reversible adders that use no ancilla.
+
+We improve on \cite{takahashi2005} by allowing addition into larger registers.
+
+\subsection{Offset}
+
+An offset gate adds a compile-time constant into a register.
+Many reversible adder constructions do not work in this case because there is no input register to use as temporary workspace.
+
+\cite{draper2000}'s QFT-based addition doesn't need the source register as workspace, and so can be applied to this problem.
+
+\cite{haner2016} describes a divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $1$ dirty ancilla (or 0 if using quantum operations).
+
+This paper describes a circuit for the special case of offset-by-1, i.e. incrementation, that uses $O(N)$ depth and size and a single dirty ancilla.
+We also show how to remove the ancilla with quantum operations.
+The construction is a simplified version of the incrementer described in \cite{gidney2015}.
+
+\subsection{Comparison}
+
+\cite{takahashi2005} describes inline comparison operations.
+We improve on this construction by allowing the comparison of registers of different sizes.
+
+\subsection{Modular Addition}
+
+\cite{vedral1995} describes a modular adder that uses $O(N)$ clean ancilla, based on performing normal addition and subtraction and temporarily storing a comparison in a clean ancilla.
+Most of the clean ancilla are only used for the addition and subtraction, so using inline adders reduces the number of clean ancilla required to 1, as used in \cite{haner2016}.
+
+We improve on this construction by allowing the ancilla to be dirty.
+
+\section{$O(N)$ inline adders} \label{sec:first-circuit}
+
+[[[ Should I cut most of this? Cited papers do just as well. ]]]
 
@@
 
-The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the source and target registers before and after the addition when the carry is on.
+The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the source and target registers before and after the addition when the ancilla is on.
 Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
@@
 
-On top of allowing the ancilla to be dirty, the ancilla's role can be merged into the most-significant-bit (MSB) of the source register.
-As shown in figure \ref{fig:inlineadder}, addition of same-sized registers can be performed {\em inline}, with no ancilla at all!
+On top of allowing the ancilla to be dirty, the ancilla's role can be merged into the most-significant-bit (MSB) of the source register as shown in figure \ref{fig:inlineadder}.
 
@@
 When the target register is larger, the workaround is more difficult.
-Not only is there insufficient workspace, the register sizes being different breaks the logical-negation trick that allowed the carry signal bit to be dirty.
+Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the carry signalling bit to be dirty.
 
 The problems caused by the target register being larger can be overcome by using the increment and decrement constructions from the next section.
-(Note that they use only the same-size adder.)
+(Note that they use only the same-register-size adder, not the adder about to be defined.)
 The carry signal can be forwarded into the high part of the target register with a controlled increment.
@@
 A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
-When $N$ dirty bits are available, $x$ comes from the register defined by those $N$ bits, as shown in figure \ref{fig:double-sub-increment}.
+When $N$ dirty bits are available, $x$ can come from the register defined by those $N$ bits, as shown in figure \ref{fig:double-sub-increment}.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[totalheight=4cm]{fourier-addition.png}
-  \caption{\em Controlled-increment from fourier addition. [[[NOT MENTIONED]]] [[[CITE THE PAPER THAT INTRODUCED THE IDEA]]]}
-  \label{fig:fourier-addition}
-\end{figure}
-
 To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
@@
 
-To work around not being able to operate on the borrowed bits while them as a control, start with the double-subtraction trick but change the second subtraction to an addition and conditionally toggle the target bits instead of the input bits.
+To work around not being able to operate on the borrowed bits while using them as a control, start with the double-subtraction trick but change the second subtraction to an addition and conditionally toggle the target bits instead of the input bits.
 When the condition isn't satisfied, the addition and subtraction will cancel each other.
@@
 
-The incrementer construction shown in .
 Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the construction only works on odd-sized registers.
 For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
-The construction is shown in \ref{fig:compact-increment-even}.
+The construction is shown in figure \ref{fig:compact-increment-even}.
 
@@
 
-Note that, unlike with addition, it is impossible to create a classical incrementer that uses no ancilla.
+Note that, unlike with register-to-register addition, it is impossible to create a classical incrementer that uses no ancilla.
 The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
@@
 
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
-As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a)$ flips below the pivot but also above the pivot.
-However, because $x \rightarrow \lnot(x - a)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the input can be controlled by dirty comparisons against the pivot.
-By applying $x \rightarrow \lnot(x - a)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved (see figure \ref{fig:const-pivot-flip}).
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x + \lnot a)$ nearly does what is required.
+As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x + \lnot a)$ flips the range below-and-including the pivot but also flips the range above the pivot.
+However, because $x \rightarrow \lnot(x + \lnot a)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the operation can be restricted by dirty comparisons against the pivot.
 
+By applying $x \rightarrow \lnot(x + \lnot a)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved.
 For states that are less than or equal to $A$, only one of the controlled flips happens.
 For other states, the controlled flip happens no times or two times, undoing itself.
+See figure \ref{fig:const-pivot-flip}, which also includes a variant for pivot-flipping while excluding the pivot.
 
@@
   \centering
-  \includegraphics[totalheight=2.7cm]{pivot-flip.png}
-  \caption{\em Pivot flip circuit.
+  \includegraphics[totalheight=4.5cm]{pivot-flip.png}
+  \caption{\em Pivot flip circuits.
   When the top wire bundle is a register, the subtractions are controlled by replacing the CNOTs in figure \ref{fig:inlineadder} with CCNOTs.
@@
 When the pivot is input from another register, the $O(N)$ inline adder can be used.
-But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead (note: they refer to it as a constant addition gate).
+But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead.
 
-\section{$O(N \lg N)$ Modular Addition and Modular Offset}
+\section{$O(N \lg N)$ Modular Addition and Modular Offset} \label{sec:last-circuit}
 
 A modular addition is three pivot flips.
-To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a$.
+To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a-1$.
 See figure \ref{fig:modular-add}.
@@
   \centering
-  \includegraphics[totalheight=1.7cm]{modular-addition.png}
+  \includegraphics[totalheight=1.4cm]{modular-addition.png}
   \caption{\em Modular addition circuit.
@@
   \centering
-  \includegraphics[totalheight=18cm]{modular-add-graph.png}
+  \includegraphics[totalheight=16cm]{modular-add-graph.png}
   \caption{\em
@@
 
-\section{Applications}
+\section{Factoring with $2n+1$ qubits} \label{sec:applications}
 
-Shor's algorithm in $2n+1$ qubits.
+\cite{haner2016} describes how to perform Shor's algorithm with Toffoli-based circuits using $2n+2$ qubits.
+The first $n$ qubits are needed to hold the input.
+The second $n$ qubits are for holding the output of exponentiating the input modulo a constant.
+One of the remaining qubits is used for the phase estimation process.
+The last qubit is a clean ancilla used by modular addition.
 
-[[[Surely there are more? Need to look around.]]]
+The modular addition circuit we have presented does not require a clean ancilla.
+The need for the last ancilla is avoided by instead borrowing the phase estimation qubit as a temporary dirty bit.
 
-\section{Conclusion}
+[[[What about underlying details I didn't mention, like reducing many-controlled many-nots?]]]
+
+[[[ Maybe some more details...? ]]]
+
+\section{Conclusion} \label{sec:conclusion}
 
@@
 
-[[[Surely there are more?]]]
+\section{Test Circuit Links}
 
-[[[Previous paper constructing each type of circuit shown here, and how they compare.]]]
+Convenience links for myself when checking that things work.
+
+\href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
+{5+5-bit inline adder}
+(or \href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
+{with test harness})
+
+\href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
+{4-bit incrementer w/ 4 dirty}
+
+\href
+{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%5D%7D}
+{8-bit incrementer w/ 1 dirty}
+(or \href
+{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting9%22%5D%2C%5B%22QFT6%22%5D%2C%5B1%2C1%2C1%2C%22QFT6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec8%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A06%22%5D%2C%5B%22QFT%E2%80%A06%22%5D%2C%5B%22Uncounting9%22%5D%5D%7D}
+{with test harness})
+
+\href
+{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%5D%7D}
+{9-bit incrementer w/ 1 dirty}
+(or \href
+{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec9%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
+{with test harness})
+
+\href
+{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9B%22%2C%22Z%5E%C2%BC%22%2C%22H%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22inc4%22%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B1%2C%22Z%5E-%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E-%E2%85%9B%22%2C%22Z%5E-%C2%BC%22%2C%22H%22%5D%5D%7D}
+{Quantum incrementer ancilla bootstrap}
+
+NOTE: have to open these next ones against local version, because online version doesn't have the comparison gates
+
+\href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
+{4+4-Bit $<$ PivotFlip}
+(or \href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
+{with test harness})
+
+\href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
+{4+4-Bit $\leq$ PivotFlip}
+(or \href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
+{with test harness})
+
+\href
+{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B%22X%22%2C1%2C1%2C%22X%22%2C%22Counting4%22%2C1%2C1%2C1%2C%22QFT4%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%7C0%E2%9F%A9%E2%9F%A80%7C%22%5D%2C%5B%22inputB4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22~a82g%22%5D%5D%2C%22gates%22%3A%5B%7B%22id%22%3A%22~a82g%22%2C%22name%22%3A%22IGNORE%22%2C%22matrix%22%3A%22%7B%7B1%2C0%7D%2C%7B0%2C1%7D%7D%22%7D%5D%7D}
+{Modular Addition with test Harness}
 

### 2017-03-11 paper.tex
First draft

@@
 \em
-We present reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using no ancilla or 1 ancilla in an unknown state.
-We improve the best known space requirements for factoring an $n$-bit number from $2n+2$ qubits to $2n+1$ qubits, without increasing the asymptotic depth or size of the circuit.
+We present size-efficient reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using no ancilla or 1 ancilla in an unknown state.
+We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ \cite{takahashi2006, haner2016} to $2n+1$, without increasing the asymptotic cost of the circuit.
 \end{abstract}
@@
 
-When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
-An ancilla is an extra bit or qubit, on top of those used for input and output, available for use as temporary workspace.
+When constructing quantum circuits, or classical reversible circuits, the number of available ancilla is an important resource.
+An ancilla is an extra bit or qubit, on top of those used in input and output roles, available for use as temporary workspace.
 Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
@@
 
-The ability for part of a circuit to borrow dirty bits from another part of the same circuit makes circuit constructions that require only dirty bits very flexible.
-Especially in situations with space constraints, or on circuit topologies where clean ancilla are too far away to be acquired quickly.
-For example, \cite{haner2016} improves the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size by introducing an offset gate that borrows dirty bits from one register while offsetting the other.
+In this paper we focus on circuits that use at most one dirty bit.
+
+Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
+For example, H\"{a}ner et al.
+used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size \cite{haner2016}.
 We improve on \cite{haner2016} by cutting the number of required qubits from $2n+2$ to $2n+1$.
 
-The paper is structured as follows.
+Our paper is structured as follows.
 The introduction describes the various sections and also itself for no discernible reason.
 Section \ref{sec:review} discusses existing circuit constructions.
-In sections \ref{sec:first-circuit} through \ref{sec:last-circuit} we describe our various arithmetic circuit constructions.
-Section \ref{sec:applications} discusses how to use our constructions to factor an $n$ bit number using $2n+1$ qubits.
-Section \ref{sec:conclusion} concludes.
+In sections \ref{sec:first-circuit} through \ref{sec:last-circuit} we describe various arithmetic circuit constructions that use 0 or 1 dirty bit, building up to modular addition.
+We then use our single-dirty-bit modular addition to factor an $n$ bit number using $2n+1$ qubits in section \ref{sec:applications}.
 
 All constructions assume a 2s-complement representation of integers.
-All circuit diagrams with multi-wire integer operations put the least significant bit towards the top.
+All circuit diagrams have the least significant bit towards the top.
+For clarity, circuit diagrams will divide compound operations into separate input and output parts, with inputs shaded gray.
 
@@
 
-\cite{draper2000} describes an addition circuit that uses the quantum Fourier transform to convert phasing into shifting.
-Their naive construction uses $O(N)$ depth and $O(N^2)$ controlled rotations, but this can be improved to $O(\lg N)$ depth and $O(N \lg N)$ size by using approximation techniques \cite{barenco1996} \cite{cleve2000}.
+In \cite{draper2000}, Draper describes an addition circuit that uses the quantum Fourier transform to convert controlled phasing into addition.
+Draper notes that the construction uses $O(N)$ depth and $O(N^2)$ controlled rotations, but approximation techniques \cite{barenco1996} \cite{cleve2000} can improve the depth and size to $O(\lg N)$ and $O(N \lg N)$ respectively.
 
-\cite{cuccaro2004} and \cite{van2004} describe classical reversible adders that use a single clean ancilla, $O(N)$ depth, and $O(N)$ size.
+Both \cite{cuccaro2004} and \cite{van2004} describe classical reversible ripple-carry adders that use a single clean ancilla, $O(N)$ depth, and $O(N)$ size.
+Takahashi \cite{takahashi2005} showed how to use the most-significant-bit (MSB) of the input register, instead of a separate ancilla, to propagate the carry signal.
+Takashi later showed how to trade depth against size \cite{takahashi2009}.
 
-\cite{takahashi2005} and \cite{takahashi2009} describe classical reversible adders that use no ancilla.
-
-We improve on \cite{takahashi2005} by allowing addition into larger registers.
+We generalize the ideas in \cite{takahashi2005} to apply to additions where the input register may be smaller than the target register.
 
@@
 An offset gate adds a compile-time constant into a register.
-Many reversible adder constructions do not work in this case because there is no input register to use as temporary workspace.
+Many reversible adder constructions do not work as offset gates because there is no input register to use as temporary workspace.
 
-\cite{draper2000}'s QFT-based addition doesn't need the source register as workspace, and so can be applied to this problem.
+Draper's QFT-based addition \cite{draper2000} doesn't need the source register as workspace, and so can be applied to this problem.
 
-\cite{haner2016} describes a divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $1$ dirty ancilla (or 0 if using quantum operations).
+\cite{haner2016} describes a divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $n/2$ dirty ancilla.
+[[[ Actually, only 1 is required by their construction.
+0 if you go quantum.
+And when I emailed them about it they said they would fix the oversight.
+So I'm not sure how to describe that.
+It is important for this paper that we have an non-modular offset circuit that only needs 1 dirty ancilla.
+]]]
 
-This paper describes a circuit for the special case of offset-by-1, i.e. incrementation, that uses $O(N)$ depth and size and a single dirty ancilla.
-We also show how to remove the ancilla with quantum operations.
-The construction is a simplified version of the incrementer described in \cite{gidney2015}.
+We describe a circuit for the special case of offset-by-1, i.e. incrementation.
+Our circuit uses $O(N)$ depth, $O(N)$ size, and a single dirty ancilla.
+We also note how to remove the ancilla with quantum operations.
+Our construction is a simplified version of our unpublished incrementer described in \cite{gidney2015}.
 
@@
 
-\cite{takahashi2005} describes inline comparison operations.
-We improve on this construction by allowing the comparison of registers of different sizes.
+Comparison operators, which toggle a target bit based on if one input is larger, smaller, and/or equal to another, can be implemented efficiently using adders \cite{takahashi2005}.
+Due to generalizing Takahashi's adder \cite{takahashi2005} to apply to different sizes, we also generalize Takahashi's comparators to apply to registers of different sizes.
 
@@
 
+[[[ do better here ]]]
+
 We improve on this construction by allowing the ancilla to be dirty.
@@
 
-[[[ Should I cut most of this? Cited papers do just as well. ]]]
-
 In \cite{van2004}, a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla, is described.
 If the ancilla is not initialized in the off state, the target register ends up storing $a+b+1$ instead of $a+b$.
+For clarity, we review how to elide the ancilla before generalizing the construction to apply to registers of differing size.
 
-The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the source and target registers before and after the addition when the ancilla is on.
+The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the input and target registers before and after the addition when the ancilla is on.
 Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
-Temporarily inverting the source turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
+Temporarily inverting the input turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
 
-On top of allowing the ancilla to be dirty, the ancilla's role can be merged into the most-significant-bit (MSB) of the source register as shown in figure \ref{fig:inlineadder}.
+Once the ancilla can be dirty the MSB of the input register can play the role of the ancilla, as shown in figure \ref{fig:inlineadder}.
 
@@
   \includegraphics[totalheight=3cm]{inline-adder.png}
-  \caption{\em Same-register-size adder using no ancilla, $6N-3$ CNOTs, and $2N-2$ CSWAPs. Based on \cite{van2004}.}
+  \caption{\em Adder for input and target registers of the same size.
+  Requires no ancilla, and uses $O(N)$ gates and depth.
+  Based on \cite{van2004, takahashi2005}.}
   \label{fig:inlineadder}
@@
 
-Because the source register is used as workspace, the construction in figure \ref{fig:inlineadder} only works when the source register is the same size as the target register.
-When the source register is larger, the issue is easily avoided by ignoring the high bits of the source register.
-When the target register is larger, the workaround is more difficult.
-Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the carry signalling bit to be dirty.
+Because this adder construction uses the input register as workspace, it only works when the input register is the same size as the target register.
+When the input register is larger the high bits of the input register can simply be ignored, but when the target register is larger the workaround is more difficult.
+Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the ancilla to be dirty.
 
-The problems caused by the target register being larger can be overcome by using the increment and decrement constructions from the next section.
-(Note that they use only the same-register-size adder, not the adder about to be defined.)
-The carry signal can be forwarded into the high part of the target register with a controlled increment.
-The MSB of the source register can be free'd for use as a borrowed dirty bit by performing the MSB's controlled increment.
-Finally, the extra increment caused by the borrowed dirty bit being on can be cancelled with a controlled decrement.
-See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
+To work around the target register being larger, three fixes are needed.
+Each uses a controlled increment or decrement gate, which will be described in the next section.
+(Note that, to avoid cyclic dependencies, the increment and decrement constructions will only use the same-register-size adder.)
+First, the carry signal must be forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
+Second, to free the MSB of the input register for use as the carry signal, we add it into the target ahead of time with a controlled increment.
+Finally, when the MSB is on instead of off its role as carry signal causes an extra increment of the target.
+This extra increment is cancelled with a controlled decrement.
+See the circuit diagram in figure \ref{fig:inline-adder-into-large} for the complete construction.
 
@@
       Adder with target larger than source, using no ancilla.
-      Uses increment constructions that use the same-size adder.}
+      Uses $O(N)$ gates and depth.
+      The increment and decrement gates all have at least one free wire to borrow as a dirty ancilla.}
   \label{fig:inline-adder-into-large}
@@
 A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
-When $N$ dirty bits are available, $x$ can come from the register defined by those $N$ bits, as shown in figure \ref{fig:double-sub-increment}.
+When $N$ dirty bits are available, $x$ can come from a register defined by those $N$ arbitrary bits, as shown in figure \ref{fig:double-sub-increment}.
 
@@
 The bottom half can be incremented with the double-subtraction trick by borrowing the top half.
-But the trick doesn't work on the top half, because the bottom half can't be borrowed when it is being used as a control.
+But the double-subtraction trick doesn't work on the top half, because the bottom half can't be borrowed when it is being used as a control.
 
@@
 When the condition isn't satisfied, the addition and subtraction will cancel each other.
-When the condition is satisfied, the addition is inverted into a subtraction.
-Both the subtraction and addition-turned-subtraction fire, subtracting the input register from the target register twice.
-However, because we are conditioning on the input bits all being on, the input must be -1.
+When the condition is satisfied, the addition will be inverted into a subtraction, and both the subtraction and addition-turned-subtraction will fire, subtracting the input register from the target register twice.
+However, because we are conditioning on the input bits all being on, we know the input must be -1.
 Therefore the target was incremented by 2.
@@
 
-Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the construction only works on odd-sized registers.
+Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the construction described so far only works on odd-sized registers.
 For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
-The construction is shown in figure \ref{fig:compact-increment-even}.
+That construction is shown in figure \ref{fig:compact-increment-even}.
 
@@
 As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting all of the controls.
-Also, increment gates can be controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).
+Also, increment gates can be efficiently controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).
 
@@
 
-Note that, unlike with register-to-register addition, it is impossible to create a classical incrementer that uses no ancilla.
+Note that it is impossible to create a classical incrementer that uses no ancilla.
 The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
@@
   \includegraphics[totalheight=1.4cm]{ancilla-bootstrap.png}
-  \caption{\em Bootstrapping dirty ancilla out of an increment gate using quantum operations.
-  The $i^{\text{frac}}$ operation is implemented by a column of $Z^{2^{-k}}$ gates.
-  The 'frac' refers to the fraction $v/2^d$, where $v$ is the computational-basis value in the register and $d$ is the size of the register.}
+  \caption{\em Bootstrapping a dirty ancilla out of an increment gate using quantum operations.
+  The $i^{\text{frac}}$ gate is a ``phase gradient" operation that phases each computational basis state $|v\rangle$ by an amount proportional to $v/2^d$, where $d$ is the size of the register.
+  In this case each state is phased by $e^{i \frac{\pi}{2} v/2^d}$.
+  The phase gradient is implemented by a column of $Z^{2^{-k}}$ gates, .}
   \label{fig:bootstrap-ancilla}
@@
 
-Comparison operations toggle a target bit based on the relationship between two input registers.
+We review how to implement comparison operations, which toggle a target bit based on the relationship between two input registers, using addition and subtraction operations.
 
-The is-B-less-than-A comparison is implemented by subtracting $A$ out of $B$ with the target bit appended to $B$, then adding $A$ into $B$ without the target bit appended.
+The is-B-less-than-A comparison is implemented by subtracting $A$ out of $B$, with a target bit appended as $B$'s MSB, then adding $A$ into $B$ without the target bit appended.
 See figure \ref{fig:comparison-less}.
@@
 The same idea is used for is-B-less-than-or-equal-to-A, except $A+1$ is subtracted and added instead of $A$.
-The extra $\pm 1$ is performed by increment and decrement gates, as shown in figure \ref{fig:comparison-leq}.
+The extra $\pm 1$ can be performed by increment and decrement gates, as shown in figure \ref{fig:comparison-leq}.
 When the two registers are the same size, the increment gate can be omitted by inverting the $A$ and $B$ registers during the addition.
 
-$>$ and $\geq$ comparisons are done by performing the opposite comparison ($\leq$ and $<$ respectively) and inverting the result by toggling the target.
-To perform a comparison when the $B$ register is smaller than the $A$ register, swap their roles and use the opposite comparison.
+$>$ and $\geq$ comparisons are done by performing the corresponding opposite comparison ($\leq$ and $<$ respectively) and inverting the result by toggling the target.
+
+Beware that, when the $B$ register is smaller than the $A$ register, the high bits of $A$ are not accounted for by the subtraction and addition and so the construction breaks.
+To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
+Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by all of the high bits of $A$ being off (and then, if any of those high bits are on, meaning $A$ must be larger than $B$, toggle the result bit as appropriate for the comparison).
 
@@
 
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x + \lnot a)$ nearly does what is required.
-As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x + \lnot a)$ flips the range below-and-including the pivot but also flips the range above the pivot.
-However, because $x \rightarrow \lnot(x + \lnot a)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the operation can be restricted by dirty comparisons against the pivot.
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a - 1)$ nearly does what is required.
+As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a - 1)$ flips the range below-and-including the pivot but also flips the range above the pivot.
+However, because $x \rightarrow \lnot(x - a - 1)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the operation can be controlled by dirty comparisons against the pivot.
 
-By applying $x \rightarrow \lnot(x + \lnot a)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved.
-For states that are less than or equal to $A$, only one of the controlled flips happens.
-For other states, the controlled flip happens no times or two times, undoing itself.
-See figure \ref{fig:const-pivot-flip}, which also includes a variant for pivot-flipping while excluding the pivot.
+By applying $x \rightarrow \lnot(x - a - 1)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved.
+For states that are less than or equal to the pivot $a$, the dirty ancilla is being toggled and exactly one of the controlled flips will happen.
+For other states, the ancilla does not change and so the controlled flip happens no times or two times, undoing itself.
+See figure \ref{fig:const-pivot-flip} fot the construction, as well as a variant that only flips values strictly less than the pivot instead of including the pivot.
 
@@
   \caption{\em Pivot flip circuits.
-  When the top wire bundle is a register, the subtractions are controlled by replacing the CNOTs in figure \ref{fig:inlineadder} with CCNOTs.
+  When the top wire bundle is a register, the subtractions are controlled by adding a control to any CNOTs, increments, and decrements in figures \ref{fig:inlineadder} or \ref{fig:inline-adder-into-large}.
   When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
@@
 
+Be aware that this construction requires that the input register be in range.
+If the input is greater than or equal to $r$, the pivot-flips will trash the target register instead of adding the input mod $r$ into it.
+
 \begin{figure}
@@
   \caption{\em Modular addition circuit.
-  The $+r$ offset gates require $O(N \lg N)$ size.
+  The $+r$ operations are offset gates requiring $O(N \lg N)$ size when the modulus is a constant.
   The pivot-flips require a dirty bit.}
@@
 
-Note that we have implemented modular addition in terms of normal addition.
-This fact is obfuscated by the number of constructions used but, as shown in figure \ref{fig:modular-dependencies}, addition and offset (constant addition) are the only operations not defined in terms of another operations.
+Although the fact is obfuscated by the number of constructions used, our modular addition construction is performed entirely in terms of normal addition (with various tweaks and tricks and inversions along the way).
+See figure \ref{fig:modular-dependencies} for the full dependency graph.
+
+The modular addition uses $O(N \lg N)$ gates because that is the cost of the offset gate in \cite{haner2016}, and we subtract the modulus (typically a compile-time constant) out of a register during one of the pivot-flips.
+If both the amount to add {\em and the modulus} are in registers, then the construction only requires $O(N)$ gates.
+The total depth is $O(N)$ regardless.
 
@@
 
-\cite{haner2016} describes how to perform Shor's algorithm with Toffoli-based circuits using $2n+2$ qubits.
+In \cite{haner2016}, H\"{a}ner et al. describe how to perform Shor's algorithm with Toffoli-based circuits using $2n+2$ qubits.
 The first $n$ qubits are needed to hold the input.
 The second $n$ qubits are for holding the output of exponentiating the input modulo a constant.
-One of the remaining qubits is used for the phase estimation process.
-The last qubit is a clean ancilla used by modular addition.
+One of the remaining qubits is used in the phase estimation process.
+The last qubit is a clean ancilla used by their modular offset operation to temporarily hold a comparison against the modulus.
 
-The modular addition circuit we have presented does not require a clean ancilla.
-The need for the last ancilla is avoided by instead borrowing the phase estimation qubit as a temporary dirty bit.
-
-[[[What about underlying details I didn't mention, like reducing many-controlled many-nots?]]]
-
-[[[ Maybe some more details...? ]]]
+The modular addition circuit we have presented does not require that last clean ancilla.
+Instead, we borrow the phase estimation qubit as a temporary dirty bit.
+Our modular offset circuit is a drop-in replacement for the one used by H\"{a}ner et all.
+It even has the same asymptotic costs: $O(N \lg N)$ size and $O(N)$ depth.
+Using our modular offset circuit, and borrowing the phase estimation qubit instead of using a separate clean bit, creates an implementation of Shor's factoring algorithm that uses only $2n+1$ qubits without increasing the size or depth complexity.
 
@@
 
-[[[TODO]]]
+We present classical reversible circuits for incrementing, modular addition, and modular offset that use only a single dirty bit.
+Our modular offset circuit improves the minimum number of qubits used when factoring an $n$-bit integer from $2n+2$ to $2n+1$.
+
+For incrementing, a single dirty ancilla is classically optimal because the permutation defined by an increment operation is odd.
+This obstacle does not apply to modular additions, because any odd parities can made even with a useless swap in the range of states above the modulus.
+So it may be possible to elide the final dirty ancilla our modular addition construction requires, even when restricted to Toffoli-based circuits.
 
@@
 
-\section{Test Circuit Links}
+\section{Temporary Test Circuit Links}
 

### 2017-03-28 paper.tex
Cleaning

@@
 \em
-We present size-efficient reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using no ancilla or 1 ancilla in an unknown state.
+We present size-efficient reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using either no ancilla or 1 ancilla in an unknown state.
 We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ \cite{takahashi2006, haner2016} to $2n+1$, without increasing the asymptotic cost of the circuit.
@@
 When constructing quantum circuits, or classical reversible circuits, the number of available ancilla is an important resource.
-An ancilla is an extra bit or qubit, on top of those used in input and output roles, available for use as temporary workspace.
+An ancilla is an extra bit or qubit available for use by a circuit as temporary workspace.
 Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
 Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, because any temporarily unused qubit is a borrowable dirty bit.
-
-In this paper we focus on circuits that use at most one dirty bit.
 
@@
 used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size \cite{haner2016}.
-We improve on \cite{haner2016} by cutting the number of required qubits from $2n+2$ to $2n+1$.
+The modular addition construction in this paper improves on \cite{haner2016} by cutting the number of required qubits to factor an $n$-bit number from $2n+2$ to $2n+1$.
 
@@
 In sections \ref{sec:first-circuit} through \ref{sec:last-circuit} we describe various arithmetic circuit constructions that use 0 or 1 dirty bit, building up to modular addition.
-We then use our single-dirty-bit modular addition to factor an $n$ bit number using $2n+1$ qubits in section \ref{sec:applications}.
+We then describe how to use our single-dirty-bit modular addition to factor an $n$ bit number using $2n+1$ qubits in section \ref{sec:applications}.
 
@@
 
-We generalize the ideas in \cite{takahashi2005} to apply to additions where the input register may be smaller than the target register.
+We generalize the ideas in \cite{takahashi2005} to apply to additions where the input and target registers have differing sizes.
 
@@
 An offset gate adds a compile-time constant into a register.
-Many reversible adder constructions do not work as offset gates because there is no input register to use as temporary workspace.
+Some reversible adder constructions can be used as offset gates, but others can't because they use qubits in the input register as temporary workspace.
+An offset gate has no input register.
 
-Draper's QFT-based addition \cite{draper2000} doesn't need the source register as workspace, and so can be applied to this problem.
+Draper's QFT-based addition construction \cite{draper2000} doesn't use the input register as workspace, and so trivially generalizes into an offset gate.
 
-\cite{haner2016} describes a divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $n/2$ dirty ancilla.
+\cite{haner2016} describes a Toffoli-based divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $n/2$ dirty ancilla.
 [[[ Actually, only 1 is required by their construction.
@@
 We describe a circuit for the special case of offset-by-1, i.e. incrementation.
-Our circuit uses $O(N)$ depth, $O(N)$ size, and a single dirty ancilla.
-We also note how to remove the ancilla with quantum operations.
-Our construction is a simplified version of our unpublished incrementer described in \cite{gidney2015}.
+Our increment circuit uses $O(N)$ depth, $O(N)$ size, one dirty ancilla, and no non-classical operations.
+We also note how to cut the ancilla with quantum operations.
+This construction is a simplified version of our unpublished incrementer described in \cite{gidney2015}.
 
@@
 Comparison operators, which toggle a target bit based on if one input is larger, smaller, and/or equal to another, can be implemented efficiently using adders \cite{takahashi2005}.
+
 Due to generalizing Takahashi's adder \cite{takahashi2005} to apply to different sizes, we also generalize Takahashi's comparators to apply to registers of different sizes.
@@
 
-\cite{vedral1995} describes a modular adder that uses $O(N)$ clean ancilla, based on performing normal addition and subtraction and temporarily storing a comparison in a clean ancilla.
-Most of the clean ancilla are only used for the addition and subtraction, so using inline adders reduces the number of clean ancilla required to 1, as used in \cite{haner2016}.
+\cite{vedral1995} describes a modular adder that uses $O(N)$ clean ancilla.
+It temporarily stores a comparison in a clean ancilla, performs a normal addition, uses a subtraction conditioned on the ancilla to wrap the result back into range, then performs a cleanup comparison to clear the ancilla.
 
-[[[ do better here ]]]
+All but one of the $O(N)$ clean ancilla used by \cite{vedral1995} are due to using inefficient addition circuits.
+Using later zero-ancilla adder constructions reduces xthe number of clean ancilla required to perform a modular addition to just one.
 
-We improve on this construction by allowing the ancilla to be dirty.
+Our modular adder construction also requires a single ancilla, but the ancilla is no longer required to be clean.
 
@@
 Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
-Temporarily inverting the input turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
+Temporarily inverting the input effectively turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
 
@@
 Because this adder construction uses the input register as workspace, it only works when the input register is the same size as the target register.
-When the input register is larger the high bits of the input register can simply be ignored, but when the target register is larger the workaround is more difficult.
+When the input register is larger, the high bits of the input register can simply be ignored.
+But, when the target register is larger, the workaround is more difficult.
 Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the ancilla to be dirty.
@@
 (Note that, to avoid cyclic dependencies, the increment and decrement constructions will only use the same-register-size adder.)
+
 First, the carry signal must be forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
@@
 This extra increment is cancelled with a controlled decrement.
-See the circuit diagram in figure \ref{fig:inline-adder-into-large} for the complete construction.
+See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
 
@@
 To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
-A top half that is incremented only if all of the bottom bits are on, and a bottom half that is unconditionally incremented.
-The bottom half can be incremented with the double-subtraction trick by borrowing the top half.
-But the double-subtraction trick doesn't work on the top half, because the bottom half can't be borrowed when it is being used as a control.
+A high half that is incremented only if all of the bottom bits are on, and a low half that is unconditionally incremented.
+The low half can be incremented with the double-subtraction trick by borrowing the high half.
+But the double-subtraction trick doesn't work on the high half, because the low half can't be borrowed when it is being used as a control.
 
@@
 Therefore the target was incremented by 2.
-To halve the +2 into a +1, prepend a dirty LSB onto the target register as shown in \ref{fig:compact-increment}.
+To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register as shown in \ref{fig:compact-increment}.
 
@@
 
-Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the construction described so far only works on odd-sized registers.
+Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
 For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
@@
 
-As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting all of the controls.
+As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting controls.
 Also, increment gates can be efficiently controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).
@@
 To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
-Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by all of the high bits of $A$ being off (and then, if any of those high bits are on, meaning $A$ must be larger than $B$, toggle the result bit as appropriate for the comparison).
+Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by all of the high bits of $A$ being off and otherwise (if any of the high bits are on) concluding $A>B$ and toggling the result bit as appropriate for the comparison being implemented.
 
@@
 The first $n$ qubits are needed to hold the input.
-The second $n$ qubits are for holding the output of exponentiating the input modulo a constant.
-One of the remaining qubits is used in the phase estimation process.
-The last qubit is a clean ancilla used by their modular offset operation to temporarily hold a comparison against the modulus.
+The second $n$ qubits are used for performing a non-power-of-2 QFT [[[ cite what they used ]]].
+One qubit is used for the phase estimation process [[[ cite what they used ]]].
+The last qubit is a clean ancilla used by their modular offset operation (to temporarily hold a comparison result).
 
@@
 Instead, we borrow the phase estimation qubit as a temporary dirty bit.
-Our modular offset circuit is a drop-in replacement for the one used by H\"{a}ner et all.
-It even has the same asymptotic costs: $O(N \lg N)$ size and $O(N)$ depth.
-Using our modular offset circuit, and borrowing the phase estimation qubit instead of using a separate clean bit, creates an implementation of Shor's factoring algorithm that uses only $2n+1$ qubits without increasing the size or depth complexity.
+Our modular offset circuit is a drop-in replacement for the one used by H\"{a}ner et all, with the same asymptotic costs ($O(N \lg N)$ size and $O(N)$ depth).
+Doing so creates an implementation of Shor's factoring algorithm that uses $2n+1$ qubits, without increasing the asymptotic size or depth with respect to the $2n+2$ qubit algorithm.
 
@@
 
-We present classical reversible circuits for incrementing, modular addition, and modular offset that use only a single dirty bit.
+We present linear sized classical reversible circuits for incrementing, modular addition, and modular offset that use only a single dirty bit.
 Our modular offset circuit improves the minimum number of qubits used when factoring an $n$-bit integer from $2n+2$ to $2n+1$.

### 2017-04-05 paper.tex
Half-rewritten into high-to-low order

@@
 \usepackage{hyperref}
-\title{Reversible Modular Addition using a Dirty Bit}
+\title{Factoring with $n+1$ clean qubits and $n$ dirty qubits}
 \author{Craig Gidney}
@@
 \em
-We present size-efficient reversible classical circuits for performing addition, incrementation, comparisons, modular addition, and modular offset on registers of differing sizes using either no ancilla or 1 ancilla in an unknown state.
-We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ \cite{takahashi2006, haner2016} to $2n+1$, without increasing the asymptotic cost of the circuit.
+We present size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
+We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+1$ clean qubits and $n$ dirty qubits, without increasing the asymptotic size or depth of the circuit.
 \end{abstract}
@@
 Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
-Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, because any temporarily unused qubit is a borrowable dirty bit.
+Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, since any temporarily unused bit is a borrowable dirty bit.
 
 Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
-For example, H\"{a}ner et al.
-used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, to $O(n^3)$ depth and $O(n^3 \lg n)$ size \cite{haner2016}.
-The modular addition construction in this paper improves on \cite{haner2016} by cutting the number of required qubits to factor an $n$-bit number from $2n+2$ to $2n+1$.
+For example, in \cite{haner2016} H\"{a}ner et al.
+used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, from ???????
+to $O(n^3)$ depth and $O(n^3 \lg n)$ size.
+The construction in this paper improve on \cite{haner2016} by cutting the number of required qubits to factor an $n$-bit number from $2n+2$ to $2n+1$.
 
@@
 The introduction describes the various sections and also itself for no discernible reason.
-Section \ref{sec:review} discusses existing circuit constructions.
-In sections \ref{sec:first-circuit} through \ref{sec:last-circuit} we describe various arithmetic circuit constructions that use 0 or 1 dirty bit, building up to modular addition.
-We then describe how to use our single-dirty-bit modular addition to factor an $n$ bit number using $2n+1$ qubits in section \ref{sec:applications}.
+Section \ref{sec:review} discusses existing circuit constructions for various arithmetic tasks.
+In section \ref{sec:construct} we iteratively reduce period finding into simpler and simpler arithmetic operations, until we reach constant-sized gates, while tracking the number of required dirty ancilla.
+Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented arithmetic circuit constructions.
+Finally, \ref{sec:conclusions} concludes and summarizes.
 
-All constructions assume a 2s-complement representation of integers.
-All circuit diagrams have the least significant bit towards the top.
+All circuit daigrams assume a 2s-complement representation of integers, and have the least significant bit towards the top.
 For clarity, circuit diagrams will divide compound operations into separate input and output parts, with inputs shaded gray.
@@
 
-\section{$O(N)$ inline adders} \label{sec:first-circuit}
+\subsection{Period Finding}
+
+...
+
+\section{From Period Finding to Toffolis, using $2n+1$ qubits} \label{sec:first-circuit} \label{sec:construct}
+
+\subsection{Period Finding}
+
+We begin with the abstract period finding circuit from ???, shown in figure \ref{fig:period-finding-solo-phase-qubit}.
+This circuit uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
+The only non-trivial operation is (a), controlled modular multiplication an $m$ qubit register, for which we will need $m$ clean ancilla.
+
+\begin{figure}
+  \centering
+  \makebox[\linewidth]{
+    \includegraphics[width=\linewidth]{assets/shor-period-finding.png}
+  }
+  \caption{\em
+	High-level period finding circuit [[[ cite ]]].
+    The triangle indicates how many ancilla are needed, by our constructions, to perform the linked operation.
+  }
+  \label{fig:period-finding}
+\end{figure}
+
+\begin{figure*}
+  \centering
+  \makebox[\linewidth]{
+    \includegraphics[width=\linewidth]{assets/shor-period-finding-solo-phase-qubit.png}
+  }
+  \caption{\em
+	Period finding with a single phase-estimation qubit [[[ cite ]]].
+	The small $\tiny \oplus$ are ``X-axis controls".
+	An X-axis control is equivalent to a normal control, but with a Hadamard gate applied before and after.
+	It conditions on the state $\frac{1}{\sqrt 2}|0\rangle - \frac{1}{\sqrt 2}|1\rangle$ instead of on the state $|1\rangle$.
+  }
+  \label{fig:period-finding-solo-phase-qubit}
+\end{figure*}
+
+\subsection{Controlled Modular Multiply}
+
+We do controlled modular multiplication with the same technique used in ???: multiply-accumulate into an extra register, then erase the input with an inverse multiply-accumulate.
+See figure \ref{fig:controlled-modular-multiply}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply.png}
+  \caption{\em
+    Reducing a controlled modular multiplication into two controlled modular multiply-accumulates.
+    Based on ???.
+  }
+  \label{fig:controlled-modular-multiply}
+\end{figure}
+
+The controlled modular multiply-accumulates require no ancilla.
+
+\subsection{Controlled Modular Multiply-Accumulate}
+
+To perform multiply-accumulation, we use a shift-and-add algorithm.
+We right-shift (i.e.
+divide by 2) the target register $m-1 {\pmod R}$ times, then begin iteratively left-shifting (i.e.
+multiplying by 2) and conditionally adding $K$ into the target based on the appropriate input bit.
+See figure \ref{fig:controlled-modular-multiply-accumulate}.
+
+Both the modular bit-shift and conditional offset operations require a single dirty bit, but have more than enough available to borrow.
+Also, both operations will be implemented in terms of what we call a pivot flip operation.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
+  \caption{\em
+    Reducing a controlled modular multiply-accumulate into multiplications by 2 and controlled modular additions.
+    Note that, although the controlled modular offset operations require a dirty ancilla, they can borrow it from one of the wires they don't affect.
+    Based on ???.
+  }
+  \label{fig:controlled-modular-multiply-accumulate}
+\end{figure}
+
+\subsection{Pivot-Flip}
+
+A "pivot-flip" is an operation that reverses the order of states less than a given pivot value, without affecting other states.
+For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
+Pivot-flips are useful for implementing cyclic operations, such as modular addition.
+
+The exact permutation performed by a pivot-flip with pivot equal to $k$ is:
+
+$$\text{PivotFlip}_k = \sum_{i=0}^{k-1} |k-i-1\rangle \langle i| + \sum_{i=k}^{N-1} |i\rangle \langle i|$$
+
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
+As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a)$ flips the range below $a$ but also flips the range above-and-including $a$.
+However, because $x \rightarrow \lnot(x - a)$ is its own inverse and doesn't change whether any value is less than the pivot, the operation can be controlled by a dirty ancilla being toggled by a comparison against $a$.
+
+See figure \ref{fig:const-pivot-flip} for the construction (as well as a less-than-or-equal-to-pivot variant).
+For states that are less than the pivot $a$, the dirty ancilla is being toggled so exactly one of the controlled flips will happen.
+For other states, the ancilla does not change and so the controlled flip happens either no times or two times (undoing itself) for a net effect of no effect.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/double-flip.png}
+  \caption{\em Subtraction followed by logical negation performs flips both above and below $k$.}
+  \label{fig:double-flip}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/pivot-flip.png}
+  \caption{\em Pivot flip circuits.
+  When the top wire bundle is a register, the subtractions are controlled by adding a control to any CNOTs, increments, and decrements in figures \ref{fig:inlineadder} or \ref{fig:inline-adder-into-large}.
+  When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
+  \label{fig:const-pivot-flip}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-offset.png}
+  \caption{\em Offset circuit from \cite{haner2016}, modified to include a control.}
+  \label{fig:controlled-offset}
+\end{figure}
+
+The cost of a $\text{PivotFlip}$ depends on if the pivot is a compile-time constant or not.
+When the pivot is input from another register, the $O(N)$ inline adder can be used.
+But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead.
+
+\subsection{Modular Doubling}
+
+If the modulus $R$ is even, modular doubling is irreversible and therefore not unitary (it can be performed, but it must create garbage bits leading to decoherence).
+So we assume that the modulus $R$ is odd.
+
+To multiply a register by 2, we use a pivot-flip operation, a subtraction, and CNOTs.
+See \ref{fig:modular-double}.
+This construction requires a dirty ancilla because the pivot-flip operation covers all wires but requires a dirty ancilla.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/modular-double.png}
+  \caption{\em
+    Modular doubling.
+    Assumes $R$ is odd.
+    Requires 1 dirty ancilla.
+    Trashes out-of-range values (ones greater or equal to $R$).
+  }
+  \label{fig:modular-double}
+\end{figure}
+
+https://arxiv.org/abs/1202.6614
+
+https://arxiv.org/abs/1110.2574
+
+[[[diagram of their construction]]]
+
+[[[diagram of their doubling]]]
+
+The construction requires $n+1$ clean ancilla.
+$n$ for the second register, and $1$ for holding a carry bit during the doubling step.
+By using the doubling construction in figure [[[whatever]]], the extra ancilla can be dirty instead of clean.
+
+[[[diagram of my doubling]]]
+
+xxx
+
+\subsection{Controlled Modular Addition}
+
+As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot flips.
+To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a-1$.
+See figure \ref{fig:modular-add} for the circuit.
+
+Be aware that this construction requires that the input register be in range.
+If the input is greater than or equal to $r$, the pivot-flips will trash the target register instead of adding the input mod $r$ into it.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
+  \caption{\em
+     A modular offset is three pivot flips.
+     One at $r-a$, then one at $r$, then one at $a$.
+   }
+  \label{fig:modular-add}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/modular-addition.png}
+  \caption{\em Reducing modular addition to non-modular offset, pivot-flips, and NOTs.}
+  \label{fig:modular-add}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
+  \caption{\em Reducing controlled modular addition to non-modular offset, controlled pivot-flips, and NOTs.}
+  \label{fig:modular-add}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/modular-offset.png}
+  \caption{\em Reducing modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
+  \label{fig:modular-add}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
+  \caption{\em Reducing controlled modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
+  \label{fig:modular-add}
+\end{figure}
+
+\section{$O(N)$ inline adders}
 
@@
   \centering
-  \includegraphics[totalheight=3cm]{inline-adder.png}
+  \includegraphics[width=\linewidth]{assets/inline-adder.png}
   \caption{\em Adder for input and target registers of the same size.
@@
   \centering
-  \includegraphics[totalheight=3cm]{inline-adder-into-large.png}
+  \includegraphics[width=\linewidth]{assets/inline-adder-into-large.png}
   \caption{\em
@@
 A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
-When $N$ dirty bits are available, $x$ can come from a register defined by those $N$ arbitrary bits, as shown in figure \ref{fig:double-sub-increment}.
+When $N$ dirty bits are available, $x$ can come from a register defined by those $N$ arbitrary bits, as shown in figure \ref{fig:increment-many-dirty}.
 
@@
   \centering
-  \includegraphics[totalheight=2cm]{double-sub-increment.png}
+  \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
   \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(N)$ depth, size, and $N$ dirty ancilla.}
-  \label{fig:double-sub-increment}
+  \label{fig:increment-many-dirty}
 \end{figure}
@@
 Therefore the target was incremented by 2.
-To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register as shown in \ref{fig:compact-increment}.
+To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register as shown in \ref{fig:increment-odd}.
 
@@
   \centering
-  \includegraphics[totalheight=1.5cm]{compact-increment.png}
+  \includegraphics[width=\linewidth]{assets/increment-odd.png}
   \caption{\em Odd-sized $O(N)$ increment with 1 dirty ancilla.}
-  \label{fig:compact-increment}
+  \label{fig:increment-odd}
 \end{figure}
@@
   \centering
-  \includegraphics[totalheight=2cm]{compact-increment-even.png}
+  \includegraphics[width=\linewidth]{assets/increment-even.png}
   \caption{\em Even-sized $O(N)$ increment with 1 dirty ancilla.}
-  \label{fig:compact-increment-even}
+  \label{fig:increment-even}
 \end{figure}
@@
   \centering
-  \includegraphics[totalheight=2cm]{controlled-increment-odd.png}
+  \includegraphics[width=\linewidth]{assets/controlled-increment-odd.png}
   \caption{\em Odd-sized $O(N)$ controlled increment with 1 dirty ancilla.}
@@
   \centering
-  \includegraphics[totalheight=2.5cm]{controlled-increment-even.png}
+  \includegraphics[width=\linewidth]{assets/controlled-increment-even.png}
   \caption{\em Even-sized $O(N)$ controlled increment with 1 dirty ancilla.}
@@
   \centering
-  \includegraphics[totalheight=1.4cm]{ancilla-bootstrap.png}
+  \includegraphics[width=\linewidth]{assets/ancilla-bootstrap.png}
   \caption{\em Bootstrapping a dirty ancilla out of an increment gate using quantum operations.
@@
   \centering
-  \includegraphics[totalheight=3cm]{comparison-less.png}
+  \includegraphics[width=\linewidth]{assets/comparison-less.png}
   \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
@@
   \centering
-  \includegraphics[totalheight=3cm]{comparison-leq.png}
+  \includegraphics[width=\linewidth]{assets/comparison-leq.png}
   \caption{\em Toggling a target bit if one register is less than or equal to another, without using ancilla.}
@@
 
-A pivot-flip operation reverses the order of states less than or equal to a given pivot value.
-For example, a pivot-flip with the pivot equal to 3 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
-Pivot-flips are useful for implementing cyclic operations, such as modular addition.
-
-The exact permutation performed by a pivot-flip with pivot equal to $k$ is:
-
-$$\text{PivotFlip}_k = \sum_{i=0}^k |k-i\rangle \langle i| + \sum_{i=k+1}^{N-1} |i\rangle \langle i|$$
-
-And the overall permutation, including the pivot input, is:
-
-$$\text{PivotFlip} = \sum_{k=0}^{N-1} |k\rangle \langle k| \otimes \text{PivotFlip}_{k}$$
-
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a - 1)$ nearly does what is required.
-As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a - 1)$ flips the range below-and-including the pivot but also flips the range above the pivot.
-However, because $x \rightarrow \lnot(x - a - 1)$ is its own inverse and doesn't change whether any value is less than or equal to the pivot, the operation can be controlled by dirty comparisons against the pivot.
-
-By applying $x \rightarrow \lnot(x - a - 1)$ twice, controlled by a dirty bit that is toggled by the comparison, the desired effect is achieved.
-For states that are less than or equal to the pivot $a$, the dirty ancilla is being toggled and exactly one of the controlled flips will happen.
-For other states, the ancilla does not change and so the controlled flip happens no times or two times, undoing itself.
-See figure \ref{fig:const-pivot-flip} fot the construction, as well as a variant that only flips values strictly less than the pivot instead of including the pivot.
-
-\begin{figure}
-  \centering
-  \includegraphics[totalheight=4cm]{double-flip.png}
-  \caption{\em Subtraction followed by logical negation performs flips both above and below $k$.}
-  \label{fig:double-flip}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[totalheight=4.5cm]{pivot-flip.png}
-  \caption{\em Pivot flip circuits.
-  When the top wire bundle is a register, the subtractions are controlled by adding a control to any CNOTs, increments, and decrements in figures \ref{fig:inlineadder} or \ref{fig:inline-adder-into-large}.
-  When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
-  \label{fig:const-pivot-flip}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[totalheight=3.1cm]{controlled-offset.png}
-  \caption{\em Offset circuit from \cite{haner2016}, modified to include a control.}
-  \label{fig:controlled-offset}
-\end{figure}
-
-The cost of a $\text{PivotFlip}$ depends on if the pivot is a compile-time constant or not.
-When the pivot is input from another register, the $O(N)$ inline adder can be used.
-But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead.
-
 \section{$O(N \lg N)$ Modular Addition and Modular Offset} \label{sec:last-circuit}
-
-A modular addition is three pivot flips.
-To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a-1$.
-See figure \ref{fig:modular-add}.
-
-Be aware that this construction requires that the input register be in range.
-If the input is greater than or equal to $r$, the pivot-flips will trash the target register instead of adding the input mod $r$ into it.
-
-\begin{figure}
-  \centering
-  \includegraphics[totalheight=1.4cm]{modular-addition.png}
-  \caption{\em Modular addition circuit.
-  The $+r$ operations are offset gates requiring $O(N \lg N)$ size when the modulus is a constant.
-  The pivot-flips require a dirty bit.}
-  \label{fig:modular-add}
-\end{figure}
 
@@
   \centering
-  \includegraphics[totalheight=16cm]{modular-add-graph.png}
+  \includegraphics[width=\linewidth]{assets/modular-add-graph.png}
   \caption{\em
@@
 \end{figure}
+
+To perform modular multiplication, we use the construction from [[[[ reference ]]], which works by iteratively scaling and adding the target register into an ancilla register.
+The ancilla and target register then switch roles, and the old target register is erased by adding in the new target times a  multiplicative inverse.
 
@@
 Doing so creates an implementation of Shor's factoring algorithm that uses $2n+1$ qubits, without increasing the asymptotic size or depth with respect to the $2n+2$ qubit algorithm.
+
+[[[ diagram of Shor's algorithm ]]]
 

### 2017-04-06 paper.tex
Second draft

@@
 \usepackage{hyperref}
-\title{Factoring with $n+1$ clean qubits and $n$ dirty qubits}
+\title{Factoring with $n+2$ clean qubits and $n-1$ dirty qubits}
 \author{Craig Gidney}
@@
 We present size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
-We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+1$ clean qubits and $n$ dirty qubits, without increasing the asymptotic size or depth of the circuit.
+We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
 \end{abstract}
@@
 
-When constructing quantum circuits, or classical reversible circuits, the number of available ancilla is an important resource.
+When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
 An ancilla is an extra bit or qubit available for use by a circuit as temporary workspace.
@@
 Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
-For example, in \cite{haner2016} H\"{a}ner et al.
-used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, from ???????
-to $O(n^3)$ depth and $O(n^3 \lg n)$ size.
-The construction in this paper improve on \cite{haner2016} by cutting the number of required qubits to factor an $n$-bit number from $2n+2$ to $2n+1$.
+Also, when attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
+In this paper, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.
 
@@
 The introduction describes the various sections and also itself for no discernible reason.
-Section \ref{sec:review} discusses existing circuit constructions for various arithmetic tasks.
 In section \ref{sec:construct} we iteratively reduce period finding into simpler and simpler arithmetic operations, until we reach constant-sized gates, while tracking the number of required dirty ancilla.
 Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented arithmetic circuit constructions.
-Finally, \ref{sec:conclusions} concludes and summarizes.
+Finally, section \ref{sec:conclusion} concludes and summarizes.
 
-All circuit daigrams assume a 2s-complement representation of integers, and have the least significant bit towards the top.
-For clarity, circuit diagrams will divide compound operations into separate input and output parts, with inputs shaded gray.
+All constructions assume a 2s-complement representation of integers.
+Multi-qubit arithmetic operations always have the least significant bit towards the top.
+For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray.
 
-\section{Existing Constructions} \label{sec:review}
-
-\subsection{Addition}
-
-In \cite{draper2000}, Draper describes an addition circuit that uses the quantum Fourier transform to convert controlled phasing into addition.
-Draper notes that the construction uses $O(N)$ depth and $O(N^2)$ controlled rotations, but approximation techniques \cite{barenco1996} \cite{cleve2000} can improve the depth and size to $O(\lg N)$ and $O(N \lg N)$ respectively.
-
-Both \cite{cuccaro2004} and \cite{van2004} describe classical reversible ripple-carry adders that use a single clean ancilla, $O(N)$ depth, and $O(N)$ size.
-Takahashi \cite{takahashi2005} showed how to use the most-significant-bit (MSB) of the input register, instead of a separate ancilla, to propagate the carry signal.
-Takashi later showed how to trade depth against size \cite{takahashi2009}.
-
-We generalize the ideas in \cite{takahashi2005} to apply to additions where the input and target registers have differing sizes.
-
-\subsection{Offset}
-
-An offset gate adds a compile-time constant into a register.
-Some reversible adder constructions can be used as offset gates, but others can't because they use qubits in the input register as temporary workspace.
-An offset gate has no input register.
-
-Draper's QFT-based addition construction \cite{draper2000} doesn't use the input register as workspace, and so trivially generalizes into an offset gate.
-
-\cite{haner2016} describes a Toffoli-based divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $n/2$ dirty ancilla.
-[[[ Actually, only 1 is required by their construction.
-0 if you go quantum.
-And when I emailed them about it they said they would fix the oversight.
-So I'm not sure how to describe that.
-It is important for this paper that we have an non-modular offset circuit that only needs 1 dirty ancilla.
-]]]
-
-We describe a circuit for the special case of offset-by-1, i.e. incrementation.
-Our increment circuit uses $O(N)$ depth, $O(N)$ size, one dirty ancilla, and no non-classical operations.
-We also note how to cut the ancilla with quantum operations.
-This construction is a simplified version of our unpublished incrementer described in \cite{gidney2015}.
-
-\subsection{Comparison}
-
-Comparison operators, which toggle a target bit based on if one input is larger, smaller, and/or equal to another, can be implemented efficiently using adders \cite{takahashi2005}.
-
-Due to generalizing Takahashi's adder \cite{takahashi2005} to apply to different sizes, we also generalize Takahashi's comparators to apply to registers of different sizes.
-
-\subsection{Modular Addition}
-
-\cite{vedral1995} describes a modular adder that uses $O(N)$ clean ancilla.
-It temporarily stores a comparison in a clean ancilla, performs a normal addition, uses a subtraction conditioned on the ancilla to wrap the result back into range, then performs a cleanup comparison to clear the ancilla.
-
-All but one of the $O(N)$ clean ancilla used by \cite{vedral1995} are due to using inefficient addition circuits.
-Using later zero-ancilla adder constructions reduces xthe number of clean ancilla required to perform a modular addition to just one.
-
-Our modular adder construction also requires a single ancilla, but the ancilla is no longer required to be clean.
+\section{Period Finding into Toffolis} \label{sec:construct}
 
@@
 
-...
+A high-level circuit for period finding, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
 
-\section{From Period Finding to Toffolis, using $2n+1$ qubits} \label{sec:first-circuit} \label{sec:construct}
+Because period-finding measures all qubits immediately after performing a quantum Fourier transform, most of the transformed qubits can be measured earlier than shown.
+In fact, each qubit can be measured so early that the next does not even need to be initialized yet!
+Only one of the qubits in the phase-estimation register needs to be present at a time, and so the phase register can be reduced to a single repeatedly-used qubit \cite{beauregard2003}.
+Figure \ref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
 
-\subsection{Period Finding}
+The period-finding circuit in figure \ref{fig:period-finding-solo-phase-qubit} is our high-level starting point.
+It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
+The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla).
 
-We begin with the abstract period finding circuit from ???, shown in figure \ref{fig:period-finding-solo-phase-qubit}.
-This circuit uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
-The only non-trivial operation is (a), controlled modular multiplication an $m$ qubit register, for which we will need $m$ clean ancilla.
+We can perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
+However, if we want to allow the register to be dirty, it is convenient to pair each modular multiplication with an inverse multiplication on another register (a ``multiply-both-ways").
+This change on its own would trash the second register, since at the end of the circuit the second register will have been multiplied by some unknown constant.
+But, since the first register is initialized to store 1 and gets multiplied by an inverse constant, we can untrash the second register by multiplying by whatever is in the first register at the end of the circuit.
+Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
+
+Be aware that none of our modular circuit constructions guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
+To ensure the second register's value is in range, we use a clean zero'd ancilla for its most-significant-bit (MSB).
 
@@
   \caption{\em
-	High-level period finding circuit [[[ cite ]]].
-    The triangle indicates how many ancilla are needed, by our constructions, to perform the linked operation.
+	High-level period finding circuit \cite{Shor1999}.
+	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(N)$ controls the precision of the phase estimation step.
+    The triangles indicate how many ancilla are needed``behind the scenes", by our constructions, to perform an operation.
+	Recovering the period requires classically processing the sampled output with a continued fractions algorithm.
   }
@@
   \caption{\em
-	Period finding with a single phase-estimation qubit [[[ cite ]]].
-	The small $\tiny \oplus$ are ``X-axis controls".
+	Period finding with a single phase-estimation qubit \cite{beauregard2003}.
+	The small oplus' ({\tiny $\oplus$}) are ``X-axis controls".
 	An X-axis control is equivalent to a normal control, but with a Hadamard gate applied before and after.
@@
 
-\subsection{Controlled Modular Multiply}
+\begin{figure*}
+  \centering
+  \makebox[\linewidth]{
+    \includegraphics[width=\linewidth]{assets/shor-period-finding-solo-phase-qubit-double-register.png}
+  }
+  \caption{\em
+	Period finding with a single phase-estimation qubit and paired inverse multiplications.
+	The terminating bra $\langle 0 |$ indicates that the bottom wire will always end up in the state $|0\rangle$, without post-selection.
+  }
+  \label{fig:period-finding-solo-phase-qubit-explicit-dirty-register}
+\end{figure*}
 
-We do controlled modular multiplication with the same technique used in ???: multiply-accumulate into an extra register, then erase the input with an inverse multiply-accumulate.
-See figure \ref{fig:controlled-modular-multiply}.
+\subsection{Controlled Modular Multiply-Both-Ways}
+
+As in \cite{beauregard2003}, we perform controlled modular multiplication with modular multiply-accumulates into a second register.
+However, by using a third multiply-accumulate, we end up with a compound multiply-and-inverse-multiply operation instead of requiring a clean ancilla register.
+
+As with all other modular arithmetic constructions in this paper, we require that the input registers be in range (i.e. have values below $R$).
+
+Suppose that the two registers start in the state $(a, b)$, working modulo $R$.
+By multiply-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
+Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(bK^{-1}, b+aK)$.
+Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(bK^{-1}, aK)$.
+Finally, we swap to $(aK, bK^{-1})$.
+The overall effect is to multiply the first register by $K$ and the second register by $K^{-1}$, as desired.
+See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
+
+In the case where $K$ has no multiplicative inverse modulo $R$, this construction will not work (it would define an invalid irreversible operation).
+However, that would mean $K$ is a factor of $R$, a case which the very lucky user can check and handle classically before invoking a quantum period finding circuit.
 
@@
   \caption{\em
-    Reducing a controlled modular multiplication into two controlled modular multiply-accumulates.
-    Based on ???.
+    Performing a controlled modular multiply-both-ways with three controlled modular multiply-accumulates.
+    $K$ must have a multiplicative inverse modulo $R$.
   }
@@
 
-The controlled modular multiply-accumulates require no ancilla.
-
 \subsection{Controlled Modular Multiply-Accumulate}
 
-To perform multiply-accumulation, we use a shift-and-add algorithm.
+To perform multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
 We right-shift (i.e.
-divide by 2) the target register $m-1 {\pmod R}$ times, then begin iteratively left-shifting (i.e.
-multiplying by 2) and conditionally adding $K$ into the target based on the appropriate input bit.
+divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e.
+multiplying it by 2) and adding $K$ into the target conditioned on the most significant bit of the input, then the next most significant bit, and so forth.
+(The more significant bits go first because their effects must be hit by more left-shifts.)
 See figure \ref{fig:controlled-modular-multiply-accumulate}.
 
-Both the modular bit-shift and conditional offset operations require a single dirty bit, but have more than enough available to borrow.
-Also, both operations will be implemented in terms of what we call a pivot flip operation.
+A single dirty bit is required by the conditional offset operations, and two dirty bits are required by the modular doublings, but there are more than enough unused bits available to borrow, so the overall circuit doesn't require any dirty bits.
+
+Note that the modular doublings require $R$ to be odd, since otherwise the operation would be irreversible and non-unitary.
+However, given that in the intended use case $R$ is a number to be factored, it is reasonable to require callers to have factored out multiples of two beforehand.
 
@@
   \caption{\em
-    Reducing a controlled modular multiply-accumulate into multiplications by 2 and controlled modular additions.
-    Note that, although the controlled modular offset operations require a dirty ancilla, they can borrow it from one of the wires they don't affect.
-    Based on ???.
+    Reducing a controlled modular multiply-accumulate into the modular equivalent of shift-and-add.
+    Assumes $R$ is odd.
   }
@@
 
-\subsection{Pivot-Flip}
+\subsection{Controlled Pivot-Flips}
 
-A "pivot-flip" is an operation that reverses the order of states less than a given pivot value, without affecting other states.
+We will implement both modular doubling and modular offset/addition in terms of ``pivot-flips".
+A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
 For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
-Pivot-flips are useful for implementing cyclic operations, such as modular addition.
 
-The exact permutation performed by a pivot-flip with pivot equal to $k$ is:
+The exact permutation performed by a pivot-flip with pivot equal to $A$ is:
 
-$$\text{PivotFlip}_k = \sum_{i=0}^{k-1} |k-i-1\rangle \langle i| + \sum_{i=k}^{N-1} |i\rangle \langle i|$$
+$$\text{PivotFlip}_A = \sum_{i=0}^{A-1} |A-i-1\rangle \langle i| + \sum_{i=A}^{N-1} |i\rangle \langle i|$$
 
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
-As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a)$ flips the range below $a$ but also flips the range above-and-including $a$.
-However, because $x \rightarrow \lnot(x - a)$ is its own inverse and doesn't change whether any value is less than the pivot, the operation can be controlled by a dirty ancilla being toggled by a comparison against $a$.
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
+However, because $x \rightarrow \lnot(x - A)$ is its own inverse and doesn't change whether a given value is below the pivot, the operation can be toggle-controlled by the comparison against $A$.
+See figure \ref{fig:controlled-pivot-flip} for the construction.
+For states that are less than the pivot $A$, the comparison is true, the ancilla is toggled both times, and exactly one of the controlled subtract-and-inverts will fire.
+For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.
 
-See figure \ref{fig:const-pivot-flip} for the construction (as well as a less-than-or-equal-to-pivot variant).
-For states that are less than the pivot $a$, the dirty ancilla is being toggled so exactly one of the controlled flips will happen.
-For other states, the ancilla does not change and so the controlled flip happens either no times or two times (undoing itself) for a net effect of no effect.
+Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
+This distinction matters because the constant-offset case uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
+Also, the constant-offset case requires an extra dirty bit unless quantum operations are used (e.g. by using the technique demonstrated in figure \ref{fig:bootstrap-ancilla}).
 
@@
   \centering
-  \includegraphics[width=\linewidth]{assets/double-flip.png}
-  \caption{\em Subtraction followed by logical negation performs flips both above and below $k$.}
-  \label{fig:double-flip}
+  \includegraphics[width=\linewidth]{assets/controlled-pivot-flip.png}
+  \caption{\em
+    Controlled pivot-flip circuit with an enregistered pivot.
+  }
+  \label{fig:controlled-pivot-flip}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\linewidth]{assets/pivot-flip.png}
-  \caption{\em Pivot flip circuits.
-  When the top wire bundle is a register, the subtractions are controlled by adding a control to any CNOTs, increments, and decrements in figures \ref{fig:inlineadder} or \ref{fig:inline-adder-into-large}.
-  When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
-  \label{fig:const-pivot-flip}
+  \includegraphics[width=\linewidth]{assets/controlled-const-pivot-flip.png}
+  \caption{\em
+    Controlled pivot-flip circuit with a constant pivot.
+    The extra dirty ancilla, compared to the enregistered case, comes from the controlled-offset construction.
+  }
+  \label{fig:controlled-pivot-flip}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{assets/controlled-offset.png}
-  \caption{\em Offset circuit from \cite{haner2016}, modified to include a control.}
+  \caption{\em
+      Offset circuit from \cite{haner2016}, but modified to include controls.
+      Uses $O(n \lg n)$ gates and $O(n)$ depth (by overlapping the recursive cases).
+      Note that, if $c$ is not $O(1)$, the recursion introduces additional non-linear costs.
+  }
   \label{fig:controlled-offset}
@@
 
-The cost of a $\text{PivotFlip}$ depends on if the pivot is a compile-time constant or not.
-When the pivot is input from another register, the $O(N)$ inline adder can be used.
-But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead.
-
 \subsection{Modular Doubling}
 
-If the modulus $R$ is even, modular doubling is irreversible and therefore not unitary (it can be performed, but it must create garbage bits leading to decoherence).
-So we assume that the modulus $R$ is odd.
+To multiply a register by 2 modulo an odd $R$, we use pivot-flips and subtraction to align the two halves of the input space on the MSB boundaries.
+Then we left-rotate.
+Note that this trashes out-of-range values larger than $R$ instead of leaving them alone.
+See figure \ref{fig:modular-double}.
 
-To multiply a register by 2, we use a pivot-flip operation, a subtraction, and CNOTs.
-See \ref{fig:modular-double}.
-This construction requires a dirty ancilla because the pivot-flip operation covers all wires but requires a dirty ancilla.
+To invert the doubling, i.e. to divide by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
+Recurse-and-reverse works on all circuits presented in this paper (e.g.
+to perform subtraction instead of addition, or to decrement instead of increment), except period-finding, because the circuits don't generate garbage or use measurement.
 
@@
     Assumes $R$ is odd.
-    Requires 1 dirty ancilla.
-    Trashes out-of-range values (ones greater or equal to $R$).
+    Requires 2 dirty ancilla, or 1 if using quantum operations.
   }
@@
 
-https://arxiv.org/abs/1202.6614
-
-https://arxiv.org/abs/1110.2574
-
-[[[diagram of their construction]]]
-
-[[[diagram of their doubling]]]
-
-The construction requires $n+1$ clean ancilla.
-$n$ for the second register, and $1$ for holding a carry bit during the doubling step.
-By using the doubling construction in figure [[[whatever]]], the extra ancilla can be dirty instead of clean.
-
-[[[diagram of my doubling]]]
-
-xxx
-
-\subsection{Controlled Modular Addition}
+\subsection{Controlled Modular Addition and Offset}
 
 As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot flips.
-To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a-1$.
+To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
 See figure \ref{fig:modular-add} for the circuit.
 
-Be aware that this construction requires that the input register be in range.
-If the input is greater than or equal to $r$, the pivot-flips will trash the target register instead of adding the input mod $r$ into it.
+Although our modular addition circuit requires a dirty bit, {\em controlled} modular addition does {\em not}.
+See figure \ref{fig:controlled-modular-add}.
+However, when performing modular offset (i.e. adding a compile-time constant to a register), we instead require two dirty bits.
+See figures \ref{fig:modular-offset} and \ref{fig:controlled-modular-offset}.
+
+Be aware that pivot-flip-based-addition assumes that {\em both} the input and target register are in range.
+If the input is greater than or equal to $R$, the pivot-flips will trash the target register instead of adding the input mod $R$ into it.
 
@@
   \caption{\em
-     A modular offset is three pivot flips.
-     One at $r-a$, then one at $r$, then one at $a$.
+     Modular addition of $A \pmod{R}$ with three pivot flips.
+     One at $R-A$, then one at $R$, then one at $A$.
+     Requires $A \leq R$ (or else the operation will trash the target).
    }
-  \label{fig:modular-add}
+  \label{fig:mod-add-from-pivot-flip-bars}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{assets/modular-addition.png}
-  \caption{\em Reducing modular addition to non-modular offset, pivot-flips, and NOTs.}
+  \caption{\em Non-constant modular addition from pivot-flips.}
   \label{fig:modular-add}
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
-  \caption{\em Reducing controlled modular addition to non-modular offset, controlled pivot-flips, and NOTs.}
-  \label{fig:modular-add}
+  \caption{\em
+    Controlled modular addition.
+    The pivot-flips don't need to be controlled so, unlike the uncontrolled modular addition construction, this construction requires no dirty bits.
+  }
+  \label{fig:controlled-modular-add}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{assets/modular-offset.png}
-  \caption{\em Reducing modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
-  \label{fig:modular-add}
+  \caption{\em Modular offset.}
+  \label{fig:modular-offset}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
-  \caption{\em Reducing controlled modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
-  \label{fig:modular-add}
+  \caption{\em
+    Controlled modular offset.
+  }
+  \label{fig:controlled-modular-offset}
 \end{figure}
 
-\section{$O(N)$ inline adders}
+\subsection{Comparison}
+
+Comparison operations toggle a target bit based on the relationship between two input registers.
+We implement comparisons as in \cite{takahashi2005}, using an addition followed by a subtraction that is slightly smaller to clear all changes except an overflow signal into the target.
+See figures \ref{fig:comparison-less} and \ref{fig:comparison-less-const}.
+
+To perform a $\leq$ instead of a $<$, simply add/subtract $A+1$ instead of $A$.
+(Apply the extra increment/decrement to the target register, not the input register, since incrementing the input register may wrap it to zero.)
+
+Beware that,  in these constructions, if the $B$ register is smaller than the $A$ register, the high bits of $A$ are lost and the construction breaks.
+To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
+Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by the high bits of $A$ all being off and otherwise (if any of the high bits are on) concluding $A>B$ and toggling the result bit as appropriate for the comparison being implemented.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/comparison-less.png}
+  \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
+  \label{fig:comparison-less}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/comparison-less-const.png}
+  \caption{\em Toggling a target bit if a register is less then a constant, using a dirty ancilla and $O(n \lg n)$ gates.}
+  \label{fig:comparison-less-const}
+\end{figure}
+
+\subsection{Addition and Offset}
 
@@
 If the ancilla is not initialized in the off state, the target register ends up storing $a+b+1$ instead of $a+b$.
-For clarity, we review how to elide the ancilla before generalizing the construction to apply to registers of differing size.
+For clarity, we review how to elide the ancilla.
 
@@
 
-Once the ancilla can be dirty the MSB of the input register can play the role of the ancilla, as shown in figure \ref{fig:inlineadder}.
+Once the ancilla can be dirty, the MSB of the input register can play the role of the ancilla, as in \cite{takahashi2005} and shown in figure \ref{fig:inlineadder}.
 
@@
 
-To perform subtraction instead of addition: run the circuit in reverse, temporarily invert the target register by bordering the circuit with NOT gates on the target, or invert the CSWAP controls.
+Because this adder construction uses the input register as workspace, it breaks when the input and target registers differ in size.
+When the input register is larger, the high bits of the input register can simply be ignored.
+But, when the target register is larger, there is insufficient workspace and also the logical-negation trick that allowed the ancilla to be dirty breaks.
 
-Because this adder construction uses the input register as workspace, it only works when the input register is the same size as the target register.
-When the input register is larger, the high bits of the input register can simply be ignored.
-But, when the target register is larger, the workaround is more difficult.
-Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the ancilla to be dirty.
-
-To work around the target register being larger, three fixes are needed.
-Each uses a controlled increment or decrement gate, which will be described in the next section.
+To work around the target register being larger, three tweaks are needed.
+Each uses a controlled increment or decrement gate, which will be described in the next subsection.
 (Note that, to avoid cyclic dependencies, the increment and decrement constructions will only use the same-register-size adder.)
@@
 
-\section{$O(N)$ incrementer}
+Another problem with using the input register as workspace appears when adding a constant offset into a target: there {\em isn't} an input register.
+We use the construction from \cite{haner2016}, which works for constant offsets but increases the gate cost from $\Theta(n)$ to $\Theta(n \lg n)$ and also requires a dirty ancilla instead of no ancilla (unless using quantum operations).
+See figure \ref{fig:controlled-offset}.
+We note that our paper would have significantly fewer diagrams if offset and addition didn't have differing costs.
+
+\subsection{Increment}
 
 A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
-When $N$ dirty bits are available, $x$ can come from a register defined by those $N$ arbitrary bits, as shown in figure \ref{fig:increment-many-dirty}.
+When $n$ dirty bits are available, $x$ can come from a register defined by those $n$ arbitrary bits, as shown in figure \ref{fig:increment-many-dirty}.
 
@@
 
-To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
+To improve from $n$ dirty bits to a single dirty bit, we break the increment operation into two halves.
 A high half that is incremented only if all of the bottom bits are on, and a low half that is unconditionally incremented.
@@
 
-Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
+Because the adder/subtracter construction used by the incrementer in figure \ref{fig:increment-odd} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
 For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
-That construction is shown in figure \ref{fig:compact-increment-even}.
+That construction is shown in figure \ref{fig:increment-even}.
 
@@
 
-As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting controls.
-Also, increment gates can be efficiently controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).
+Increment gates can be efficiently controlled by using another invert-input-and-output trick.
+See figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}.
 
@@
 
-Note that it is impossible to create a classical incrementer that uses no ancilla.
-The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
-Quantumly, the parity barrier can be bypassed by using partial rotations (see figure \ref{fig:bootstrap-ancilla}).
-
 \begin{figure}
@@
 
-\section{$O(N)$ inline comparisons}
+\section{Overview and Improvements} \label{sec:costs}
 
-We review how to implement comparison operations, which toggle a target bit based on the relationship between two input registers, using addition and subtraction operations.
+In figure \ref{fig:dependencies} we show a dependency graph of the constructions discussed in this paper.
+The total cost of period-finding, using this construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
+We perform $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
 
-The is-B-less-than-A comparison is implemented by subtracting $A$ out of $B$, with a target bit appended as $B$'s MSB, then adding $A$ into $B$ without the target bit appended.
-See figure \ref{fig:comparison-less}.
+Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of multiply-both-ways for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
+
+Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
+Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and in the context of period-finding there are many qubits available to borrow whenever a pivot-flip is needed.
+This improvement saves a qubit, reducing the total number of qubits required for period finding to $2n+1$.
+
+Previous modular multiplication constructions used a clean ancilla register \cite{haner2016}.
+For these constructions, any garbage in the ancilla register would infect the target register and trash it.
+We use an extra multiply-accumulate to cancel out garbage entering the target register, but the ancilla register is still multiplified by the inverse of whatever the target register is multiplied by, creating a "multiply-both-ways" circuit.
+We don't want to multiply the ancilla register, since it must be returned to its starting state before our circuit ends, but we can fix the problem with a final cleanup multiplication informed by a measurement of the main register.
+This improvement allows $n-1$ of the qubits in the ancilla register, all except the MSB, to be dirty.
+The MSB must be zero to ensure values are not larger than the modulus.
+
+Previous published incrementers (not counting a rough unpublished version of our construction \cite{gidney2015} being cited by \cite{haner2016}) required either $O(n^2)$ gates or $O(n)$ ancilla \cite{draper2000, barenco1995}.
+Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
+Note that, for classical reversible computation, 1 dirty ancilla is optimal.
+The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
+With quantum operations, we achieve zero ancilla and $O(n)$ gates (see figure \ref{fig:bootstrap-ancilla}).
 
@@
   \centering
-  \includegraphics[width=\linewidth]{assets/comparison-less.png}
-  \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
-  \label{fig:comparison-less}
+  \includegraphics[height=17.3cm]{assets/dependencies.png}
+  \caption{\em
+    Graph of non-trivial dependencies between constructions, from period finding to fixed-size gates.
+    Edges are labelled by an upper bound on how many times the target is used by the source.
+  }
+  \label{fig:dependencies}
 \end{figure}
-
-The same idea is used for is-B-less-than-or-equal-to-A, except $A+1$ is subtracted and added instead of $A$.
-The extra $\pm 1$ can be performed by increment and decrement gates, as shown in figure \ref{fig:comparison-leq}.
-When the two registers are the same size, the increment gate can be omitted by inverting the $A$ and $B$ registers during the addition.
-
-$>$ and $\geq$ comparisons are done by performing the corresponding opposite comparison ($\leq$ and $<$ respectively) and inverting the result by toggling the target.
-
-Beware that, when the $B$ register is smaller than the $A$ register, the high bits of $A$ are not accounted for by the subtraction and addition and so the construction breaks.
-To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
-Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by all of the high bits of $A$ being off and otherwise (if any of the high bits are on) concluding $A>B$ and toggling the result bit as appropriate for the comparison being implemented.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/comparison-leq.png}
-  \caption{\em Toggling a target bit if one register is less than or equal to another, without using ancilla.}
-  \label{fig:comparison-leq}
-\end{figure}
-
-\section{$\text{PivotFlip}$}
-
-\section{$O(N \lg N)$ Modular Addition and Modular Offset} \label{sec:last-circuit}
-
-Although the fact is obfuscated by the number of constructions used, our modular addition construction is performed entirely in terms of normal addition (with various tweaks and tricks and inversions along the way).
-See figure \ref{fig:modular-dependencies} for the full dependency graph.
-
-The modular addition uses $O(N \lg N)$ gates because that is the cost of the offset gate in \cite{haner2016}, and we subtract the modulus (typically a compile-time constant) out of a register during one of the pivot-flips.
-If both the amount to add {\em and the modulus} are in registers, then the construction only requires $O(N)$ gates.
-The total depth is $O(N)$ regardless.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/modular-add-graph.png}
-  \caption{\em
-      Dependency graph of constructions used to create a classical reversible modular adder / modular offset gates using only 1 dirty bit.}
-  \label{fig:modular-dependencies}
-\end{figure}
-
-To perform modular multiplication, we use the construction from [[[[ reference ]]], which works by iteratively scaling and adding the target register into an ancilla register.
-The ancilla and target register then switch roles, and the old target register is erased by adding in the new target times a  multiplicative inverse.
-
-\section{Factoring with $2n+1$ qubits} \label{sec:applications}
-
-In \cite{haner2016}, H\"{a}ner et al. describe how to perform Shor's algorithm with Toffoli-based circuits using $2n+2$ qubits.
-The first $n$ qubits are needed to hold the input.
-The second $n$ qubits are used for performing a non-power-of-2 QFT [[[ cite what they used ]]].
-One qubit is used for the phase estimation process [[[ cite what they used ]]].
-The last qubit is a clean ancilla used by their modular offset operation (to temporarily hold a comparison result).
-
-The modular addition circuit we have presented does not require that last clean ancilla.
-Instead, we borrow the phase estimation qubit as a temporary dirty bit.
-Our modular offset circuit is a drop-in replacement for the one used by H\"{a}ner et all, with the same asymptotic costs ($O(N \lg N)$ size and $O(N)$ depth).
-Doing so creates an implementation of Shor's factoring algorithm that uses $2n+1$ qubits, without increasing the asymptotic size or depth with respect to the $2n+2$ qubit algorithm.
-
-[[[ diagram of Shor's algorithm ]]]
 
@@
 
-We present linear sized classical reversible circuits for incrementing, modular addition, and modular offset that use only a single dirty bit.
-Our modular offset circuit improves the minimum number of qubits used when factoring an $n$-bit integer from $2n+2$ to $2n+1$.
-
-For incrementing, a single dirty ancilla is classically optimal because the permutation defined by an increment operation is odd.
-This obstacle does not apply to modular additions, because any odd parities can made even with a useless swap in the range of states above the modulus.
-So it may be possible to elide the final dirty ancilla our modular addition construction requires, even when restricted to Toffoli-based circuits.
+We presented Toffoli-based arithmetic circuits that reduce the number of qubits required to perform Shor's algorithm by 1, and the number of {\em clean} qubits required by an additional $n-1$, compared to the best previous constructions.
 

### 2017-04-22 paper.tex
Fixing issues found by implementing in ProjectQ - Adding controlled variants of more operations - Wire bundles have thick wires in diagrams - Not showing ancilla counters as 'satisfied' on input side of diagrams - Mentioning bit permutation / multi-not operations. - Merging increment diagrams.

@@
 \em
-We present size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
-We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
+We present asymptotically size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
+We improve the number of qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
 \end{abstract}
@@
 Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
-Also, when attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
-In this paper, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.
+When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
+In this paper, at great cost to the constant factors in the gate count and depth, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits without increasing the asymptotic costs.
 
@@
 Multi-qubit arithmetic operations always have the least significant bit towards the top.
-For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray.
+For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray, when applicable.
+For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we provide controllable constructions of that operation that scale well with the number of controls.
 
@@
 It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
-The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla).
+The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla using our constructions).
 
-We can perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
-However, if we want to allow the register to be dirty, it is convenient to pair each modular multiplication with an inverse multiplication on another register (a ``multiply-both-ways").
-This change on its own would trash the second register, since at the end of the circuit the second register will have been multiplied by some unknown constant.
-But, since the first register is initialized to store 1 and gets multiplied by an inverse constant, we can untrash the second register by multiplying by whatever is in the first register at the end of the circuit.
+We perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
+However, to allow the second register to be dirty, it is convenient to pair each modular multiplication with an inverse and negated multiplication on another register.
+We will refer to this combined operation as a ``bimultiply".
+
+Note that we must preserve the state of the second register, but bimultiplication trashes that state with many multiplications by unknown constants.
+However, because the first register is initialized to store 1 and gets multiplied by constants inverse to the ones trashing the second register, we can fix the second register by measuring the first register and using its value to perform a clean-up bimultiplication after the usual end of the circuit.
 Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
 
-Be aware that none of our modular circuit constructions guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
-To ensure the second register's value is in range, we use a clean zero'd ancilla for its most-significant-bit (MSB).
+If the second register contained a value larger than $R$, it would end up trashed despite the final clean-up operation.
+Our modular circuit constructions do not guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
+To avoid this problem, we require that the registers have minimal size $n = \lceil \lg_2(R-1) \rceil$ and that the second register's most-significant-bit (MSB) be a clean qubit initialized to zero.
 
@@
 	High-level period finding circuit \cite{Shor1999}.
-	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(N)$ controls the precision of the phase estimation step.
-    The triangles indicate how many ancilla are needed``behind the scenes", by our constructions, to perform an operation.
+	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(n)$ controls the precision of the phase estimation step.
+    The triangles indicate how many ancilla are needed ``behind the scenes", by our constructions, to perform an operation.
 	Recovering the period requires classically processing the sampled output with a continued fractions algorithm.
@@
 
-\subsection{Controlled Modular Multiply-Both-Ways}
+\subsection{Modular Bimultiplication}
 
-As in \cite{beauregard2003}, we perform controlled modular multiplication with modular multiply-accumulates into a second register.
-However, by using a third multiply-accumulate, we end up with a compound multiply-and-inverse-multiply operation instead of requiring a clean ancilla register.
+As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-multiply-accumulates operations.
+However, because the second register is dirty, we need three scaled-multiply-accumulates instead of the usual two.
 
-As with all other modular arithmetic constructions in this paper, we require that the input registers be in range (i.e. have values below $R$).
+As with all other modular arithmetic circuits in this paper, we make no guarantees about the circuit's behavior when either of the input registers is out of range (i.e. have values not below $R$).
 
@@
 By multiply-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
-Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(bK^{-1}, b+aK)$.
-Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(bK^{-1}, aK)$.
+Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(-bK^{-1}, b+aK)$.
+Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
 Finally, we swap to $(aK, bK^{-1})$.
-The overall effect is to multiply the first register by $K$ and the second register by $K^{-1}$, as desired.
+The overall effect is to multiply the first register by $K$ and the second register by $-K^{-1}$, as desired.
 See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
@@
   \caption{\em
-    Performing a controlled modular multiply-both-ways with three controlled modular multiply-accumulates.
+    Performing a modular bimultiplication with three scaled-multiply-accumulates and a swap.
     $K$ must have a multiplicative inverse modulo $R$.
@@
 
-\subsection{Controlled Modular Multiply-Accumulate}
+\subsection{Modular Scaled-Multiply-Accumulation}
 
-To perform multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
+To perform scaled-multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
 We right-shift (i.e.
@@
 
-A single dirty bit is required by the conditional offset operations, and two dirty bits are required by the modular doublings, but there are more than enough unused bits available to borrow, so the overall circuit doesn't require any dirty bits.
+The conditional offset operations need one dirty bit, and the modular doubling need two, but there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
 
@@
   \caption{\em
-    Reducing a controlled modular multiply-accumulate into the modular equivalent of shift-and-add.
+    Reducing a modular multiply-accumulate into the modular equivalent of shift-and-add.
     Assumes $R$ is odd.
@@
 
-\subsection{Controlled Pivot-Flips}
+\subsection{Modular Doubling}
 
-We will implement both modular doubling and modular offset/addition in terms of ``pivot-flips".
+To multiply a register by 2 modulo an odd $R$, we use pivot-flips (see section \ref{sec:pivot-flips}) and subtraction to align the two halves of the input space on the MSB boundaries.
+Then we finish the job with a left-rotate.
+See figure \ref{fig:modular-double}.
+
+To invert the doubling, i.e. to divide by 2 instead of multiplying by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
+Except for period finding, all circuits presented in this paper can be inverted (e.g. from addition to subtraction) with recurse-and-reverse.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-double.png}
+  \caption{\em
+    Controlled modular doubling in $O(c + n \lg n)$ gates, $O(c + n)$ depth, with 2 dirty ancilla.
+    $R$ must be odd.
+  }
+  \label{fig:modular-double}
+\end{figure}
+
+\subsection{Modular Addition / Offset}
+
+As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot-flips (see section \ref{sec:pivot-flips}).
+To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
+See figure \ref{fig:controlled-modular-add} for the circuit.
+
+Interestingly, modular addition only requires a dirty bit if it's {\em not} controlled.
+And when performing modular offset (i.e. adding a compile-time constant to a register), {\em two} dirty bits are required whether or not the operation is controlled.
+See figure \ref{fig:controlled-modular-offset}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
+  \caption{\em
+     Modular addition of $A \pmod{R}$ can be done with three pivot flips.
+     One at $R-A$, then one at $R$, then one at $A$.
+     Requires $A \leq R$ (or else the operation will trash the target).
+   }
+  \label{fig:mod-add-from-pivot-flip-bars}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
+  \caption{\em
+    Controlled modular addition.
+    The circuit as a whole only needs a dirty bit if no control qubits are available to be borrowed by the uncontrolled pivot-flip operations.
+  }
+  \label{fig:controlled-modular-add}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
+  \caption{\em
+    Controlled modular offset.
+  }
+  \label{fig:controlled-modular-offset}
+\end{figure}
+
+\subsection{Pivot-Flips} \label{sec:pivot-flips}
+
+We implemented both modular doubling and modular offset/addition in terms of a non-standard operation we call ``pivot-flips".
 A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
@@
 To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
-However, because $x \rightarrow \lnot(x - A)$ is its own inverse and doesn't change whether a given value is below the pivot, the operation can be toggle-controlled by the comparison against $A$.
+However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'': unless a controlling qubit is toggled between two applications of the operation, it will have no effect or undo itself.
+Also, because the operation doesn't change whether a given value is below the pivot, the controlling qubit can be toggled by a comparison of the {\em target} against the pivot $A$.
 See figure \ref{fig:controlled-pivot-flip} for the construction.
@@
 This distinction matters because the constant-offset case uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
-Also, the constant-offset case requires an extra dirty bit unless quantum operations are used (e.g. by using the technique demonstrated in figure \ref{fig:bootstrap-ancilla}).
+Also, the constant-offset case requires an extra dirty bit (unless quantum operations are used, e.g. see figure \ref{fig:bootstrap-ancilla}).
 
@@
 
-\subsection{Modular Doubling}
-
-To multiply a register by 2 modulo an odd $R$, we use pivot-flips and subtraction to align the two halves of the input space on the MSB boundaries.
-Then we left-rotate.
-Note that this trashes out-of-range values larger than $R$ instead of leaving them alone.
-See figure \ref{fig:modular-double}.
-
-To invert the doubling, i.e. to divide by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
-Recurse-and-reverse works on all circuits presented in this paper (e.g.
-to perform subtraction instead of addition, or to decrement instead of increment), except period-finding, because the circuits don't generate garbage or use measurement.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/modular-double.png}
-  \caption{\em
-    Modular doubling.
-    Assumes $R$ is odd.
-    Requires 2 dirty ancilla, or 1 if using quantum operations.
-  }
-  \label{fig:modular-double}
-\end{figure}
-
-\subsection{Controlled Modular Addition and Offset}
-
-As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot flips.
-To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
-See figure \ref{fig:modular-add} for the circuit.
-
-Although our modular addition circuit requires a dirty bit, {\em controlled} modular addition does {\em not}.
-See figure \ref{fig:controlled-modular-add}.
-However, when performing modular offset (i.e. adding a compile-time constant to a register), we instead require two dirty bits.
-See figures \ref{fig:modular-offset} and \ref{fig:controlled-modular-offset}.
-
-Be aware that pivot-flip-based-addition assumes that {\em both} the input and target register are in range.
-If the input is greater than or equal to $R$, the pivot-flips will trash the target register instead of adding the input mod $R$ into it.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
-  \caption{\em
-     Modular addition of $A \pmod{R}$ with three pivot flips.
-     One at $R-A$, then one at $R$, then one at $A$.
-     Requires $A \leq R$ (or else the operation will trash the target).
-   }
-  \label{fig:mod-add-from-pivot-flip-bars}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/modular-addition.png}
-  \caption{\em Non-constant modular addition from pivot-flips.}
-  \label{fig:modular-add}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
-  \caption{\em
-    Controlled modular addition.
-    The pivot-flips don't need to be controlled so, unlike the uncontrolled modular addition construction, this construction requires no dirty bits.
-  }
-  \label{fig:controlled-modular-add}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/modular-offset.png}
-  \caption{\em Modular offset.}
-  \label{fig:modular-offset}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
-  \caption{\em
-    Controlled modular offset.
-  }
-  \label{fig:controlled-modular-offset}
-\end{figure}
-
 \subsection{Comparison}
@@
 
-\subsection{Addition and Offset}
+\subsection{Addition / Offset}
 
@@
   \label{fig:inlineadder}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-addition.png}
+  \caption{\em
+  	Reducing controlled addition to uncontrolled addition.
+  	Uses $O(N + C)$ gates and depth.
+  }
+  \label{fig:controlled-addition}
 \end{figure}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/increment-odd.png}
-  \caption{\em Odd-sized $O(N)$ increment with 1 dirty ancilla.}
-  \label{fig:increment-odd}
-\end{figure}
-
 Because the adder/subtracter construction used by the incrementer in figure \ref{fig:increment-odd} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
@@
 That construction is shown in figure \ref{fig:increment-even}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/increment-even.png}
-  \caption{\em Even-sized $O(N)$ increment with 1 dirty ancilla.}
-  \label{fig:increment-even}
-\end{figure}
 
@@
 
+\subsection{Bit Swaps, Rotations, and Reversals}
+
+Bit permuting operations are easy to overlook in circuits, because they can be emulated by re-labelling qubits.
+But some of our circuit diagrams have used controlled bit permutations, requiring actual operations, so we provide the constructions for completeness.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-bit-rotate.png}
+  \caption{\em
+    A controlled bit rotation / bit swap is three $O(N+C)$ controlled bit-reverses.
+  }
+  \label{fig:dependencies}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-reverse.png}
+  \caption{\em
+    An inline controlled bit order reversal using $O(N + C)$ gates.
+    Each XOR operation is a series of CNOTs (note that the inputs have opposite endian-ness to the outputs).
+    For reversals on an even number of bits, swap the center two bits beforehand then use one of them as the center workspace bit from this construction.
+  }
+  \label{fig:dependencies}
+\end{figure}
+
+\subsection{Multi-Nots}
+
+Several of our constructions have used CNOTs with many controls and many targets.
+To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others. [[[ FIGURE ]]]
+We then use a dirty ancilla (i.e. one of the other targets) to efficiently reduce the big CNOT into Toffolis.
+
 \section{Overview and Improvements} \label{sec:costs}
@@
 
-Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of multiply-both-ways for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
+We tested individual constructions in Quirk [[[ https://github.com/Strilanc/Quirk ]]] and the full construction down to Toffoli gates in ProjectQ [[[ https://github.com/ProjectQ-Framework/ProjectQ ]]] [[[ link to code ]]].
+
+Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
 
 Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
-Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and in the context of period-finding there are many qubits available to borrow whenever a pivot-flip is needed.
-This improvement saves a qubit, reducing the total number of qubits required for period finding to $2n+1$.
+Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
+This improvement saves a qubit, reducing the total number of qubits required for period finding from $2n+2$ to $2n+1$.
 
-Previous modular multiplication constructions used a clean ancilla register \cite{haner2016}.
-For these constructions, any garbage in the ancilla register would infect the target register and trash it.
-We use an extra multiply-accumulate to cancel out garbage entering the target register, but the ancilla register is still multiplified by the inverse of whatever the target register is multiplied by, creating a "multiply-both-ways" circuit.
-We don't want to multiply the ancilla register, since it must be returned to its starting state before our circuit ends, but we can fix the problem with a final cleanup multiplication informed by a measurement of the main register.
-This improvement allows $n-1$ of the qubits in the ancilla register, all except the MSB, to be dirty.
-The MSB must be zero to ensure values are not larger than the modulus.
+Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}, where any garbage in the ancilla register would have infected the target register and trashed it.
+We deal with the trash with an extra multiply-accumulate per modular multiplication, and cancel the overall effects on the ancilla register with a clean-up multiplication informed by measuring the main register at the end of the circuit.
+This improvement allows $n-1$ of the qubits in the ancilla register  to be dirty (only the MSB must be zero; to ensure the register's value is in range).
 
@@
 
-We presented Toffoli-based arithmetic circuits that reduce the number of qubits required to perform Shor's algorithm by 1, and the number of {\em clean} qubits required by an additional $n-1$, compared to the best previous constructions.
+We presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm by 1, and allowed an additional $n-1$ qubits to be dirty.
 

### 2017-04-24 paper.tex
Third draft

@@
 The introduction describes the various sections and also itself for no discernible reason.
-In section \ref{sec:construct} we iteratively reduce period finding into simpler and simpler arithmetic operations, until we reach constant-sized gates, while tracking the number of required dirty ancilla.
-Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented arithmetic circuit constructions.
+In section \ref{sec:construct} we describe all the circuit constructions needed to reduce period finding into constant-sized gates, while tracking the number of required dirty ancilla.
+Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
 Finally, section \ref{sec:conclusion} concludes and summarizes.
@@
 Multi-qubit arithmetic operations always have the least significant bit towards the top.
+All dirty qubit counts are for classical Toffoli-based constructions, without using the quantum ancilla bootstrap trick noted in figure \ref{fig:bootstrap-ancilla}.
 For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray, when applicable.
-For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we provide controllable constructions of that operation that scale well with the number of controls.
+For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we always provide controllable constructions that scale linearly with the number of controls.
 
@@
 A high-level circuit for period finding, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
+The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $p$ of that operation.
+That is to say, w.l.o.g. the state ends up being $|\psi_{1,x}\rangle = \sqrt{2^{-n}/p} \sum_{k=0}^{2^n/p} |pk + x \rangle$ for some $x$.
+The circuit then applies an inverse Fourier transform and samples the output.
+Using a classical continued fraction algorithm, the sample can be used to estimate the period.
 
-Because period-finding measures all qubits immediately after performing a quantum Fourier transform, most of the transformed qubits can be measured earlier than shown.
-In fact, each qubit can be measured so early that the next does not even need to be initialized yet!
+Because period-finding measures all qubits immediately after performing a QFT, most of the transformed qubits can be measured earlier than shown.
+In fact, each qubit can be measured so early that the next qubit needed for the QFT does not even need to be initialized yet!
 Only one of the qubits in the phase-estimation register needs to be present at a time, and so the phase register can be reduced to a single repeatedly-used qubit \cite{beauregard2003}.
@@
 The period-finding circuit in figure \ref{fig:period-finding-solo-phase-qubit} is our high-level starting point.
-It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
+It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
 The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla using our constructions).
@@
 We perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
-However, to allow the second register to be dirty, it is convenient to pair each modular multiplication with an inverse and negated multiplication on another register.
+However, to allow the second register to be mostly dirty, we note that each modular multiplication on the first register causes an inverse and negated multiplication on the second register.
 We will refer to this combined operation as a ``bimultiply".
@@
 Note that we must preserve the state of the second register, but bimultiplication trashes that state with many multiplications by unknown constants.
-However, because the first register is initialized to store 1 and gets multiplied by constants inverse to the ones trashing the second register, we can fix the second register by measuring the first register and using its value to perform a clean-up bimultiplication after the usual end of the circuit.
+However, because the first register is initialized to $|1\rangle$ and gets multiplied by constants inverse to the ones trashing the second register, we can fix the second register by measuring the first register and using its value to perform a clean-up bimultiplication after the usual end of the circuit.
 Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
 
+Our modular circuit constructions do not guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
 If the second register contained a value larger than $R$, it would end up trashed despite the final clean-up operation.
-Our modular circuit constructions do not guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
 To avoid this problem, we require that the registers have minimal size $n = \lceil \lg_2(R-1) \rceil$ and that the second register's most-significant-bit (MSB) be a clean qubit initialized to zero.
@@
     The triangles indicate how many ancilla are needed ``behind the scenes", by our constructions, to perform an operation.
-	Recovering the period requires classically processing the sampled output with a continued fractions algorithm.
+	Recovering the period requires classical post-processing of the sampled output.
   }
@@
 
-As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-multiply-accumulates operations.
-However, because the second register is dirty, we need three scaled-multiply-accumulates instead of the usual two.
+As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-accumulate operations.
+However, because the second register is dirty, we need three scaled-accumulates instead of the usual two.
 
@@
 Suppose that the two registers start in the state $(a, b)$, working modulo $R$.
-By multiply-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
-Then a multiply-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(-bK^{-1}, b+aK)$.
-Next, we cancel the $b$ term in the second register by multiply-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
+By scale-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
+Then a scale-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(-bK^{-1}, b+aK)$.
+Next, we cancel the $b$ term in the second register by scale-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
 Finally, we swap to $(aK, bK^{-1})$.
@@
 In the case where $K$ has no multiplicative inverse modulo $R$, this construction will not work (it would define an invalid irreversible operation).
-However, that would mean $K$ is a factor of $R$, a case which the very lucky user can check and handle classically before invoking a quantum period finding circuit.
+However, that would mean $K$ is a factor of $R$, a case that the very lucky user can check for and handle classically before invoking a quantum period finding circuit.
 
@@
   \caption{\em
-    Performing a modular bimultiplication with three scaled-multiply-accumulates and a swap.
+    Controlled modular bimultiplication, with a constant multiplier $K$, using three modular scaled-accumulates and a swap.
     $K$ must have a multiplicative inverse modulo $R$.
+    Uses no ancilla, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
 
-\subsection{Modular Scaled-Multiply-Accumulation}
+\subsection{Modular Scaled-Accumulation}
 
-To perform scaled-multiply-accumulation, we use a shift-and-add approach similar to \cite{beauregard2003}.
-We right-shift (i.e.
-divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e.
-multiplying it by 2) and adding $K$ into the target conditioned on the most significant bit of the input, then the next most significant bit, and so forth.
+To perform scaled-accumulate, we use a shift-and-add approach similar to \cite{beauregard2003}.
+We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying it by 2) and adding $K$ into the target.
+We condition the first modular addition on the most significant bit of the input, the second addition on the next most significant bit, and so forth.
 (The more significant bits go first because their effects must be hit by more left-shifts.)
-See figure \ref{fig:controlled-modular-multiply-accumulate}.
+See figure \ref{fig:controlled-modular-scale-accumulate}.
 
-The conditional offset operations need one dirty bit, and the modular doubling need two, but there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
+The conditional offset operations need one dirty bit, and the modular doublings need two, but for non-trivial $n$ there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
 
@@
 However, given that in the intended use case $R$ is a number to be factored, it is reasonable to require callers to have factored out multiples of two beforehand.
+
+As will all circuits presented in this paper (except the period-finding circuit), we define the construction of the inverse operation to be the same circuit but run in reverse order with each sub-operation inverted.
+A decrement is just a reversed increment, a division is just a reversed multiplication, etc.
 
@@
   \caption{\em
-    Reducing a modular multiply-accumulate into the modular equivalent of shift-and-add.
-    Assumes $R$ is odd.
+    Reducing a controlled modular scale-accumulate, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
+    Requires the modulus $R$ to be odd.
+    Uses no ancilla, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
-  \label{fig:controlled-modular-multiply-accumulate}
+  \label{fig:controlled-modular-scale-accumulate}
 \end{figure}
@@
 
-To multiply a register by 2 modulo an odd $R$, we use pivot-flips (see section \ref{sec:pivot-flips}) and subtraction to align the two halves of the input space on the MSB boundaries.
-Then we finish the job with a left-rotate.
+To multiply a register by 2 modulo an odd $R$, we use pivot-flips (see section \ref{sec:pivot-flips}) and subtraction to align the two halves of the input space to MSB boundaries.
+After the two halves are aligned, a left-rotate completes the operation.
 See figure \ref{fig:modular-double}.
-
-To invert the doubling, i.e. to divide by 2 instead of multiplying by 2, recurse and reverse: invert each operation in the construction, then apply the operations in reverse order. 
-Except for period finding, all circuits presented in this paper can be inverted (e.g. from addition to subtraction) with recurse-and-reverse.
 
@@
     $R$ must be odd.
+    Uses 2 dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
 
-Interestingly, modular addition only requires a dirty bit if it's {\em not} controlled.
-And when performing modular offset (i.e. adding a compile-time constant to a register), {\em two} dirty bits are required whether or not the operation is controlled.
+Interestingly, modular addition can use its own controls as dirty bits that would otherwise be required.
+For modular offset (i.e. adding a compile-time constant into a register), the two dirty bits are required whether or not the operation is controlled.
 See figure \ref{fig:controlled-modular-offset}.
@@
      One at $R-A$, then one at $R$, then one at $A$.
-     Requires $A \leq R$ (or else the operation will trash the target).
+     Requires $A \leq R$.
    }
@@
   \caption{\em
-    Controlled modular addition.
-    The circuit as a whole only needs a dirty bit if no control qubits are available to be borrowed by the uncontrolled pivot-flip operations.
+    Controlled modular addition construction based on pivot-flips.
+    Uses $2-c$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
   \caption{\em
-    Controlled modular offset.
+    Controlled modular offset construction based on pivot-flips.
+    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
 
-We implemented both modular doubling and modular offset/addition in terms of a non-standard operation we call ``pivot-flips".
+We implemented both modular doubling and modular offset/addition in terms of a non-standard operation we call a ``pivot-flip".
 A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
@@
 To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
-However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'': unless a controlling qubit is toggled between two applications of the operation, it will have no effect or undo itself.
-Also, because the operation doesn't change whether a given value is below the pivot, the controlling qubit can be toggled by a comparison of the {\em target} against the pivot $A$.
-See figure \ref{fig:controlled-pivot-flip} for the construction.
-For states that are less than the pivot $A$, the comparison is true, the ancilla is toggled both times, and exactly one of the controlled subtract-and-inverts will fire.
+However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'', where the operation undoes itself unless a controlling qubit is toggled between two applications.
+Also, because the operation doesn't change whether a given value is below the pivot, the controlling qubit can be toggled by a comparison of the target against the pivot $A$ even though the target is being operated on.
+
+See figure \ref{fig:controlled-pivot-flip} for our pivot-flip construction.
+For states that are less than the pivot $A$, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled subtract-and-inverts will fire.
 For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.
@@
 Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
-This distinction matters because the constant-offset case uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
-Also, the constant-offset case requires an extra dirty bit (unless quantum operations are used, e.g. see figure \ref{fig:bootstrap-ancilla}).
+This distinction matters because the constant-offset construction uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
 
@@
     Controlled pivot-flip circuit with an enregistered pivot.
+    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Controlled pivot-flip circuit with a constant pivot.
-    The extra dirty ancilla, compared to the enregistered case, comes from the controlled-offset construction.
+    Uses $2$ dirty ancilla, $O(c + n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
   \label{fig:controlled-pivot-flip}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-offset.png}
-  \caption{\em
-      Offset circuit from \cite{haner2016}, but modified to include controls.
-      Uses $O(n \lg n)$ gates and $O(n)$ depth (by overlapping the recursive cases).
-      Note that, if $c$ is not $O(1)$, the recursion introduces additional non-linear costs.
-  }
-  \label{fig:controlled-offset}
 \end{figure}
@@
 We implement comparisons as in \cite{takahashi2005}, using an addition followed by a subtraction that is slightly smaller to clear all changes except an overflow signal into the target.
-See figures \ref{fig:comparison-less} and \ref{fig:comparison-less-const}.
-
-To perform a $\leq$ instead of a $<$, simply add/subtract $A+1$ instead of $A$.
-(Apply the extra increment/decrement to the target register, not the input register, since incrementing the input register may wrap it to zero.)
-
-Beware that, in these constructions, if the $B$ register is smaller than the $A$ register, the high bits of $A$ are lost and the construction breaks.
-To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
-Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by the high bits of $A$ all being off and otherwise (if any of the high bits are on) concluding $A>B$ and toggling the result bit as appropriate for the comparison being implemented.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/comparison-less.png}
-  \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
-  \label{fig:comparison-less}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/comparison-less-const.png}
-  \caption{\em Toggling a target bit if a register is less then a constant, using a dirty ancilla and $O(n \lg n)$ gates.}
-  \label{fig:comparison-less-const}
-\end{figure}
+For comparisons against another register this uses $1$ dirty ancilla (if controlled, otherwise no ancilla), $O(c + n)$ gates, and $O(c + n)$ depth.
+For comparisons against a constant, the number of gates increases to $O(c + n \lg n)$ and the dirty ancilla is always required.
 
@@
 
-In \cite{van2004}, a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla, is described.
-If the ancilla is not initialized in the off state, the target register ends up storing $a+b+1$ instead of $a+b$.
-For clarity, we review how to elide the ancilla.
-
-The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the input and target registers before and after the addition when the ancilla is on.
-Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
-Temporarily inverting the input effectively turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.
-
-Once the ancilla can be dirty, the MSB of the input register can play the role of the ancilla, as in \cite{takahashi2005} and shown in figure \ref{fig:inlineadder}.
+\cite{takahashi2005} provides a reversible adder circuit that uses $O(n)$ gates, $O(n)$ depth, and no ancilla.
+We show an equivalent circuit in figure \ref{fig:inlineadder}.
+Note that the construction uses the input register as workspace, so it breaks when the input is a compile-time constant or when the target register is larger than the input register.
 
@@
 
+When the value to be added into a register is a compile-time constant (i.e. when applying an offset gate), we use the construction from \cite{haner2016}.
+Their offset circuit, shown in figure \ref{fig:offset}, uses $O(n \lg n)$ gates, $O(n)$ depth, and a dirty ancilla.
+
+When the target register is larger than the input register, three tweaks based on increment/decrement gates can be used (instead of increasing the asymptotic costs).
+(Note that, to avoid cyclic dependencies, the increment and decrement constructions described in the next subsection will only use the same-register-size adder.)
+First, the carry signal is forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
+Second, to free the MSB of the input register for use as a dirty ancilla holding the carry signal, we add it into the target ahead of time with a controlled increment.
+Finally, when the MSB is on its role as carry signal causes an extra increment of the target.
+This extra increment is cancelled with a controlled decrement.
+See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
+
+To control addition gates, and offset gates, we do an uncontrolled addition followed by an uncontrolled subtraction bordered by NOT gates that fire when the controls are satisfied.
+When the NOT gates don't fire, the subtraction cancels the addition.
+When the NOT gates do fire, the subtraction is inverted into an addition and $2 \cdot a$ is added to the target.
+To add $a$ instead of $2 \cdot a$ into the target, we temporarily prepend a dirty bit onto the target register.
+See \ref{fig:controlled-addition}.
+
 \begin{figure}
   \centering
-  \includegraphics[width=\linewidth]{assets/controlled-addition.png}
+  \includegraphics[width=\linewidth]{assets/offset.png}
   \caption{\em
-  	Reducing controlled addition to uncontrolled addition.
-  	Uses $O(N + C)$ gates and depth.
+      Offset circuit from \cite{haner2016}.
+      Uses $1$ dirty ancilla, $O(n \lg n)$ gates, and $O(n)$ depth (by overlapping the recursive cases).
   }
-  \label{fig:controlled-addition}
+  \label{fig:offset}
 \end{figure}
-
-Because this adder construction uses the input register as workspace, it breaks when the input and target registers differ in size.
-When the input register is larger, the high bits of the input register can simply be ignored.
-But, when the target register is larger, there is insufficient workspace and also the logical-negation trick that allowed the ancilla to be dirty breaks.
-
-To work around the target register being larger, three tweaks are needed.
-Each uses a controlled increment or decrement gate, which will be described in the next subsection.
-(Note that, to avoid cyclic dependencies, the increment and decrement constructions will only use the same-register-size adder.)
-
-First, the carry signal must be forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
-Second, to free the MSB of the input register for use as the carry signal, we add it into the target ahead of time with a controlled increment.
-Finally, when the MSB is on instead of off its role as carry signal causes an extra increment of the target.
-This extra increment is cancelled with a controlled decrement.
-See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
 
@@
       Adder with target larger than source, using no ancilla.
-      Uses $O(N)$ gates and depth.
+      Uses $O(n)$ gates and depth.
       The increment and decrement gates all have at least one free wire to borrow as a dirty ancilla.}
@@
 
-Another problem with using the input register as workspace appears when adding a constant offset into a target: there {\em isn't} an input register.
-We use the construction from \cite{haner2016}, which works for constant offsets but increases the gate cost from $\Theta(n)$ to $\Theta(n \lg n)$ and also requires a dirty ancilla instead of no ancilla (unless using quantum operations).
-See figure \ref{fig:controlled-offset}.
-We note that our paper would have significantly fewer diagrams if offset and addition didn't have differing costs.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-addition.png}
+  \caption{\em
+  	Reducing controlled addition to uncontrolled addition.
+  	Uses $O(c + n)$ gates, $O(c + n)$ depth, and a dirty ancilla.
+  	When applied to offset gates, uses $O(c + n \lg n)$ gates, $O(n + c)$ depth, and two dirty ancilla.
+  }
+  \label{fig:controlled-addition}
+\end{figure}
 
@@
 Therefore the target was incremented by 2.
-To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register as shown in \ref{fig:increment-odd}.
+To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register.
 
-Because the adder/subtracter construction used by the incrementer in figure \ref{fig:increment-odd} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
-For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
-That construction is shown in figure \ref{fig:increment-even}.
-
-Increment gates can be efficiently controlled by using another invert-input-and-output trick.
-See figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}.
+Note that, to avoid a cyclic dependency, we must require that the additions and subtractions use input and target registers of the same size.
+So the construction described in the previous paragraph only works for odd-sized registers.
+For even-sized registers, we simply handle the LSB separately.
+Also, we support contorls with another conditional-inversion trick.
+See figure \ref{fig:controlled-increment-odd}.
 
@@
   \includegraphics[width=\linewidth]{assets/controlled-increment-odd.png}
-  \caption{\em Odd-sized $O(N)$ controlled increment with 1 dirty ancilla.}
+  \caption{\em
+    Odd-sized controlled increment.
+    For the even-sized case, separate the LSB from the rest of the register, increment the rest of the register using the LSB as an extra control, then toggle the LSB.
+    Uses $O(c+n)$ gates, $O(c+n)$ depth, and 1 dirty ancilla.
+  }
   \label{fig:controlled-increment-odd}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-increment-even.png}
-  \caption{\em Even-sized $O(N)$ controlled increment with 1 dirty ancilla.}
-  \label{fig:controlled-increment-even}
 \end{figure}
@@
 
-Bit permuting operations are easy to overlook in circuits, because they can be emulated by re-labelling qubits.
-But some of our circuit diagrams have used controlled bit permutations, requiring actual operations, so we provide the constructions for completeness.
+Bit permuting operations can usually be emulated by re-labelling qubits, so they are easy to overlook in circuits.
+But some of our circuit diagrams have used controlled bit rotations, requiring actual operations, so we provide the constructions for completeness.
 
@@
   \caption{\em
-    A controlled bit rotation / bit swap is three $O(N+C)$ controlled bit-reverses.
+    A controlled bit rotation / bit swap is three controlled bit-reverses.
+    Uses no ancilla, $O(n+c)$ gates, and $O(n+c)$ depth.
   }
@@
   \caption{\em
-    An inline controlled bit order reversal using $O(N + C)$ gates.
-    Each XOR operation is a series of CNOTs (note that the inputs have opposite endian-ness to the outputs).
+    An inline controlled bit order reversal.
+    Each XOR operation is a series of independent CNOTs (note that the inputs have opposite endian-ness to the outputs).
     For reversals on an even number of bits, swap the center two bits beforehand then use one of them as the center workspace bit from this construction.
+    Uses no ancilla, $O(c + n)$ gates, and $O(c + n)$ depth.
   }
@@
 Several of our constructions have used CNOTs with many controls and many targets.
-To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others. [[[ FIGURE ]]]
+To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others.
 We then use a dirty ancilla (i.e. one of the other targets) to efficiently reduce the big CNOT into Toffolis.
@@
 In figure \ref{fig:dependencies} we show a dependency graph of the constructions discussed in this paper.
-The total cost of period-finding, using this construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
+The total cost of period-finding, using the full construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
 We perform $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
 
-We tested individual constructions in Quirk [[[ https://github.com/Strilanc/Quirk ]]] and the full construction down to Toffoli gates in ProjectQ [[[ https://github.com/ProjectQ-Framework/ProjectQ ]]] [[[ link to code ]]].
+We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
+[[[ insert link to code ]]] [[[ maybe make a plot of gate counts vs size?
+]]] [[[ definitely need another constant-factor joke ]]]
 
@@
 Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
-Pivot flips also require an ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
+Pivot flips also require ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
 This improvement saves a qubit, reducing the total number of qubits required for period finding from $2n+2$ to $2n+1$.
@@
 
-Previous published incrementers (not counting a rough unpublished version of our construction \cite{gidney2015} being cited by \cite{haner2016}) required either $O(n^2)$ gates or $O(n)$ ancilla \cite{draper2000, barenco1995}.
+Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $O(n)$ ancilla \cite{draper2000, barenco1995}.
 Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
@@
   \caption{\em
-    Graph of non-trivial dependencies between constructions, from period finding to fixed-size gates.
-    Edges are labelled by an upper bound on how many times the target is used by the source.
+    Meta-figure showing simplified dependencies between constructions, from period finding to constant-size gates.
+    Edge labels indicate which constructions use their dependencies more than a constant number of times.
   }
@@
 
-\section{Temporary Test Circuit Links}
-
-Convenience links for myself when checking that things work.
-
-\href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
-{5+5-bit inline adder}
-(or \href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
-{with test harness})
-
-\href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
-{4-bit incrementer w/ 4 dirty}
-
-\href
-{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%5D%7D}
-{8-bit incrementer w/ 1 dirty}
-(or \href
-{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting9%22%5D%2C%5B%22QFT6%22%5D%2C%5B1%2C1%2C1%2C%22QFT6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec8%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A06%22%5D%2C%5B%22QFT%E2%80%A06%22%5D%2C%5B%22Uncounting9%22%5D%5D%7D}
-{with test harness})
-
-\href
-{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%5D%7D}
-{9-bit incrementer w/ 1 dirty}
-(or \href
-{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec9%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
-{with test harness})
-
-\href
-{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9B%22%2C%22Z%5E%C2%BC%22%2C%22H%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22inc4%22%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B1%2C%22Z%5E-%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E-%E2%85%9B%22%2C%22Z%5E-%C2%BC%22%2C%22H%22%5D%5D%7D}
-{Quantum incrementer ancilla bootstrap}
-
-NOTE: have to open these next ones against local version, because online version doesn't have the comparison gates
-
-\href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
-{4+4-Bit $<$ PivotFlip}
-(or \href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
-{with test harness})
-
-\href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
-{4+4-Bit $\leq$ PivotFlip}
-(or \href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
-{with test harness})
-
-\href
-{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B%22X%22%2C1%2C1%2C%22X%22%2C%22Counting4%22%2C1%2C1%2C1%2C%22QFT4%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%7C0%E2%9F%A9%E2%9F%A80%7C%22%5D%2C%5B%22inputB4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22~a82g%22%5D%5D%2C%22gates%22%3A%5B%7B%22id%22%3A%22~a82g%22%2C%22name%22%3A%22IGNORE%22%2C%22matrix%22%3A%22%7B%7B1%2C0%7D%2C%7B0%2C1%7D%7D%22%7D%5D%7D}
-{Modular Addition with test Harness}
-
 \end{document}

### 2017-05-08 test/decompositions/addition_test.py
Start table in README

@@ -21,7 +21,6 @@ from dirty_period_finding.decompositions.addition_rules import (
 )
 from dirty_period_finding.extensions import (
     LimitedCapabilityEngine,
-    Swap,
     AutoReplacerEx,
 )
 from dirty_period_finding.gates import (

### 2017-05-08 paper/doc/paper.tex

--- 2017-04-24 paper.tex
@@
 
-See figure \ref{fig:controlled-pivot-flip} for our pivot-flip construction.
+See figures \ref{fig:controlled-pivot-flip} and \ref{fig:controlled-const-pivot-flip} for the circuit diagrams.
 For states that are less than the pivot $A$, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled subtract-and-inverts will fire.
@@
   }
-  \label{fig:controlled-pivot-flip}
+  \label{fig:controlled-const-pivot-flip}
 \end{figure}
@@
 Bit permuting operations can usually be emulated by re-labelling qubits, so they are easy to overlook in circuits.
-But some of our circuit diagrams have used controlled bit rotations, requiring actual operations, so we provide the constructions for completeness.
+But some of our circuit diagrams have used controlled bit rotations, requiring actual operations, so we provide the constructions in figures \ref{fig:bit-rotate} and \ref{fig:bit-reverse} for completeness.
 
@@
   }
-  \label{fig:dependencies}
+  \label{fig:bit-rotate}
 \end{figure}
@@
   }
-  \label{fig:dependencies}
+  \label{fig:bit-reverse}
 \end{figure}

### 2017-05-08 src/dirty_period_finding/extensions/_monkey_patching.py
Monkey-patch measure methods onto Qureg/Qubit

@@ -6,7 +6,7 @@ from projectq import MainEngine
 from projectq.backends import ResourceCounter
 from projectq.backends import Simulator
 from projectq.cengines import BasicEngine, UnsupportedEngineError
-from projectq.ops import Command
+from projectq.ops import Command, Measure
 from projectq.types import Qubit, Qureg
 
 
@@ -150,6 +150,22 @@ def MainEngine_init(self, backend=None, engine_list=None):
     self._measurements = dict()
     self.dirty_qubits = set()
 
+
+def measure_register(qureg):
+    if not len(qureg):
+        return 0
+    eng = [q.engine for q in qureg][0]
+    Measure | qureg
+    eng.flush()
+    return sum(int(qureg[i]) << i for i in range(len(qureg)))
+
+
+def measure_qubit(qubit):
+    Measure | qubit
+    qubit.engine.flush()
+    return bool(qubit)
+
+
 Command.__str__ = command_str
 Command.__repr__ = command_str
 Qureg.__str__ = qureg_str
@@ -159,3 +175,5 @@ Qubit.__del__ = lambda x: None
 Qureg.__int__ = lambda x: sum(int(x[i]) << i for i in range(len(x)))
 BasicEngine.__del__ = lambda x: None
 MainEngine.__init__ = MainEngine_init
+Qureg.measure = measure_register
+Qubit.measure = measure_qubit

### 2017-05-08 src/dirty_period_finding/find_period.py

@@ -22,6 +22,15 @@ from dirty_period_finding.extensions import (
 from dirty_period_finding.gates import *
 
 
+def measure_register(qureg):
+    if not len(qureg):
+        return 0
+    eng = [q.engine for q in qureg][0]
+    Measure | qureg
+    eng.flush()
+    return int(qureg)
+
+
 def sample_period(eng,
                   base,
                   modulus,
@@ -36,7 +45,7 @@ def sample_period(eng,
         modulus (int):
         precision (int):
             The number of iterations.
-        phase_qubit (Qubit):
+        phase_qubit (Qureg):
             A clean zero-initialized qubit.
         work_qureg (Qureg):
             A clean zero-initialized register with lg(modulus) bits.
@@ -62,9 +71,7 @@ def sample_period(eng,
         Z**frac | phase_qubit
         H | phase_qubit
 
-        Measure | phase_qubit
-        eng.flush()
-        b = bool(phase_qubit)
+        b = phase_qubit.measure()
         if b:
             X | phase_qubit
             frac += Fraction(1, 2 << i)
@@ -120,9 +127,7 @@ def main():
     for q in ancilla_qureg[:-1]:
         if random.random() < 0.5:
             X | q
-    Measure | ancilla_qureg
-    eng.flush()
-    before = int(ancilla_qureg)
+    before = ancilla_qureg.measure()
 
     p = sample_period(eng,
                       base=base,
@@ -132,9 +137,7 @@ def main():
                       work_qureg=eng.allocate_qureg(n),
                       ancilla_qureg=ancilla_qureg)
 
-    Measure | ancilla_qureg
-    eng.flush()
-    after = int(ancilla_qureg)
+    after = ancilla_qureg.measure()
     print("before", before, "after", after)
     u = pow(base, p, modulus)
     print("result", p, "u", u if u < modulus / 2 else '-' + str(-u % modulus))

### 2017-05-08 doc/paper.tex
Better modular double

--- 2017-05-08 paper/doc/paper.tex
@@
 
-The conditional offset operations need one dirty bit, and the modular doublings need two, but for non-trivial $n$ there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
+The conditional offset and modular doubling operations need one dirty bit, but for non-trivial $n$ there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
 
@@
 
-To multiply a register by 2 modulo an odd $R$, we use pivot-flips (see section \ref{sec:pivot-flips}) and subtraction to align the two halves of the input space to MSB boundaries.
-After the two halves are aligned, a left-rotate completes the operation.
+To multiply a register by 2 modulo an odd $R$, we use controlled offsets and nots to shift the top half of the input space until it starts on the MSB boundary.
+Once that's done, a left-rotate completes the operation by turning the MSB into the LSB (converting the alignment into interleaving).
 See figure \ref{fig:modular-double}.
@@
     $R$ must be odd.
-    Uses 2 dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses 1 dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }

### 2017-05-08 src/dirty_period_finding/decompositions/comparison_rules.py
Typo fixes

@@ -99,13 +99,13 @@ def do_predict_overflow(offset, query_reg, target_bit, controls, dirty_reg):
         control ━━/━━━●━━━━━━━━━━━━━●━━━━━━━━━━━━●━━━
                       │             │            │
                  n-1  │  ┌──────┐   │  ┌──────┐  │
-         offset ══/═══╪══╡offset╞═══╪══╡offset╞══╪═══
-                      │  └──┬───┘   │  └──┬───┘  │
+                ══/═══╪══╡offset╞═══╪══╡offset╞══╪═══
+         offset       │  └──┬───┘   │  └──┬───┘  │
                 ──────●─────┼──────●┼●────┼─────●┼●──
                       │     │      │││    │     │││
                  n-1  │  ┌──┴───┐  │││ ┌──┴───┐ │││
-          query ━━/━━━┿━━┥query ┝━━┿┿┿━┥query ┝━┿┿┿━━
-                      │  └──┬───┘  │││ └──┬───┘ │││
+                ━━/━━━┿━━┥query ┝━━┿┿┿━┥query ┝━┿┿┿━━
+          query       │  └──┬───┘  │││ └──┬───┘ │││
                 ──────●─────┼──────⊕●⊕────┼─────⊕●⊕──
                       │     │       │     │      │
                  n-2  │  ┌──┴───┐   │  ┌──┴───┐  │

### 2017-05-09 doc/paper.tex
More quirk links in README
Addressing issues - Switched dep graph to be transitive reduction instead of just saying vaguely 'simplified' - Fixed href in projectq citation - Cited C standard for undefined behavior - Added missing dirty indicators to big-add - Added missing = signs to diagrams - Better modular double - Sign flip fix - Various diagram fixes

@@
 \em
-We present asymptotically size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
-We improve the number of qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, $n-1$ of which can be an unknown state, without increasing the asymptotic size or depth of the circuit.
+We present reversible classical circuits for performing various arithmetic operations aided by dirty ancilla in an unknown state that must be preserved.
+We improve the number of qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, without increasing the asymptotic depth of the circuit.
+We also allow $n-1$ of the $2n+1$ qubits to be dirty.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction} \label{sec:introduction}
 
 When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
-An ancilla is an extra bit or qubit available for use by a circuit as temporary workspace.
+An ancilla is an extra bit or qubit that is available for use by a circuit as temporary workspace.
 Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
@@
 
-Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
-When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla makes a useful intermediate goal.
-In this paper, at great cost to the constant factors in the gate count and depth, we reduce the number of qubits required to perform period finding from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits without increasing the asymptotic costs.
+One part of a circuit can borrow dirty bits from another part of the same circuit, so circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
+When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla is a useful intermediate goal.
+In this paper, at great cost to the constant factors in the gate count and depth, we reduce the number of qubits required to perform period finding on modular exponentiation from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits without increasing the asymptotic costs.
 
 Our paper is structured as follows.
-The introduction describes the various sections and also itself for no discernible reason.
-In section \ref{sec:construct} we describe all the circuit constructions needed to reduce period finding into constant-sized gates, while tracking the number of required dirty ancilla.
+Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions will follow.
+In section \ref{sec:construct} we describe the circuit constructions we use to convert period finding into constant-sized gates, while tracking the number of required dirty ancilla.
 Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
@@
 
-All constructions assume a 2s-complement representation of integers.
-Multi-qubit arithmetic operations always have the least significant bit towards the top.
-All dirty qubit counts are for classical Toffoli-based constructions, without using the quantum ancilla bootstrap trick noted in figure \ref{fig:bootstrap-ancilla}.
+All constructions in this paper use a 2s-complement representation of integers, and all diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
+Nearly all presented circuit constructions are classical, and the required dirty qubit counts we note and show in diagrams are in terms of those classical constructions.
+The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla from several of our constructions, but only in places that don't affect the overall total for period-finding.
+
 For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray, when applicable.
-For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we always provide controllable constructions that scale linearly with the number of controls.
+For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we nevertheless provide controllable constructions that scale linearly with the number of controls.
+For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a division is a reversed multiplication, and so forth).
+
+We do not attempt to give reasonable behavior to modular arithmetic circuits applied to out of range values (ones equal to or larger than the modulus).
+This {\em includes input values}.
+For example, if our modular addition construction ($x \mathrel{{+}{=}} y \pmod{R}$) is applied to a $y \geq R$, then we don't guarantee that $y \mod R$ will be added into $x$ or that the operation will commute with other additions or that anything sensible at all will happen.
+If the precondition $y < R$ is violated, what occurs is undefined behavior \cite{c11}.
 
@@
 
-A high-level circuit for period finding, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
-The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $p$ of that operation.
-That is to say, w.l.o.g. the state ends up being $|\psi_{1,x}\rangle = \sqrt{2^{-n}/p} \sum_{k=0}^{2^n/p} |pk + x \rangle$ for some $x$.
-The circuit then applies an inverse Fourier transform and samples the output.
-Using a classical continued fraction algorithm, the sample can be used to estimate the period.
+A high-level view of the circuit for period finding on modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
+The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $p$ of said operation.
+That is to say, w.l.o.g. the state ends up equal to $|\psi_{1,x}\rangle = \sqrt{2^{-n}/p} \sum_{k=0}^{2^n/p} |pk + x \rangle$ for some $x$.
+The circuit then applies an inverse Fourier transform to the state.
+Fourier transforming a uniform signal with period $p$ produces a spectrum with peaks near $N \cdot 0/p$, $N \cdot 1/p$, $N \cdot 2/p$, ..., $N \cdot (p-1)/p$.
+Shor's algorithm recovers the period $p$ by sampling values $s_i$ from said spectrum, then using a continued fractions algorithm to compute the denominator of a fraction closest to $s_i/N$ given some maximum denominator.
 
@@
 Figure \ref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
+It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
+Note that the only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register.
 
-The period-finding circuit in figure \ref{fig:period-finding-solo-phase-qubit} is our high-level starting point.
-It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
-The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register (requiring $n-1$ dirty ancilla and 1 clean ancilla using our constructions).
+We perform modular multiplication with multiply-accumulates and an ancilla register as in \cite{beauregard2003}.
+However, to allow our ancilla register to be mostly dirty, we extend the operation so that it has a well defined effect on the second register: multiplying by the inverse factor.
+We will refer to this combined multiply-A-and-inverse-multiply-B operation as a ``bimultiply".
 
-We perform modular multiplication with multiply-accumulates and a second clean register as in \cite{beauregard2003}.
-However, to allow the second register to be mostly dirty, we note that each modular multiplication on the first register causes an inverse and negated multiplication on the second register.
-We will refer to this combined operation as a ``bimultiply".
-
-Note that we must preserve the state of the second register, but bimultiplication trashes that state with many multiplications by unknown constants.
-However, because the first register is initialized to $|1\rangle$ and gets multiplied by constants inverse to the ones trashing the second register, we can fix the second register by measuring the first register and using its value to perform a clean-up bimultiplication after the usual end of the circuit.
+The bimultiplications we perform throughout the algorithm do change the value of the ancilla register, but after the usual end of the circuit we can undo the damage.
+The key insight is that the work register is initialized to $|1\rangle$ and gets multiplied by constants inverse to the ones trashing the dirty ancilla register.
+Instead of discarding the work register after performing phase estimation, we measure it to recover the fixup factor needed to restore the dirty register's original value.
 Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
 
-Our modular circuit constructions do not guarantee that out-of-range values (i.e. values not below the modulus $R$) are well behaved.
-If the second register contained a value larger than $R$, it would end up trashed despite the final clean-up operation.
-To avoid this problem, we require that the registers have minimal size $n = \lceil \lg_2(R-1) \rceil$ and that the second register's most-significant-bit (MSB) be a clean qubit initialized to zero.
+Because our modular circuit constructions would have undefined behavior if any register has a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
+We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e. clean) and that the registers be as small as possible (i.e. have size $n = \lceil \lg_2(R-1) \rceil$).
 
@@
   \caption{\em
-	Period finding with a single phase-estimation qubit and paired inverse multiplications.
-	The terminating bra $\langle 0 |$ indicates that the bottom wire will always end up in the state $|0\rangle$, without post-selection.
+	Period finding with a single phase-estimation qubit and paired inverse multiplications (``bimultiplications").
+	Uses $O(n^3 \lg n)$ gates, $O(n^3)$ depth, and no additional qubits beyond those shown.
   }
@@
 As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-accumulate operations.
-However, because the second register is dirty, we need three scaled-accumulates instead of the usual two.
+However, to allow the second register to be dirty, we use an additional scaled-accumulate as well as a negation operation.
+See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
 
-As with all other modular arithmetic circuits in this paper, we make no guarantees about the circuit's behavior when either of the input registers is out of range (i.e. have values not below $R$).
-
-Suppose that the two registers start in the state $(a, b)$, working modulo $R$.
+To show that the circuit works, suppose that the two registers start in the state $(a, b)$, working modulo $R$.
 By scale-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
@@
 Next, we cancel the $b$ term in the second register by scale-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
-Finally, we swap to $(aK, bK^{-1})$.
-The overall effect is to multiply the first register by $K$ and the second register by $-K^{-1}$, as desired.
-See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
+Finally, we swap the terms and negate the second term to get $(aK, bK^{-1})$ as desired.
 
 In the case where $K$ has no multiplicative inverse modulo $R$, this construction will not work (it would define an invalid irreversible operation).
-However, that would mean $K$ is a factor of $R$, a case that the very lucky user can check for and handle classically before invoking a quantum period finding circuit.
+However, that would mean $K$ is a factor of $R$; a case that the very lucky user can check for and handle classically before bothering to factor $R$ with a quantum computer.
 
@@
 
-To perform scaled-accumulate, we use a shift-and-add approach similar to \cite{beauregard2003}.
-We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying it by 2) and adding $K$ into the target.
+To perform scale-accumulate operations, we use a shift-and-add approach similar to \cite{beauregard2003}.
+We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying by $2 {\pmod R}$) and adding $K$ into the target.
 We condition the first modular addition on the most significant bit of the input, the second addition on the next most significant bit, and so forth.
@@
 
-The conditional offset and modular doubling operations need one dirty bit, but for non-trivial $n$ there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
+The conditional offset and modular doubling operations need a dirty bit, but for non-trivial $n$ there are more than enough unused bits available to borrow, so the circuit as a whole doesn't require any dirty bits.
 
-Note that the modular doublings require $R$ to be odd, since otherwise the operation would be irreversible and non-unitary.
+Note that the modular doublings require $R$ to be odd, since otherwise the operation would be irreversible.
 However, given that in the intended use case $R$ is a number to be factored, it is reasonable to require callers to have factored out multiples of two beforehand.
-
-As will all circuits presented in this paper (except the period-finding circuit), we define the construction of the inverse operation to be the same circuit but run in reverse order with each sub-operation inverted.
-A decrement is just a reversed increment, a division is just a reversed multiplication, etc.
 
@@
 
-To multiply a register by 2 modulo an odd $R$, we use controlled offsets and nots to shift the top half of the input space until it starts on the MSB boundary.
+To multiply a register by 2 modulo an odd $R$, we use controlled offsets and nots to align each half of the input state space against MSB boundaries.
 Once that's done, a left-rotate completes the operation by turning the MSB into the LSB (converting the alignment into interleaving).
@@
 To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
+(The pivot-flip at $R$ can be simplified because the values above $R$ don't matter.)
 See figure \ref{fig:controlled-modular-add} for the circuit.
@@
 
+\subsection{Modular Negation}
+
+To negate a number mod $R$, we need to reverse the order of the states $|1\rangle$ to $|R-1\rangle$.
+We do so by temporarily moving $|0\rangle$ out the way with a decrement, pivot-flipping at $R-1$, then undoing the decrement.
+See figure \ref{fig:negate-mod}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/negate-mod.png}
+  \caption{\em
+    Controlled modular negation.
+    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+  }
+  \label{fig:negate-mod}
+\end{figure}
+
 \subsection{Pivot-Flips} \label{sec:pivot-flips}
 
-We implemented both modular doubling and modular offset/addition in terms of a non-standard operation we call a ``pivot-flip".
+We implemented both modular negation and modular offset/addition in terms of a non-standard operation we call a ``pivot-flip".
 A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
@@
 However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'', where the operation undoes itself unless a controlling qubit is toggled between two applications.
-Also, because the operation doesn't change whether a given value is below the pivot, the controlling qubit can be toggled by a comparison of the target against the pivot $A$ even though the target is being operated on.
+Also, because the operation doesn't change whether any state is above or below the pivot, the controlling qubit can be toggled by a comparison of the target itself against the pivot (even though the target is being operated on).
 
@@
     Controlled pivot-flip circuit with a constant pivot.
-    Uses $2$ dirty ancilla, $O(c + n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
 Comparison operations toggle a target bit based on the relationship between two input registers.
-We implement comparisons as in \cite{takahashi2005}, using an addition followed by a subtraction that is slightly smaller to clear all changes except an overflow signal into the target.
+We implement comparisons as in \cite{takahashi2005}, using an addition followed by a slightly smaller subtraction that clears all changes except the overflow signal into the target bit.
 For comparisons against another register this uses $1$ dirty ancilla (if controlled, otherwise no ancilla), $O(c + n)$ gates, and $O(c + n)$ depth.
 For comparisons against a constant, the number of gates increases to $O(c + n \lg n)$ and the dirty ancilla is always required.
+
+When dirty qubits are plentiful, the overflow-predicting construction from \cite{haner2016} with better constant factors can be used instead.
 
@@
 We show an equivalent circuit in figure \ref{fig:inlineadder}.
-Note that the construction uses the input register as workspace, so it breaks when the input is a compile-time constant or when the target register is larger than the input register.
+Because the construction uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
 
@@
 
-When the value to be added into a register is a compile-time constant (i.e. when applying an offset gate), we use the construction from \cite{haner2016}.
-Their offset circuit, shown in figure \ref{fig:offset}, uses $O(n \lg n)$ gates, $O(n)$ depth, and a dirty ancilla.
+When the value to be added into a register is a compile-time constant (i.e. when applying an offset gate), we use the offset construction from \cite{haner2016}.
+Their offset circuit, shown in figure \ref{fig:offset}, uses $O(n \lg n)$ gates, $O(n)$ depth, and one dirty ancilla.
 
-When the target register is larger than the input register, three tweaks based on increment/decrement gates can be used (instead of increasing the asymptotic costs).
+When the target register is larger than the input register, three increment/decrement tweaks to the same-register-size adder fix the insufficient-input-space problem without increasing asymptotic costs.
 (Note that, to avoid cyclic dependencies, the increment and decrement constructions described in the next subsection will only use the same-register-size adder.)
@@
 Second, to free the MSB of the input register for use as a dirty ancilla holding the carry signal, we add it into the target ahead of time with a controlled increment.
-Finally, when the MSB is on its role as carry signal causes an extra increment of the target.
+Finally, when the MSB is on (instead of off) its role as carry signal causes an extra increment of the target.
 This extra increment is cancelled with a controlled decrement.
@@
 When the NOT gates don't fire, the subtraction cancels the addition.
-When the NOT gates do fire, the subtraction is inverted into an addition and $2 \cdot a$ is added to the target.
-To add $a$ instead of $2 \cdot a$ into the target, we temporarily prepend a dirty bit onto the target register.
+When the NOT gates do fire, the subtraction is inverted into an addition and $2 \cdot A$ is added to the target.
+To add $A$ instead of $2 \cdot A$ into the target, we temporarily prepend a dirty bit onto the target register.
 See \ref{fig:controlled-addition}.
@@
       Uses $O(n)$ gates and depth.
-      The increment and decrement gates all have at least one free wire to borrow as a dirty ancilla.}
+      The increment and decrement gates all have at least one free wire to borrow as a dirty ancilla.
+  }
   \label{fig:inline-adder-into-large}
@@
 For even-sized registers, we simply handle the LSB separately.
-Also, we support contorls with another conditional-inversion trick.
+Also, we support controls with another conditional-inversion trick.
 See figure \ref{fig:controlled-increment-odd}.
@@
 We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
-[[[ insert link to code ]]] [[[ maybe make a plot of gate counts vs size?
-]]] [[[ definitely need another constant-factor joke ]]]
+The source code, and an issue tracker for submitting errata, can be found at \href{https://github.com/Strilanc/PaperImpl-PeriodFindCutQubits}{github.com/Strilanc/PaperImpl-PeriodFindCutQubits}.
+[[[ maybe make a plot of gate counts vs size? ]]] [[[ definitely need another constant-factor joke ]]]
 
@@
 Pivot flips also require ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
-This improvement saves a qubit, reducing the total number of qubits required for period finding from $2n+2$ to $2n+1$.
+This improvement saves a qubit in the overall circuit, reducing the total number of qubits required for period finding (on modular exponentiation against an $n$-bit modulus) from $2n+2$ to $2n+1$.
 
-Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}, where any garbage in the ancilla register would have infected the target register and trashed it.
-We deal with the trash with an extra multiply-accumulate per modular multiplication, and cancel the overall effects on the ancilla register with a clean-up multiplication informed by measuring the main register at the end of the circuit.
+Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}.
+We use an an extra multiply-accumulate and negation to prevent the ancilla register from leaking into the work register, and add a clean-up multiplication informed by the work register at the end of the circuit to undo any effects on the ancilla register.
 This improvement allows $n-1$ of the qubits in the ancilla register to be dirty (only the MSB must be zero; to ensure the register's value is in range).
 
-Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $O(n)$ ancilla \cite{draper2000, barenco1995}.
+Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancilla \cite{draper2000, barenco1995}.
 Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
 Note that, for classical reversible computation, 1 dirty ancilla is optimal.
-The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
-With quantum operations, we achieve zero ancilla and $O(n)$ gates (see figure \ref{fig:bootstrap-ancilla}).
+The parity of the state permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
+When quantum operations are available, the final ancilla can be cut without increasing the asymptotic cost (see figure \ref{fig:bootstrap-ancilla}).
 
@@
   \includegraphics[height=17.3cm]{assets/dependencies.png}
-  \caption{\em
-    Meta-figure showing simplified dependencies between constructions, from period finding to constant-size gates.
-    Edge labels indicate which constructions use their dependencies more than a constant number of times.
+  \caption{
+    Meta-figure showing the transitive reduction of dependencies between constructions in our paper.
+    Edge labels indicate which constructions use a dependency more than a constant number of times.
   }

### 2017-05-09 src/dirty_period_finding/find_period.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
-from __future__ import unicode_literals
 from __future__ import print_function
+from __future__ import unicode_literals
 
 import math
 import random
@@ -9,7 +9,6 @@ from fractions import Fraction
 from projectq import MainEngine
 from projectq.backends import Simulator
 from projectq.cengines import DecompositionRuleSet
-from projectq.ops import Measure
 from projectq.setups.decompositions import swap2cnot
 from projectq.types import Qureg
 
@@ -22,17 +21,7 @@ from dirty_period_finding.extensions import (
 from dirty_period_finding.gates import *
 
 
-def measure_register(qureg):
-    if not len(qureg):
-        return 0
-    eng = [q.engine for q in qureg][0]
-    Measure | qureg
-    eng.flush()
-    return int(qureg)
-
-
-def sample_period(eng,
-                  base,
+def sample_period(base,
                   modulus,
                   precision,
                   phase_qubit,
@@ -40,7 +29,6 @@ def sample_period(eng,
                   ancilla_qureg):
     """
     Args:
-        eng (projectq.cengines.MainEngine):
         base (int):
         modulus (int):
         precision (int):
@@ -82,10 +70,7 @@ def sample_period(eng,
               frac.limit_denominator(modulus - 1).denominator)
 
     # Fix ancilla register using factor in work register.
-    Measure | work_qureg
-    eng.flush()
-
-    fixup_factor = int(work_qureg)
+    fixup_factor = work_qureg.measure()
     fixup_gate = ModularBimultiplicationGate(fixup_factor, modulus)
     fixup_gate | (ancilla_qureg, work_qureg)
 
@@ -129,8 +114,7 @@ def main():
             X | q
     before = ancilla_qureg.measure()
 
-    p = sample_period(eng,
-                      base=base,
+    p = sample_period(base=base,
                       modulus=modulus,
                       precision=n * 2,
                       phase_qubit=eng.allocate_qubit(),

### 2017-05-12 doc/paper.tex
Rewrote rewrote rewrote conclusion

@@
 
-\section{Conclusion} \label{sec:conclusion}
+\section{Future Work and Conclusion} \label{sec:conclusion}
 
-We presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm by 1, and allowed an additional $n-1$ qubits to be dirty.
+Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancilla used to implement modular multiplication in terms of scaled modular addition.
+All the other constructions, for simpler arithmetic, need two or fewer dirty ancilla.
+Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there's always more than enough dirty ancilla available to implement the simpler arithmetic inline.
+There's so much slack in the simpler arithmetic constructions that improving the number of qubits used by period-finding effectively means improving the very first step, from multiplication to some other operation, and nothing else.
+
+For example, consider a situation where we could cut multiplication operations into two pieces.
+If we knew the factorization $p \cdot q$ of the modulus $R$, with $p$ and $q$ each having size $\approx n/2$, then the Chinese remainder theorem guarantees we could uniquely represent $x \mod R$ as the pair $(x \mod p, x \mod q)$.
+Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, so there would be three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) for a total qubit count of $\approx 1.5n$.
+Despite the large reduction in the number of ancilla available to them, the underlying constructions would still contribute no additional ancilla.
+(Note: because the initial value $x_0 = 1$ is trivial, and the final value is discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)
+Of course, since we use period-finding to compute the factorization of $R$, using the factorization of $R$ to optimize period-finding would be paradoxical.
+
+We presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
+The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.
+We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by months or years (depending on how quickly quantum computers gain error-corrected capacity).
 

### 2017-05-17 doc/paper.tex
Use new test methods in multi_not_test

@@
 The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.
-We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by months or years (depending on how quickly quantum computers gain error-corrected capacity).
+We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by years (depending on how quickly quantum computers gain error-corrected capacity).
 

### 2017-05-17 src/dirty_period_finding/decompositions/multi_not_rules.py

@@ -142,30 +142,36 @@ def cut_not_max_controls_into_toffolis(target, controls, dirty_reg):
         sweep(range(d - 1))
 
 
+# Reduce from many targets to one target.
+decompose_multi_not_into_cnots = DecompositionRule(
+    gate_class=MultiNotGate,
+    gate_decomposer=lambda cmd: do_multi_not_with_one_big_not_and_friends(
+        targets=cmd.qubits[0],
+        controls=cmd.control_qubits))
+
+# Use many dirty bits created by other decompositions to cut all the way down
+# to Toffolis.
+decompose_cnot_with_big_workspace = DecompositionRule(
+    gate_class=XGate,
+    gate_recognizer=min_controls(3) & min_workspace_vs_controls(factor=1,
+                                                                offset=-2),
+    gate_decomposer=lambda cmd: cut_not_max_controls_into_toffolis(
+        target=cmd.qubits[0],
+        controls=cmd.control_qubits,
+        dirty_reg=workspace(cmd)))
+
+# Use a dirty bit to cut control counts in half.
+decompose_halve_cnot_with_single_workspace = DecompositionRule(
+    gate_class=XGate,
+    gate_recognizer=min_controls(3) & min_workspace(1),
+    gate_decomposer=lambda cmd: cut_not_max_controls_in_half(
+        target=cmd.qubits[0],
+        controls=cmd.control_qubits,
+        dirty=workspace(cmd)[0]))
+
+
 all_defined_decomposition_rules = [
-    # Use many dirty bits from last step to cut all the way down to Toffolis.
-    DecompositionRule(
-        gate_class=XGate,
-        gate_recognizer=min_controls(3) &
-                        min_workspace_vs_controls(factor=1, offset=-2),
-        gate_decomposer=lambda cmd: cut_not_max_controls_into_toffolis(
-            target=cmd.qubits[0],
-            controls=cmd.control_qubits,
-            dirty_reg=workspace(cmd))),
-
-    # Reduce from many targets to one target.
-    DecompositionRule(
-        gate_class=MultiNotGate,
-        gate_decomposer=lambda cmd: do_multi_not_with_one_big_not_and_friends(
-            targets=cmd.qubits[0],
-            controls=cmd.control_qubits)),
-
-    # Use a dirty bit to cut control counts in half.
-    DecompositionRule(
-        gate_class=XGate,
-        gate_recognizer=min_controls(3) & min_workspace(1),
[…]

### 2017-05-17 src/dirty_period_finding/gates/_vector_phaser.py

@@ -144,7 +144,7 @@ class XGate(projectq.ops.XGate, BasicGateEx):
         return FancyXGate()**power
 
 
-class FancyXGate(VectorPhaserGate, BasicMathGateEx, projectq.ops.XGate):
+class FancyXGate(VectorPhaserGate, BasicMathGateEx, XGate):
     def __init__(self, half_turns=1.0):
         BasicMathGateEx.__init__(self)
         projectq.ops.XGate.__init__(self)
@@ -154,7 +154,7 @@ class FancyXGate(VectorPhaserGate, BasicMathGateEx, projectq.ops.XGate):
         return 1 if r == 0 else 0,
 
     def with_half_turns(self, half_turns):
-        return XGate(half_turns)
+        return FancyXGate(half_turns)
 
     def base_str(self):
         return 'X'

### 2017-05-17 test/_test_util.py

@@ -384,10 +384,11 @@ def decomposition_to_ascii(gate,
                            decomposition_rule,
                            register_sizes,
                            control_size=0,
-                           workspace=0):
+                           workspace=0,
+                           ascii_only=True):
     return commands_to_ascii_circuit(record_decomposition(gate,
                                                           decomposition_rule,
                                                           register_sizes,
                                                           control_size,
                                                           workspace),
-                                     ascii_only=True)
+                                     ascii_only=ascii_only)

### 2017-05-17 test/decompositions/multi_not_test.py

@@ -2,151 +2,140 @@
 from __future__ import division
 from __future__ import unicode_literals
 
-import random
-
-from projectq import MainEngine
-from projectq.cengines import DecompositionRuleSet, DummyEngine
-
-from dirty_period_finding.decompositions import multi_not_rules
 from dirty_period_finding.decompositions.multi_not_rules import (
-    do_multi_not_with_one_big_not_and_friends,
-    cut_not_max_controls_in_half,
-    cut_not_max_controls_into_toffolis,
+    decompose_multi_not_into_cnots,
+    decompose_halve_cnot_with_single_workspace,
+    decompose_cnot_with_big_workspace,
 )
-from dirty_period_finding.extensions import (
-    LimitedCapabilityEngine,
-    AutoReplacerEx,
-)
-from dirty_period_finding.gates import MultiNot
-from dirty_period_finding.gates import X
+from dirty_period_finding.gates import MultiNot, X
 from .._test_util import (
-    fuzz_permutation_circuit, check_permutation_circuit
+    decomposition_to_ascii,
+    check_permutation_decomposition,
+    cover,
 )
 
 
-def test_do_multi_not_with_one_big_not_and_friends():
-    backend = DummyEngine(save_commands=True)
-    eng = MainEngine(backend=backend, engine_list=[])
-    controls = eng.allocate_qureg(5)
-    targets = eng.allocate_qureg(5)
-    backend.received_commands = []
-
-    do_multi_not_with_one_big_not_and_friends(targets, controls)
-
-    assert backend.received_commands == [
-        (X & targets[3]).generate_command(targets[4]),
-        (X & targets[2]).generate_command(targets[3]),
-        (X & targets[1]).generate_command(targets[2]),
-        (X & targets[0]).generate_command(targets[1]),
-        (X & controls).generate_command(targets[0]),
-        (X & targets[0]).generate_command(targets[1]),
-        (X & targets[1]).generate_command(targets[2]),
-        (X & targets[2]).generate_command(targets[3]),
-        (X & targets[3]).generate_command(targets[4]),
-    ]
+def test_multinot_operation():
+    assert MultiNot.do_operation(5) == (~5,)
 
 
-def test_do_multi_not_with_one_big_not_and_friends_trivial():
-    backend = DummyEngine(save_commands=True)
-    eng = MainEngine(backend=backend, engine_list=[])
-    controls = eng.allocate_qureg(1)
-    targets = eng.allocate_qureg(4)
[…]

### 2017-05-17 src/dirty_period_finding/decompositions/modular_scaled_addition_rules.py
Optimize modular scaled addition with classical doublings

@@ -18,24 +18,22 @@ def do_modular_scaled_addition(gate, input_reg, target_reg, controls):
 
     N: len(input_reg) + len(target_reg)
     C: len(controls)
-    Size: O(N² lg N + NC????????????????????????????????????????)
+    Size: O(N² lg N + NC)
     Depth: O(N² + NC)
     Diagram:
         c              c
-       ━/━━━━●━━━     ━/━━━━━━━━━━━━━━●━━━━━━━━━━━●━━━━━━━━━━━●━━━━━━━━━━━━●━━
-             │                        │           │           │            │
-          ┌──┴──┐                     │           │           │            │
-       ───┤     ├  =  ────────────────┼───────────┼───────────┼────────────●──
-       n-3│     │     n-3             │           │             .`         │
-       ━/━┥     ┝  =  ━/━━━━━━━━━━━━━━┿━━━━━━━━━━━┿━━━━━━━━━━━.━━━━━━━━━━━━┿━━
-          │  A  │                     │           │         .`             │
-       ───┤     ├  =  ────────────────┼───────────●───────────┼────────────┼──
-          │     │                     │           │           │            │
-       ───┤     ├  =  ────────────────●───────────┼───────────┼────────────┼──
-          └──┬──┘                     │           │           │            │
-        n ┌──┴──┐      n ┌────┐⊗n-1┌──┴─┐┌────┐┌──┴─┐┌────┐   │   ┌────┐┌──┴─┐
-       ━/━┥+AK%R┝     ━/━┥÷2%R┝━━━━┥+K%R┝┥×2%R┝┥+K%R┝┥×2%R┝━ ... ━┥×2%R┝┥+K%R┝
-          └─────┘        └────┘    └────┘└────┘└────┘└────┘       └────┘└────┘
+       ━/━━━━●━━━     ━/━━━━●━━━━━●━━━━━━━━●━━━━━━━
+             │              │     │        │
+          ┌──┴──┐           │     │        │
+       ───┤     ├  =  ──────●─────┼────────┼───────
+       n-2│     │     n-2   │   `.         │
+       ━/━┥  A  ┝  =  ━/━━━━┿━━━━━.━━━━━━━━┿━━━━━━━
+          │     │           │      `.      │
+       ───┤     ├  =  ──────┼─────┼────────●───────
+          └──┬──┘           │     │        │
+        n ┌──┴──┐      n ┌──┴─┐   │  ┌─────┴─────┐
[…]

### 2017-05-17 test/decompositions/modular_bimultiplication_test.py

@@ -70,7 +70,7 @@ def test_toffoli_size_of_bimultiplication():
 
     ModularBimultiplicationGate(factor, modulus) & controls | (t1, t2)
 
-    assert 5000 < len(rec.received_commands) < 15000
+    assert 5000 < len(rec.received_commands) < 10000
 
 
 def test_bimultiplication_operation():

### 2017-05-17 test/decompositions/modular_scaled_addition_test.py

@@ -61,7 +61,7 @@ def test_toffoli_size_of_scaled_modular_addition():
 
     ModularScaledAdditionGate(factor, modulus) & controls | (target1, target2)
 
-    assert 30000 < len(rec.received_commands) < 60000
+    assert 20000 < len(rec.received_commands) < 40000
 
 
 def test_scaled_modular_addition_operation():
@@ -92,22 +92,22 @@ def test_diagram_decompose_into_doubled_addition():
         control_size=1)
     print(text_diagram)
     assert text_diagram == """
-|0>-------------------------------------------------@---------------------------@---------------------------@---------------------------@-------
-                                                    |                           |                           |                           |
-|0>-------------------------------------------------|---------------------------|---------------------------|---------------------------@-------
-                                                    |                           |                           |                           |
-|0>-------------------------------------------------|---------------------------|---------------------------@---------------------------|-------
-                                                    |                           |                           |                           |
-|0>-------------------------------------------------|---------------------------@---------------------------|---------------------------|-------
-                                                    |                           |                           |                           |
-|0>-------------------------------------------------@---------------------------|---------------------------|---------------------------|-------
-    .-----------. .-----------. .-----------. .-----|-----. .-----------. .-----|-----. .-----------. .-----|-----. .-----------. .-----|-----.
-|0>-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-
-    |           | |           | |           | |           | |           | |           | |           | |           | |           | |           |
-|0>-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-
-    |           | |           | |           | |           | |           | |           | |           | |           | |           | |           |
-|0>-|  /2 % 13  |-|  /2 % 13  |-|  /2 % 13  |-|  +7 % 13  |-|  *2 % 13  |-|  +7 % 13  |-|  *2 % 13  |-|  +7 % 13  |-|  *2 % 13  |-|  +7 % 13  |-
-    |           | |           | |           | |           | |           | |           | |           | |           | |           | |           |
-|0>-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-|           |-
-    `-----------` `-----------` `-----------` `-----------` `-----------` `-----------` `-----------` `-----------` `-----------` `-----------`
+|0>-------@-------------@-------------@-------------@-------
[…]

### 2017-05-17 src/dirty_period_finding/decompositions/addition_rules.py
Try to get simulator performance up

@@ -3,9 +3,9 @@ from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
 from projectq.meta import Control
+from projectq.ops import X
 
 from dirty_period_finding.extensions import (
-    Swap,
     min_workspace,
     min_controls,
     max_controls,
@@ -14,7 +14,6 @@ from dirty_period_finding.extensions import (
     workspace,
 )
 from dirty_period_finding.gates import (
-    X,
     AdditionGate,
     Increment,
     Decrement,

### 2017-05-17 src/dirty_period_finding/decompositions/bootstrap_ancilla_rules.py

@@ -3,6 +3,7 @@ from __future__ import division
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import XGate, X, Z
 
 from dirty_period_finding.extensions import (
     H,
@@ -15,9 +16,6 @@ from dirty_period_finding.gates import (
     IncrementGate,
     OffsetGate,
     PhaseGradient,
-    X,
-    XGate,
-    Z,
 )
 
 

### 2017-05-17 src/dirty_period_finding/decompositions/comparison_rules.py

@@ -2,6 +2,7 @@
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import X
 
 from dirty_period_finding.extensions import (
     workspace,
@@ -12,7 +13,6 @@ from dirty_period_finding.gates import (
     LessThanConstantGate,
     MultiNot,
     PredictOffsetOverflowGate,
-    X,
     XorOffsetCarrySignalsGate,
     OffsetGate,
 )

### 2017-05-17 src/dirty_period_finding/decompositions/increment_rules.py

@@ -1,6 +1,7 @@
 # -*- coding: utf-8 -*-
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import X
 
 from dirty_period_finding.extensions import (
     min_workspace,
@@ -8,14 +9,12 @@ from dirty_period_finding.extensions import (
     max_controls,
     workspace,
     max_register_sizes,
-    max_workspace,
 )
 from dirty_period_finding.gates import (
     Add,
     Subtract,
     IncrementGate,
     MultiNot,
-    X
 )
 
 

### 2017-05-17 src/dirty_period_finding/decompositions/modular_double_rules.py

@@ -2,9 +2,9 @@
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import X
 
 from dirty_period_finding.gates import (
-    X,
     ModularBimultiplicationGate,
     ModularDoubleGate,
     OffsetGate,

### 2017-05-17 src/dirty_period_finding/decompositions/multi_not_rules.py

@@ -2,6 +2,7 @@
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import XGate, X
 
 from dirty_period_finding.extensions import (
     min_workspace,
@@ -9,7 +10,7 @@ from dirty_period_finding.extensions import (
     min_controls,
     workspace,
 )
-from dirty_period_finding.gates import MultiNotGate, X, XGate
+from dirty_period_finding.gates import MultiNotGate
 
 
 def do_multi_not_with_one_big_not_and_friends(targets, controls):

### 2017-05-17 src/dirty_period_finding/decompositions/phase_gradient_rules.py

@@ -3,9 +3,10 @@ from __future__ import division
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import Z
 
 from dirty_period_finding.extensions import max_controls
-from dirty_period_finding.gates import PhaseGradientGate, Z
+from dirty_period_finding.gates import PhaseGradientGate
 
 
 def do_phase_gradient(target_reg, controls, factor):

### 2017-05-17 src/dirty_period_finding/decompositions/reverse_bits_rules.py

@@ -3,13 +3,14 @@ from __future__ import division
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
+from projectq.ops import X
 
 from dirty_period_finding.extensions import (
     Swap,
     min_controls,
     max_controls,
 )
-from dirty_period_finding.gates import ReverseBitsGate, X
+from dirty_period_finding.gates import ReverseBitsGate
 
 
 def do_naive_bit_reverse(target_reg, controls):

### 2017-05-17 src/dirty_period_finding/extensions/__init__.py

@@ -1,6 +1,7 @@
 from . import _monkey_patching
 from ._basic_gate_ex import (
     BasicGateEx,
+    gate_and,
     H,
     SelfInverseGateEx,
     GateWithCurriedControls,

### 2017-05-17 src/dirty_period_finding/extensions/_basic_gate_ex.py

@@ -6,12 +6,16 @@ from projectq.ops import BasicGate, Command
 from projectq.types import BasicQubit
 
 
+def gate_and(self, controls):
+    if isinstance(controls, BasicQubit):
+        controls = [controls]
+    controls = list(controls)
+    return GateWithCurriedControls(self, controls)
+
+
 class BasicGateEx(BasicGate):
     def __and__(self, controls):
-        if isinstance(controls, BasicQubit):
-            controls = [controls]
-        controls = list(controls)
-        return GateWithCurriedControls(self, controls)
+        return gate_and(self, controls)
 
     def generate_command(self, qubits):
         qubits = self.make_tuple_of_qureg(qubits)

### 2017-05-17 src/dirty_period_finding/extensions/_monkey_patching.py

@@ -6,8 +6,10 @@ from projectq import MainEngine
 from projectq.backends import ResourceCounter
 from projectq.backends import Simulator
 from projectq.cengines import BasicEngine, UnsupportedEngineError
-from projectq.ops import Command, Measure
+from projectq.ops import Command, Measure, FlushGate, XGate, ZGate
 from projectq.types import Qubit, Qureg
+
+from ._basic_gate_ex import gate_and
 
 
 def command_str(cmd):
@@ -166,6 +168,16 @@ def measure_qubit(qubit):
     return bool(qubit)
 
 
+def simulator_receive(self, command_list):
+    if self.is_last_engine:
+        for cmd in command_list:
+            self._handle(cmd)
+    else:
+        for cmd in command_list:
+            self._handle(cmd)
+            self.send([cmd])
+
+
 Command.__str__ = command_str
 Command.__repr__ = command_str
 Qureg.__str__ = qureg_str
@@ -177,3 +189,27 @@ BasicEngine.__del__ = lambda x: None
 MainEngine.__init__ = MainEngine_init
 Qureg.measure = measure_register
 Qubit.measure = measure_qubit
+Simulator.receive = simulator_receive
+XGate.__and__ = gate_and
+ZGate.__and__ = gate_and
+XGate.__eq__ = lambda self, other: other.__class__ is XGate
+XGate.__hash__ = lambda self: hash((XGate,))
+
+original_handle = Simulator._handle
+
+
+_x_list = [[0, 1], [1, 0]]
+
+
+def sim_handle(self, cmd):
+    if cmd.gate.__class__ is XGate:
+        target = [cmd._qubits[0][0].id]
+        controls = [qb.id for qb in cmd.control_qubits]
+        self._simulator.apply_controlled_gate(_x_list, target, controls)
+        if not self._gate_fusion:
+            self._simulator.run()
+    elif cmd.gate.__class__ is FlushGate:
+        self._simulator.run()
+    else:
+        original_handle(self, cmd)
+Simulator._handle = sim_handle

### 2017-05-17 src/dirty_period_finding/find_period.py

@@ -9,6 +9,7 @@ from fractions import Fraction
 from projectq import MainEngine
 from projectq.backends import Simulator
 from projectq.cengines import DecompositionRuleSet
+from projectq.ops import X, Z
 from projectq.setups.decompositions import swap2cnot
 from projectq.types import Qureg
 

### 2017-05-17 src/dirty_period_finding/gates/__init__.py

@@ -18,4 +18,4 @@ from ._phase_gradient import PhaseGradientGate, PhaseGradient
 from ._pivot_flip import PivotFlipGate, ConstPivotFlipGate, PivotFlip
 from ._reverse_bits import ReverseBitsGate, ReverseBits
 from ._rotate_bits import RotateBitsGate, LeftRotateBits, RightRotateBits
-from ._vector_phaser import VectorPhaserGate, X, Z, XGate, ZGate
+from ._vector_phaser import VectorPhaserGate, XPowGate, ZPowGate

### 2017-05-17 src/dirty_period_finding/gates/_vector_phaser.py

@@ -115,12 +115,12 @@ class VectorPhaserGate(BasicGateEx):
         return hash((VectorPhaserGate, self.half_turns, tuple(self.vector)))
 
 
-class ZGate(VectorPhaserGate):
+class ZPowGate(VectorPhaserGate):
     def __init__(self, half_turns=1.0):
         VectorPhaserGate.__init__(self, [0, 1], half_turns)
 
     def with_half_turns(self, half_turns):
-        return ZGate(half_turns)
+        return ZPowGate(half_turns)
 
     def base_str(self):
         return "Z"
@@ -137,28 +137,16 @@ class ZGate(VectorPhaserGate):
         return VectorPhaserGate.__str__(self)
 
 
-# A complicated hierarchy causes extreme overhead in the simulator loop.
-# So only add the very basics.
-class XGate(projectq.ops.XGate, BasicGateEx):
-    def __pow__(self, power):
-        return FancyXGate()**power
-
-
-class FancyXGate(VectorPhaserGate, BasicMathGateEx, XGate):
+class XPowGate(VectorPhaserGate):
     def __init__(self, half_turns=1.0):
-        BasicMathGateEx.__init__(self)
         projectq.ops.XGate.__init__(self)
         VectorPhaserGate.__init__(self, [1, -1], half_turns)
 
-    def do_operation(self, r):
-        return 1 if r == 0 else 0,
-
     def with_half_turns(self, half_turns):
-        return FancyXGate(half_turns)
+        return XPowGate(half_turns)
 
     def base_str(self):
         return 'X'
 
-
-X = XGate()
-Z = ZGate()
+projectq.ops.XGate.__pow__ = lambda self, power: XPowGate(power)
+projectq.ops.ZGate.__pow__ = lambda self, power: ZPowGate(power)

### 2017-05-17 test/_test_util.py

@@ -12,7 +12,7 @@ from projectq import MainEngine
 from projectq.backends import Simulator
 from projectq.cengines import DummyEngine, DecompositionRule
 from projectq.meta import Dagger
-from projectq.ops import H, All, Rz, Measure
+from projectq.ops import H, All, Rz, Measure, X
 
 from dirty_period_finding.extensions import (
     ClassicalSimulator,
@@ -21,7 +21,6 @@ from dirty_period_finding.extensions import (
     CommandEx,
     BasicGateEx,
 )
-from dirty_period_finding.gates import X
 
 
 def check_phase_circuit(register_sizes,

### 2017-05-17 test/decompositions/addition_test.py

@@ -8,6 +8,7 @@ import random
 from projectq import MainEngine
 from projectq.cengines import (DummyEngine,
                                DecompositionRuleSet)
+from projectq.ops import X
 from projectq.setups.decompositions import swap2cnot
 from projectq.types import Qureg
 
@@ -29,8 +30,6 @@ from dirty_period_finding.gates import (
     MultiNot,
     IncrementGate,
     DecrementGate,
-    X,
-    XGate,
 )
 from .._test_util import fuzz_permutation_circuit, check_permutation_circuit
 
@@ -59,7 +58,6 @@ def test_exact_commands_for_small_circuit():
 
 
 def test_toffoli_size_of_addition():
-    x = XGate()
     backend = DummyEngine(save_commands=True)
     eng = MainEngine(backend=backend, engine_list=[
         AutoReplacerEx(DecompositionRuleSet(modules=[

### 2017-05-17 test/decompositions/bootstrap_ancilla_test.py

@@ -3,6 +3,7 @@ from __future__ import division
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRuleSet
+from projectq.ops import X
 from projectq.setups.decompositions import swap2cnot
 
 from dirty_period_finding.decompositions import (
@@ -16,7 +17,7 @@ from dirty_period_finding.extensions import (
     AutoReplacerEx,
     LimitedCapabilityEngine,
 )
-from dirty_period_finding.gates import Increment, X
+from dirty_period_finding.gates import Increment
 from .._test_util import check_quantum_permutation_circuit
 
 

### 2017-05-17 test/decompositions/multi_not_test.py

@@ -2,12 +2,14 @@
 from __future__ import division
 from __future__ import unicode_literals
 
+from projectq.ops import X
+
 from dirty_period_finding.decompositions.multi_not_rules import (
     decompose_multi_not_into_cnots,
     decompose_halve_cnot_with_single_workspace,
     decompose_cnot_with_big_workspace,
 )
-from dirty_period_finding.gates import MultiNot, X
+from dirty_period_finding.gates import MultiNot
 from .._test_util import (
     decomposition_to_ascii,
     check_permutation_decomposition,
@@ -34,7 +36,7 @@ def test_decompose_halve_cnot_with_single_workspace():
         if control_size >= 3:
             check_permutation_decomposition(
                 decomposition_rule=decompose_halve_cnot_with_single_workspace,
-                gate=X**1,
+                gate=X,
                 register_sizes=[1],
                 control_size=control_size,
                 workspace=1)
@@ -45,7 +47,7 @@ def test_decompose_cnot_with_big_workspace():
         if control_size >= 3:
             check_permutation_decomposition(
                 decomposition_rule=decompose_cnot_with_big_workspace,
-                gate=X**1,
+                gate=X,
                 register_sizes=[1],
                 control_size=control_size,
                 workspace=max(0, control_size - 2))

### 2017-05-17 test/extensions/limited_capability_engine_test.py

@@ -1,13 +1,10 @@
 import pytest
 from projectq import MainEngine
 from projectq.cengines import DummyEngine
-from projectq.ops import (
-    CNOT, H, Y, Z, FlushGate, Measure, BasicMathGate
-)
+from projectq.ops import CNOT, H, Y, Z, X, FlushGate, Measure, BasicMathGate
 from projectq.types import Qubit
 
 from dirty_period_finding.extensions import LimitedCapabilityEngine
-from dirty_period_finding.gates import X
 
 
 def test_limited_capability_engine_default():

### 2017-05-17 test/extensions/permutation_simulator_test.py

@@ -1,8 +1,7 @@
 from projectq import MainEngine
-from projectq.ops import BasicMathGate
+from projectq.ops import BasicMathGate, X
 
 from dirty_period_finding.extensions import PermutationSimulator
-from dirty_period_finding.gates import X
 
 
 def test_simulator_triangle_increment():

### 2017-05-18 doc/paper.tex
Addressing Dave's comments

@@
 We present reversible classical circuits for performing various arithmetic operations aided by dirty ancilla in an unknown state that must be preserved.
-We improve the number of qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ qubits \cite{takahashi2006, haner2016} to $2n+1$ qubits, without increasing the asymptotic depth of the circuit.
-We also allow $n-1$ of the $2n+1$ qubits to be dirty.
+We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
 \end{abstract}
@@
 When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla is a useful intermediate goal.
-In this paper, at great cost to the constant factors in the gate count and depth, we reduce the number of qubits required to perform period finding on modular exponentiation from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits without increasing the asymptotic costs.
+In this paper, we reduce the number of qubits required to find the period of exponentiation, modulo an $n$-bit modulus, from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.
+We do so without increasing the asymptotic costs, though the constant factors of our constructions are worse.
 
@@
 Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions will follow.
-In section \ref{sec:construct} we describe the circuit constructions we use to convert period finding into constant-sized gates, while tracking the number of required dirty ancilla.
+In section \ref{sec:construct} we describe the circuit constructions we use to reduce period finding into constant-sized gates, while tracking the number of required dirty ancilla.
 Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
-Finally, section \ref{sec:conclusion} concludes and summarizes.
+Finally, section \ref{sec:conclusion} concludes with a discussion of possible future improvements.
 
@@
 Nearly all presented circuit constructions are classical, and the required dirty qubit counts we note and show in diagrams are in terms of those classical constructions.
-The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla from several of our constructions, but only in places that don't affect the overall total for period-finding.
+(The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla from several of our constructions, but only in places that don't affect the overall total for period-finding.)
 
@@
 For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we nevertheless provide controllable constructions that scale linearly with the number of controls.
-For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a division is a reversed multiplication, and so forth).
+For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a multiply-add is a reversed multiply-subtract, a modular division is a reversed modular multiplication, and so forth).
 
@@
 
+Our reduction from period finding to constant-sized gates uses many circuit constructions.
+Figure \ref{fig:dependencies} shows an overview of the various operations we will use, and the path through them that we will take.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=17.3cm]{assets/dependencies.png}
+  \caption{
+    Meta-figure showing the tagged transitive reduction of dependencies between constructions in our paper.
+    Edge labels indicate which constructions use a dependency more than a constant number of times.
+  }
+  \label{fig:dependencies}
+\end{figure}
+
 \subsection{Period Finding}
-
-A high-level view of the circuit for period finding on modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
-The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $p$ of said operation.
-That is to say, w.l.o.g. the state ends up equal to $|\psi_{1,x}\rangle = \sqrt{2^{-n}/p} \sum_{k=0}^{2^n/p} |pk + x \rangle$ for some $x$.
-The circuit then applies an inverse Fourier transform to the state.
-Fourier transforming a uniform signal with period $p$ produces a spectrum with peaks near $N \cdot 0/p$, $N \cdot 1/p$, $N \cdot 2/p$, ..., $N \cdot (p-1)/p$.
-Shor's algorithm recovers the period $p$ by sampling values $s_i$ from said spectrum, then using a continued fractions algorithm to compute the denominator of a fraction closest to $s_i/N$ given some maximum denominator.
-
-Because period-finding measures all qubits immediately after performing a QFT, most of the transformed qubits can be measured earlier than shown.
-In fact, each qubit can be measured so early that the next qubit needed for the QFT does not even need to be initialized yet!
-Only one of the qubits in the phase-estimation register needs to be present at a time, and so the phase register can be reduced to a single repeatedly-used qubit \cite{beauregard2003}.
-Figure \ref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
-It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
-Note that the only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register.
-
-We perform modular multiplication with multiply-accumulates and an ancilla register as in \cite{beauregard2003}.
-However, to allow our ancilla register to be mostly dirty, we extend the operation so that it has a well defined effect on the second register: multiplying by the inverse factor.
-We will refer to this combined multiply-A-and-inverse-multiply-B operation as a ``bimultiply".
-
-The bimultiplications we perform throughout the algorithm do change the value of the ancilla register, but after the usual end of the circuit we can undo the damage.
-The key insight is that the work register is initialized to $|1\rangle$ and gets multiplied by constants inverse to the ones trashing the dirty ancilla register.
-Instead of discarding the work register after performing phase estimation, we measure it to recover the fixup factor needed to restore the dirty register's original value.
-Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
-
-Because our modular circuit constructions would have undefined behavior if any register has a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
-We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e. clean) and that the registers be as small as possible (i.e. have size $n = \lceil \lg_2(R-1) \rceil$).
 
@@
 \end{figure}
+
+A high-level view of the circuit for period finding on modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
+The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n-1} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $l$ of said operation.
+That is to say, w.l.o.g. the state ends up equal to $|\psi_{1,x}\rangle = \sqrt{l 2^{-n}} \sum_{k=0}^{(2^n/l)-1} |l k + x \rangle$ for some $x$.
+The circuit then applies an inverse Fourier transform to the state.
+Fourier transforming a uniform signal with period $l$ produces a spectrum with peaks near $N \cdot 0/l$, $N \cdot 1/l$, $N \cdot 2/l$, ..., $N \cdot (p-1)/l$.
+Shor's algorithm recovers the period $l$ by sampling values $s_i$ from said spectrum, then using a continued fractions algorithm to compute the denominator of the fraction (with denominator below $R$) that is closest to $s_i/N$.
+
+Because period-finding measures all qubits immediately after performing a QFT, most of the transformed qubits can be measured earlier than shown.
+In fact, each qubit can be measured so early that the next qubit needed for the QFT does not even need to be initialized yet!
+Only one of the qubits in the phase-estimation register needs to be present at a time, and so the phase register can be reduced to a single repeatedly-used qubit \cite{zalka1998, mosca1999, parker2000, beauregard2003}.
+Figure \ref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
+It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
+The only non-trivial operation is (a), the controlled modular multiplication of an $n$-qubit register.
+
+We perform modular multiplication with multiply-accumulates and an ancilla register as in \cite{beauregard2003}.
+However, to allow our ancilla register to be mostly dirty, we extend the operation so that it has a well defined effect on the second register: multiplying by the inverse factor.
+We will refer to this combined multiply-A-and-inverse-multiply-B operation as a ``bimultiply".
+
+The bimultiplications we perform throughout the algorithm do change the value of the ancilla register, but after the usual end of the circuit we can undo the damage.
+The key insight is that the work register is initialized to $|1\rangle$ and gets multiplied by constants inverse to the ones trashing the dirty ancilla register.
+Instead of discarding the work register after performing phase estimation, as is normally done, we have a use for it.
+We measure the work register to recover the fixup factor needed to restore the dirty register's original value.
+Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
+
+Because our modular circuit constructions would have undefined behavior if any register has a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
+We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e. clean) and that the registers be as small as possible (i.e. have size $n = \lceil \lg_2(R-1) \rceil$).
 
@@
 
-To control addition gates, and offset gates, we do an uncontrolled addition followed by an uncontrolled subtraction bordered by NOT gates that fire when the controls are satisfied.
+To control addition gates, and offset gates, we do an uncontrolled addition followed by an uncontrolled subtraction bordered by NOT gates conditioned on all the controls.
 When the NOT gates don't fire, the subtraction cancels the addition.
@@
 To add $A$ instead of $2 \cdot A$ into the target, we temporarily prepend a dirty bit onto the target register.
-See \ref{fig:controlled-addition}.
+See figure \ref{fig:controlled-addition}.
 
@@
   \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
-  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(N)$ depth, size, and $N$ dirty ancilla.}
+  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancilla.}
   \label{fig:increment-many-dirty}
@@
   In this case each state is phased by $e^{i \frac{\pi}{2} v/2^d}$.
-  The phase gradient is implemented by a column of $Z^{2^{-k}}$ gates, .}
+  The phase gradient is implemented by a column of $Z^{2^{-k}}$ gates.}
   \label{fig:bootstrap-ancilla}
@@
 To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others.
-We then use a dirty ancilla (i.e. one of the other targets) to efficiently reduce the big CNOT into Toffolis.
+We then use the available dirty ancilla to reduce the single remaining CNOT with many controls into Toffoli gates.
+See figures \ref{fig:multi-not} and \ref{fig:cnot-reduction}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/multi-cnot-to-single-cnots.png}
+  \caption{\em
+    Separating controlled not gates with many controls and many targets into single-target single-control not-gates and one single-target many-control not-gate.
+    Uses no ancilla, $O(c + n)$ gates, and $O(c + \lg n)$ depth.
+  }
+  \label{fig:multi-not}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/cnot-reduction.png}
+  \caption{\em
+    When $c-2$ dirty ancilla are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates.
+  }
+  \label{fig:cnot-reduction}
+\end{figure}
 
@@
 We perform $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
-
-We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
-The source code, and an issue tracker for submitting errata, can be found at \href{https://github.com/Strilanc/PaperImpl-PeriodFindCutQubits}{github.com/Strilanc/PaperImpl-PeriodFindCutQubits}.
-[[[ maybe make a plot of gate counts vs size? ]]] [[[ definitely need another constant-factor joke ]]]
 
@@
 Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
-Note that, for classical reversible computation, 1 dirty ancilla is optimal.
-The parity of the state permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
+Note that, for classical reversible computation, 1 dirty ancilla is the minimum possible for an incrementer.
+The parity of the state permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even, and odd-parity permutations can't be implemented by composing even-parity permutations.
 When quantum operations are available, the final ancilla can be cut without increasing the asymptotic cost (see figure \ref{fig:bootstrap-ancilla}).
 
-\begin{figure}
-  \centering
-  \includegraphics[height=17.3cm]{assets/dependencies.png}
-  \caption{
-    Meta-figure showing the transitive reduction of dependencies between constructions in our paper.
-    Edge labels indicate which constructions use a dependency more than a constant number of times.
-  }
-  \label{fig:dependencies}
-\end{figure}
+We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
+The source code for the ProjectQ implementation, and an issue tracker for submitting errata, can be found at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
+
+Although the asymptotic costs of our constructions match previous work, the constant factors are significantly worse.
+For example, our equivalent of a controlled 32-qubit multiplication uses roughly 1.3 million Toffoli gates (the specific number depends on the factor, the modulus, whether or not a number of optimizations are used, and whether or not those optimizations apply to a given case).
+This cost is over an order of magnitude worse than previous work \cite{haner2016}.
+The reason for the increased cost is that using dirty ancilla results in more uncomputation work, in repeating some operations twice and conditionally to do them once unconditionally, and in less efficient constructions in general.
+These inefficiencies stack multiplicatively when one construction uses another as a subroutine.
 
@@
 Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancilla used to implement modular multiplication in terms of scaled modular addition.
-All the other constructions, for simpler arithmetic, need two or fewer dirty ancilla.
+All the other constructions, for simpler arithmetic, can be done with two or fewer dirty ancilla.
 Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there's always more than enough dirty ancilla available to implement the simpler arithmetic inline.
-There's so much slack in the simpler arithmetic constructions that improving the number of qubits used by period-finding effectively means improving the very first step, from multiplication to some other operation, and nothing else.
+There's so much slack in the simpler arithmetic constructions that improving the number of qubits used by period-finding effectively means improving the very first step, reducing multiplication into some other operation, and nothing else.
 
@@
 Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, so there would be three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) for a total qubit count of $\approx 1.5n$.
-Despite the large reduction in the number of ancilla available to them, the underlying constructions would still contribute no additional ancilla.
-(Note: because the initial value $x_0 = 1$ is trivial, and the final value is discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)
+Despite the large reduction in the number of ancilla available to the underlying arithmetic operations, they would still have more than enough ancilla and contribute no additional ancilla to the overall circuit.
+(Note: because the initial value $x_0 = 1$ is trivial, and the final value is measured or discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)
 Of course, since we use period-finding to compute the factorization of $R$, using the factorization of $R$ to optimize period-finding would be paradoxical.
@@
 The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.
-We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by years (depending on how quickly quantum computers gain error-corrected capacity).
+We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by years (depending on the efficiency of the construction and how quickly quantum computers will gain error-corrected capacity).
 

### 2017-05-18 doc/paper.tex
Cite Quirk

@@
 
-We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
-The source code for the ProjectQ implementation, and an issue tracker for submitting errata, can be found at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
+We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} \cite{quirk2016} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
+The source code for the ProjectQ implementation, and an issue tracker for submitting errata, can be found online at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
 
@@
 Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, so there would be three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) for a total qubit count of $\approx 1.5n$.
-Despite the large reduction in the number of ancilla available to the underlying arithmetic operations, they would still have more than enough ancilla and contribute no additional ancilla to the overall circuit.
+Despite the large reduction in the number of ancilla available to the underlying arithmetic operations, each operation would still have more than enough ancilla and contribute no additional ancilla to the overall circuit.
 (Note: because the initial value $x_0 = 1$ is trivial, and the final value is measured or discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)
@@
 
-We presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
+In this paper we presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
 The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.

### 2017-05-25 doc/paper.tex
Ancilla plural -> ancillae

@@
 \em
-We present reversible classical circuits for performing various arithmetic operations aided by dirty ancilla in an unknown state that must be preserved.
+We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae in an unknown state that must be preserved.
 We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
@@
 
-When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancilla.
+When constructing quantum circuits, or classical reversible circuits, an important resource is the number of available ancillae.
 An ancilla is an extra bit or qubit that is available for use by a circuit as temporary workspace.
-Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
+Ancillae may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
 Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, since any temporarily unused bit is a borrowable dirty bit.
 
-One part of a circuit can borrow dirty bits from another part of the same circuit, so circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
-When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancilla to dirty ancilla is a useful intermediate goal.
+One part of a circuit can borrow dirty bits from another part of the same circuit, so circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancillae are too far away to be acquired quickly.
+When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancillae to dirty ancillae is a useful intermediate goal.
 In this paper, we reduce the number of qubits required to find the period of exponentiation, modulo an $n$-bit modulus, from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.
@@
 Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions will follow.
-In section \ref{sec:construct} we describe the circuit constructions we use to reduce period finding into constant-sized gates, while tracking the number of required dirty ancilla.
+In section \ref{sec:construct} we describe the circuit constructions we use to reduce period finding into constant-sized gates, while tracking the number of required dirty ancillae.
 Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
@@
 Nearly all presented circuit constructions are classical, and the required dirty qubit counts we note and show in diagrams are in terms of those classical constructions.
-(The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla from several of our constructions, but only in places that don't affect the overall total for period-finding.)
+(The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla(e) from several of our constructions, but only in places that don't affect the overall total for period-finding.)
 
@@
 	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(n)$ controls the precision of the phase estimation step.
-    The triangles indicate how many ancilla are needed ``behind the scenes", by our constructions, to perform an operation.
+    The triangles indicate how many ancillae are needed ``behind the scenes", by our constructions, to perform an operation.
 	Recovering the period requires classical post-processing of the sampled output.
@@
     $K$ must have a multiplicative inverse modulo $R$.
-    Uses no ancilla, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses no ancillae, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Requires the modulus $R$ to be odd.
-    Uses no ancilla, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses no ancillae, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
   \caption{\em
-    Controlled modular doubling in $O(c + n \lg n)$ gates, $O(c + n)$ depth, with 2 dirty ancilla.
+    Controlled modular doubling in $O(c + n \lg n)$ gates, $O(c + n)$ depth, with 2 dirty ancillae.
     $R$ must be odd.
@@
     Controlled modular addition construction based on pivot-flips.
-    Uses $2-c$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2-c$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Controlled modular offset construction based on pivot-flips.
-    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Controlled modular negation.
-    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Controlled pivot-flip circuit with an enregistered pivot.
-    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
     Controlled pivot-flip circuit with a constant pivot.
-    Uses $2$ dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
   }
@@
 We implement comparisons as in \cite{takahashi2005}, using an addition followed by a slightly smaller subtraction that clears all changes except the overflow signal into the target bit.
-For comparisons against another register this uses $1$ dirty ancilla (if controlled, otherwise no ancilla), $O(c + n)$ gates, and $O(c + n)$ depth.
+For comparisons against another register this uses $1$ dirty ancilla (if controlled, otherwise no ancillae), $O(c + n)$ gates, and $O(c + n)$ depth.
 For comparisons against a constant, the number of gates increases to $O(c + n \lg n)$ and the dirty ancilla is always required.
@@
 
-\cite{takahashi2005} provides a reversible adder circuit that uses $O(n)$ gates, $O(n)$ depth, and no ancilla.
+\cite{takahashi2005} provides a reversible adder circuit that uses $O(n)$ gates, $O(n)$ depth, and no ancillae.
 We show an equivalent circuit in figure \ref{fig:inlineadder}.
@@
   \caption{\em Adder for input and target registers of the same size.
-  Requires no ancilla, and uses $O(N)$ gates and depth.
+  Requires no ancillae, and uses $O(N)$ gates and depth.
   Based on \cite{van2004, takahashi2005}.}
@@
   \caption{\em
-      Adder with target larger than source, using no ancilla.
+      Adder with target larger than source, using no ancillae.
       Uses $O(n)$ gates and depth.
@@
   	Uses $O(c + n)$ gates, $O(c + n)$ depth, and a dirty ancilla.
-  	When applied to offset gates, uses $O(c + n \lg n)$ gates, $O(n + c)$ depth, and two dirty ancilla.
+  	When applied to offset gates, uses $O(c + n \lg n)$ gates, $O(n + c)$ depth, and two dirty ancillae.
   }
@@
   \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
-  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancilla.}
+  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancillae.}
   \label{fig:increment-many-dirty}
@@
     A controlled bit rotation / bit swap is three controlled bit-reverses.
-    Uses no ancilla, $O(n+c)$ gates, and $O(n+c)$ depth.
+    Uses no ancillae, $O(n+c)$ gates, and $O(n+c)$ depth.
   }
@@
     For reversals on an even number of bits, swap the center two bits beforehand then use one of them as the center workspace bit from this construction.
-    Uses no ancilla, $O(c + n)$ gates, and $O(c + n)$ depth.
+    Uses no ancillae, $O(c + n)$ gates, and $O(c + n)$ depth.
   }
@@
 To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others.
-We then use the available dirty ancilla to reduce the single remaining CNOT with many controls into Toffoli gates.
+We then use the available dirty ancillae to reduce the single remaining CNOT with many controls into Toffoli gates.
 See figures \ref{fig:multi-not} and \ref{fig:cnot-reduction}.
@@
     Separating controlled not gates with many controls and many targets into single-target single-control not-gates and one single-target many-control not-gate.
-    Uses no ancilla, $O(c + n)$ gates, and $O(c + \lg n)$ depth.
+    Uses no ancillae, $O(c + n)$ gates, and $O(c + \lg n)$ depth.
   }
@@
   \caption{\em
-    When $c-2$ dirty ancilla are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates.
+    When $c-2$ dirty ancillae are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates.
   }
@@
 Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
-Pivot flips also require ancilla (two, actually), but the ancilla can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
+Pivot flips also require ancillae (two, actually), but the ancillae can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
 This improvement saves a qubit in the overall circuit, reducing the total number of qubits required for period finding (on modular exponentiation against an $n$-bit modulus) from $2n+2$ to $2n+1$.
@@
 
-Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancilla \cite{draper2000, barenco1995}.
+Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancillae \cite{draper2000, barenco1995}.
 Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
@@
 This cost is over an order of magnitude worse than previous work \cite{haner2016}.
-The reason for the increased cost is that using dirty ancilla results in more uncomputation work, in repeating some operations twice and conditionally to do them once unconditionally, and in less efficient constructions in general.
+The reason for the increased cost is that using dirty ancillae results in more uncomputation work, in repeating some operations twice and conditionally to do them once unconditionally, and in less efficient constructions in general.
 These inefficiencies stack multiplicatively when one construction uses another as a subroutine.
@@
 
-Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancilla used to implement modular multiplication in terms of scaled modular addition.
-All the other constructions, for simpler arithmetic, can be done with two or fewer dirty ancilla.
-Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there's always more than enough dirty ancilla available to implement the simpler arithmetic inline.
+Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancillae used to implement modular multiplication in terms of scaled modular addition.
+All the other constructions, for simpler arithmetic, can be done with two or fewer dirty ancillae.
+Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there's always more than enough dirty ancillae available to implement the simpler arithmetic inline.
 There's so much slack in the simpler arithmetic constructions that improving the number of qubits used by period-finding effectively means improving the very first step, reducing multiplication into some other operation, and nothing else.
@@
 Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, so there would be three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) for a total qubit count of $\approx 1.5n$.
-Despite the large reduction in the number of ancilla available to the underlying arithmetic operations, each operation would still have more than enough ancilla and contribute no additional ancilla to the overall circuit.
+Despite the large reduction in the number of ancillae available to the underlying arithmetic operations, each operation would still have more than enough ancillae and contribute no additional ancillae to the overall circuit.
 (Note: because the initial value $x_0 = 1$ is trivial, and the final value is measured or discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)

### 2017-05-25 src/dirty_period_finding/find_period.py
Gate counting

@@ -2,13 +2,14 @@
 from __future__ import print_function
 from __future__ import unicode_literals
 
+import fractions
 import math
 import random
 from fractions import Fraction
 
 from projectq import MainEngine
-from projectq.backends import Simulator
-from projectq.cengines import DecompositionRuleSet
+from projectq.backends import Simulator, ResourceCounter
+from projectq.cengines import DecompositionRuleSet, DummyEngine
 from projectq.ops import X, Z
 from projectq.setups.decompositions import swap2cnot
 from projectq.types import Qureg
@@ -79,8 +80,9 @@ def sample_period(base,
     return frac.limit_denominator(modulus - 1).denominator
 
 
-def main():
+def main_find_period():
     sim = Simulator()
+    cnt = ResourceCounter()
     eng = MainEngine(backend=sim, engine_list=[
         AutoReplacerEx(DecompositionRuleSet(modules=[
             addition_rules,
@@ -103,6 +105,7 @@ def main():
             allow_single_qubit_gates=True,
             allow_classes=[]
         ),
+        cnt,
     ])
 
     modulus = 11 * 5
@@ -129,6 +132,58 @@ def main():
 
     assert before == after
     print(u == 1 or u == -1 % modulus)
+    print(cnt)
+    print("Total gates: {}".format(sum(e for e in cnt.gate_counts.values())))
+
+
+def main_count_gates():
+    a = AutoReplacerEx(DecompositionRuleSet(modules=[
+        addition_rules,
+        increment_rules,
+        modular_addition_rules,
+        modular_bimultiplication_rules,
+        modular_double_rules,
+        modular_negate_rules,
+        modular_scaled_addition_rules,
+        multi_not_rules,
+        offset_rules,
+        pivot_flip_rules,
+        reverse_bits_rules,
+        rotate_bits_rules,
+        swap2cnot,
+        comparison_rules,
+    ]))
+
+    for i in range(3, 10):
+        m = 1 << i
+        m_max = 1 << m
+        x = 2
+        modulus = 2
+        while modulus % 2 == 0 or fractions.gcd(x, modulus) != 1:
+            modulus = random.randint(m_max // 2 + 1, m_max - 1)
+            x = random.randint(1, modulus - 1)
+        cnt = ResourceCounter()
+        eng = MainEngine(backend=DummyEngine(), engine_list=[
+            a,
+            LimitedCapabilityEngine(
+                allow_toffoli=True,
[…]

### 2017-05-25 doc/paper.tex
Addressing comments from maffoo - Added vertical wire-span indicators to diagrams - Clarified several other parts of diagrams - Moved pivot-flip section earlier - Significantly expanded explanations of conventions in the intro - Various other

@@
 
-All constructions in this paper use a 2s-complement representation of integers, and all diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
-Nearly all presented circuit constructions are classical, and the required dirty qubit counts we note and show in diagrams are in terms of those classical constructions.
-(The quantum trick shown in figure \ref{fig:bootstrap-ancilla} can cut dirty ancilla(e) from several of our constructions, but only in places that don't affect the overall total for period-finding.)
+All constructions in this paper use a two's-complement representation of integers, and all diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
 
-For clarity, circuit diagrams will divide operations into separate input and output parts, with inputs shaded gray, when applicable.
-For completeness, even when the period finding reduction doesn't produce a controlled version of an operation, we nevertheless provide controllable constructions that scale linearly with the number of controls.
-For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a multiply-add is a reversed multiply-subtract, a modular division is a reversed modular multiplication, and so forth).
+All our circuit diagrams annotate operations with the number of clean and/or dirty ancillae they need.
+For example, a dashed line from an operation down to a triangle inscribed with "3 dirty" means that the operation needs 3 dirty ancilla.
+If there are 3 unused wires that the operation can borrow, then the triangle will be green and a note of "(satisfied)" will be written underneath.
+If there are not enough unused wires shown in the diagram, the triangle will be yellow.
 
-We do not attempt to give reasonable behavior to modular arithmetic circuits applied to out of range values (ones equal to or larger than the modulus).
+The ancillae counts shown in diagrams and discussed in the text are not optimal; they are entirely based on what the constructions we discuss in this paper achieve.
+Nearly all the circuit constructions we present are classical reversible circuits (i.e.
+they do not use any quantum operations), so tricks such as the bootstrapping shown in figure \ref{fig:bootstrap-ancilla} could cut the ancilla count of several of the constructions.
+However, none of these improvements decrease the ancillae required at a crucial bottleneck, and so they do not reduce the number of qubits required by our construction of Shor's algorithm.
+
+To avoid ambiguity, circuit diagrams will divide multi-register operations into separate parts.
+For example, for an addition operation that adds a register $x$ into a register $y$ (i.e.
+performs $|x\rangle |y\rangle \rightarrow |x\rangle |y + x \pmod{2^n} \rangle$), we will place a light-gray box with the text ``Input A" over the wires corresponding to $x$ and a white box with the text ``+A" over the wires corresponding to $y$.
+The ``A" symbol refers to the value of the input register, in the computational basis, at the time of the operation.
+Diagrams will often use a sequence of operations each with their own ``Input A" specific to that operation.
+This re-use of ``A" does not indicate any relation between those inputs.
+The use of ``A" is merely a convention for indicating how the input-indicating box and the effect-indicating box are supposed to combine to form a single operation.
+
+In the case of operations parametrized by compile-time constants, diagrams will use the letters "K" and/or "R".
+For example, a box showing "+K (mod R)" refers to the modular-offset operation $|x\rangle \rightarrow |x+K \pmod{R}\rangle$.
+We always use $R$ t to refer to the modulus that is factored by Shor's algorithm.
+In several constructions we assume that $R$ is odd.
+Removing factors of 2 from a factoring problem is trivial, so this assumption does not decrease the generality of the construction.
+
+Because adding a constant into a register is more expensive than adding one register into another, we avoid ambiguity between the two by always referring to an operation that adds a constant as an "offset operation".
+We will refer to the operation that adds one register into another as just ``addition" or, when needed for clarity, ``enregistered addition".
+
+For completeness, even when the period finding reduction doesn't require a controlled version of an operation, we nevertheless provide a controllable construction that scale linearly with the number of controls.
+For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a multiply-add is a reversed multiply-subtract, a modular division is a reversed modular multiplication, a modular halving is a reversed modular double, and so forth).
+
+We do not attempt to define or implement reasonable behavior for modular arithmetic circuits applied to out of range values (ones equal to or larger than the modulus).
 This {\em includes input values}.
@@
 
-\section{Period Finding into Toffolis} \label{sec:construct}
+\section{Constructions} \label{sec:construct}
 
-Our reduction from period finding to constant-sized gates uses many circuit constructions.
+Our reduction, from the period finding step of Shor's algorithm down to constant-sized gates, uses many circuit constructions.
 Figure \ref{fig:dependencies} shows an overview of the various operations we will use, and the path through them that we will take.
@@
   \caption{
-    Meta-figure showing the tagged transitive reduction of dependencies between constructions in our paper.
+    Graph of the transitive reduction of dependencies between constructions in our paper.
     Edge labels indicate which constructions use a dependency more than a constant number of times.
@@
   \caption{\em
-	High-level period finding circuit \cite{Shor1999}.
-	$R$ is the modulus being queried, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(n)$ controls the precision of the phase estimation step.
+	High-level definition of the period finding circuit \cite{Shor1999} used by Shor's algorithm.
+	$R$ is the modulus and the number to be factored, $B$ is a randomly chosen base, $n$ is the number of bits needed to store $R$, and $p \in \Theta(n)$ controls the precision of the phase estimation step.
     The triangles indicate how many ancillae are needed ``behind the scenes", by our constructions, to perform an operation.
@@
 
-A high-level view of the circuit for period finding on modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
-The circuit starts by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n-1} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $l$ of said operation.
-That is to say, w.l.o.g. the state ends up equal to $|\psi_{1,x}\rangle = \sqrt{l 2^{-n}} \sum_{k=0}^{(2^n/l)-1} |l k + x \rangle$ for some $x$.
+A high-level view of the circuit for period finding applied to modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
+
+Before the quantum circuit is constructed, a base $B$ is chosen at random.
+The randomly chosen $B$ must be co-prime to $R$.
+When a $B$ is accidentally chosen that is not co-prime to $R$, the lucky victim can instead factor $R$ by recursively factoring $R_1 = \gcd(B, R)$ and $R_2 = R / \gcd(B, R)$.
+
+The circuit begins by preparing a uniform superposition $|\psi_0\rangle = \sqrt{2^{-n}} \sum_{k=0}^{2^n-1} |k\rangle$, then uses the $\times B^A {\pmod R}$ operation to separate that superposition into equivalence classes modulo the unknown period $l$ of this operation.
+That is to say, w.l.o.g. the state ends up equal to $|\psi_{1,x}\rangle = \sqrt{l 2^{-n}} \sum_{k=0}^{\approx (2^n/l)-1} |l k + x \rangle$ for some $x$.
 The circuit then applies an inverse Fourier transform to the state.
 Fourier transforming a uniform signal with period $l$ produces a spectrum with peaks near $N \cdot 0/l$, $N \cdot 1/l$, $N \cdot 2/l$, ..., $N \cdot (p-1)/l$.
-Shor's algorithm recovers the period $l$ by sampling values $s_i$ from said spectrum, then using a continued fractions algorithm to compute the denominator of the fraction (with denominator below $R$) that is closest to $s_i/N$.
+Shor's algorithm recovers the period $l$ by sampling values $s_i$ from this spectrum, then using a continued fractions algorithm to compute the denominator of the fraction (with denominator below $R$) that is closest to $s_i/N$.
 
@@
 However, to allow our ancilla register to be mostly dirty, we extend the operation so that it has a well defined effect on the second register: multiplying by the inverse factor.
-We will refer to this combined multiply-A-and-inverse-multiply-B operation as a ``bimultiply".
+We will refer to this combined operation as a ``bimultiply".
 
@@
 Because our modular circuit constructions would have undefined behavior if any register has a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
-We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e. clean) and that the registers be as small as possible (i.e. have size $n = \lceil \lg_2(R-1) \rceil$).
+We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e.
+clean) and that the registers be as small as possible (i.e.
+have size $n = \lceil \lg_2(R-1) \rceil$, because only values up to $R-1$ need to be stored).
 
@@
 
-To show that the circuit works, suppose that the two registers start in the state $(a, b)$, working modulo $R$.
-By scale-adding the first register times $K$ into the second register, the system advances to $(a, b+aK)$.
-Then a scale-subtract times $K^{-1}$ out of the first register puts the system into the state $(a-bK^{-1}-aKK^{-1}, b+aK)$, which is just $(-bK^{-1}, b+aK)$.
-Next, we cancel the $b$ term in the second register by scale-adding the first register times $K$ into it again, leaving $(-bK^{-1}, b+aK-bK^{-1} \cdot K)$ which is simply $(-bK^{-1}, aK)$.
-Finally, we swap the terms and negate the second term to get $(aK, bK^{-1})$ as desired.
+To show that the circuit works, suppose that the two registers start in the state $(x, y)$, working modulo $R$.
+We scale-add the first register times $K$ into the second register to produce the state $(x, y+xK)$.
+Then a scale-subtract times $K^{-1}$ out of the first register puts the system into the state $(x-yK^{-1}-xKK^{-1}, y+xK)$, which is just $(-yK^{-1}, y+xK)$.
+Next, we cancel the $y$ term in the second register by scale-adding the first register times $K$ into it again, leaving $(-yK^{-1}, y+xK-yK^{-1} \cdot K)$ which is simply $(-yK^{-1}, xK)$.
+Finally, we swap the terms and negate the second term to get $(xK, yK^{-1})$ as desired.
 
@@
     Uses no ancillae, $O(c + n^2 \lg n)$ gates, and $O(c + n^2)$ depth where $c$ is the number of controls and $n$ is the register size.
+    The negation operation requires 2 dirty ancillae, but notice there are free wires in the other register that can be borrowed, so the overall construction requires no additional dirty ancilla (for non-trivial $n$).
   }
@@
 
-Note that the modular doublings require $R$ to be odd, since otherwise the operation would be irreversible.
+Note that the modular doublings (and halvings) require $R$ to be odd, since otherwise the operation would be irreversible.
 However, given that in the intended use case $R$ is a number to be factored, it is reasonable to require callers to have factored out multiples of two beforehand.
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
-  \caption{\em
+  \caption{\emph{•}
     Reducing a controlled modular scale-accumulate, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
@@
 
-To multiply a register by 2 modulo an odd $R$, we use controlled offsets and nots to align each half of the input state space against MSB boundaries.
-Once that's done, a left-rotate completes the operation by turning the MSB into the LSB (converting the alignment into interleaving).
-See figure \ref{fig:modular-double}.
+To multiply a register by 2 modulo an odd $R$, we note that the permutation this operation performs is to take the values from $[(R+1)/2, R)$, the top half, and interleave them between the values from $[0, (R+1)/2)$, the bottom half.
+
+We can perform this interleaving by first offset-ing the top half up to $[2^{n-1}, 2^{n-1}+(R-1)/2)$, i.e. aligning it with the MSB boundary, and then performing a left-rotation of the register bits.
+This moves the MSB to the LSB, converting the alignment into interleaving.
+
+Because it's expensive to perform an offset that affects values above $R/2$ while leaving values below $R/2$ alone, the circuit shown in figure \ref{fig:modular-double} uses a different alignment strategy.
+It starts by offset-ing the whole range down by $\lceil R/2 \rceil$, which wraps the bottom half of the input around to the top of the register's range.
+This offset also aligns what was originally the top half against the bottom of the register range.
+Then it uses a controlled offset to move just the top half of the register up by $\lceil R/2 \rceil$, re-wrapping what was originally the bottom-half past the top and around to the middle, aligning with the MSB boundary.
+The two offsets aligned the top half against 0 and the bottom half against $2^{n-1}$, but we want the reverse, so the problem is fixed by toggling the MSB.
+With the desired alignment achieved, the circuit finishes with a left-rotate.
 
@@
   \caption{\em
-    Controlled modular doubling in $O(c + n \lg n)$ gates, $O(c + n)$ depth, with 2 dirty ancillae.
+    Controlled modular doubling.
     $R$ must be odd.
-    Uses 1 dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    Uses 1 dirty ancilla, $O(c + n \lg n)$ gates, and $O(c + n)$ depth, where $c$ is the number of controls and $n$ is the register size.
   }
@@
 
+\subsection{Pivot-Flips} \label{sec:pivot-flips}
+
+We will implement both modular negation and modular offset/addition in terms of a non-standard operation we call a ``pivot-flip".
+A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
+For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
+
+The exact permutation performed by a pivot-flip with pivot equal to $A$ is:
+
+$$\text{PivotFlip}_A = \sum_{i=0}^{A-1} |A-i-1\rangle \langle i| + \sum_{i=A}^{N-1} |i\rangle \langle i|$$
+
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
+However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'', where the operation undoes itself unless a controlling qubit is toggled between two applications.
+Also, because the operation doesn't change whether any state is above or below the pivot, the controlling qubit can be toggled by a comparison of the target itself against the pivot (even though the target is being operated on).
+
+See figures \ref{fig:controlled-pivot-flip} and \ref{fig:controlled-const-pivot-flip} for the circuit diagrams.
+For states that are less than the pivot $A$, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled subtract-and-inverts will fire.
+For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.
+
+Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
+This distinction matters because the constant-offset construction uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-pivot-flip.png}
+  \caption{\em
+    Controlled pivot-flip circuit with an enregistered pivot.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+  }
+  \label{fig:controlled-pivot-flip}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/controlled-const-pivot-flip.png}
+  \caption{\em
+    Controlled pivot-flip circuit with a constant pivot.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+  }
+  \label{fig:controlled-const-pivot-flip}
+\end{figure}
+
 \subsection{Modular Addition / Offset}
 
-As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot-flips (see section \ref{sec:pivot-flips}).
+As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition can be implemented by three pivot-flips.
 To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
@@
   \label{fig:negate-mod}
-\end{figure}
-
-\subsection{Pivot-Flips} \label{sec:pivot-flips}
-
-We implemented both modular negation and modular offset/addition in terms of a non-standard operation we call a ``pivot-flip".
-A pivot-flip is an operation that reverses the order of states less than a given pivot value, without affecting other states.
-For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
-
-The exact permutation performed by a pivot-flip with pivot equal to $A$ is:
-
-$$\text{PivotFlip}_A = \sum_{i=0}^{A-1} |A-i-1\rangle \langle i| + \sum_{i=A}^{N-1} |i\rangle \langle i|$$
-
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
-However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'', where the operation undoes itself unless a controlling qubit is toggled between two applications.
-Also, because the operation doesn't change whether any state is above or below the pivot, the controlling qubit can be toggled by a comparison of the target itself against the pivot (even though the target is being operated on).
-
-See figures \ref{fig:controlled-pivot-flip} and \ref{fig:controlled-const-pivot-flip} for the circuit diagrams.
-For states that are less than the pivot $A$, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled subtract-and-inverts will fire.
-For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.
-
-Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
-This distinction matters because the constant-offset construction uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-pivot-flip.png}
-  \caption{\em
-    Controlled pivot-flip circuit with an enregistered pivot.
-    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
-  }
-  \label{fig:controlled-pivot-flip}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{assets/controlled-const-pivot-flip.png}
-  \caption{\em
-    Controlled pivot-flip circuit with a constant pivot.
-    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
-  }
-  \label{fig:controlled-const-pivot-flip}
 \end{figure}

### 2017-05-25 doc/paper.tex
Addressing more of maffoo's comments: - Cleaning up prose in several places - Added comparison diagram back in - Noting technique - Etc

@@
 
-All constructions in this paper use a two's-complement representation of integers, and all diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
+All constructions in this paper use a two's-complement representation of integers.
+When a result would exceed the size of a register, it wraps (i.e.
+all the non-modular arithmetic constructions we discus actually perform arithmetic modulo $2^n$, where $n$ is the size of the target register).
+All diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
 
@@
 Because adding a constant into a register is more expensive than adding one register into another, we avoid ambiguity between the two by always referring to an operation that adds a constant as an "offset operation".
-We will refer to the operation that adds one register into another as just ``addition" or, when needed for clarity, ``enregistered addition".
+We will refer to the operation that adds one register into another as just ``addition" or, if needed for clarity, ``enregistered addition".
+
+When an operation is controlled by a wire bundle, it is conditioned on all wires in the bundle.
+When a single-qubit gate is applied to a wire bundle, the gate applies separately to every wire.
+For example, when a not gate controlled by one wire bundle is applied to another wire bundle, every wire in the latter is toggled if and only if every wire in the former is on.
 
@@
   \includegraphics[height=17.3cm]{assets/dependencies.png}
-  \caption{
+  \caption{\em
     Graph of the transitive reduction of dependencies between constructions in our paper.
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
-  \caption{\emph{•}
+  \caption{\em
     Reducing a controlled modular scale-accumulate, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
@@
 
-We can perform this interleaving by first offset-ing the top half up to $[2^{n-1}, 2^{n-1}+(R-1)/2)$, i.e. aligning it with the MSB boundary, and then performing a left-rotation of the register bits.
-This moves the MSB to the LSB, converting the alignment into interleaving.
+We can perform this interleaving by moving the top half up until it starts at $2^{n-1}$, i.e. aligns with the MSB boundary.
+A left-rotate of the register bits then moves the MSB to the LSB, converting the alignment into interleaving.
 
@@
 This offset also aligns what was originally the top half against the bottom of the register range.
-Then it uses a controlled offset to move just the top half of the register up by $\lceil R/2 \rceil$, re-wrapping what was originally the bottom-half past the top and around to the middle, aligning with the MSB boundary.
-The two offsets aligned the top half against 0 and the bottom half against $2^{n-1}$, but we want the reverse, so the problem is fixed by toggling the MSB.
-With the desired alignment achieved, the circuit finishes with a left-rotate.
+Then the circuit uses a controlled offset to move just the top half of the register up by $\lceil R/2 \rceil$, re-wrapping what was originally the bottom-half past the top and around to the middle, aligning with the MSB boundary.
+The circuit then toggles the MSB, fixing the fact that, although the top half is aligned with 0 and the bottom half is aligned with $2^{n-1}$, that's exactly the reverse of what we want.
+With the desired alignment achieved, the circuit interleaves the two halves with a left-rotate.
 
@@
 
-The exact permutation performed by a pivot-flip with pivot equal to $A$ is:
+The exact permutation performed by a pivot-flip with pivot equal to $K$ is:
 
-$$\text{PivotFlip}_A = \sum_{i=0}^{A-1} |A-i-1\rangle \langle i| + \sum_{i=A}^{N-1} |i\rangle \langle i|$$
+$$\text{PivotFlip}_K = \sum_{i=0}^{K-1} |K-i-1\rangle \langle i| + \sum_{i=K}^{N-1} |i\rangle \langle i|$$
 
-To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - A)$ nearly does what is required: it flips the range below $A$ but unfortunately also flips the range above-and-including $A$.
-However, because $x \rightarrow \lnot(x - A)$ is its own inverse, the operation can be ``toggle-controlled'', where the operation undoes itself unless a controlling qubit is toggled between two applications.
-Also, because the operation doesn't change whether any state is above or below the pivot, the controlling qubit can be toggled by a comparison of the target itself against the pivot (even though the target is being operated on).
+To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - K)$ nearly does what is required: it flips the range below $K$ but unfortunately also flips the range above-and-including $K$.
+It's a ``bi-flip".
+
+The bi-flip operation is its own inverse, so it can be ``toggle-controlled''.
+If we apply two bi-flips controlled by the same control qubit, then toggling the control qubit determines whether or not a bi-flip is applied to the target register.
+When the control qubit stays off, neither bi-flip fires and nothing happens to the target.
+When the control qubit stays on, both bi-flips fire and they undo each other.
+But if the control qubit is toggled after the first bi-flip but before the second, then exactly one of them will fire and the target register ends up bi-flipped.
+
+Another useful property of a bi-flip is that it doesn't move any states across the pivot.
+A bi-flip at $K$ preserves $x<K$ for all $x$.
+This allows the toggling of the control qubit, that determines whether a toggled-controlled bi-flip will happen, to be based on a comparison of the pivot against the target register we are operating on.
+Even at points where the target may or may not have been bi-flipped.
 
 See figures \ref{fig:controlled-pivot-flip} and \ref{fig:controlled-const-pivot-flip} for the circuit diagrams.
-For states that are less than the pivot $A$, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled subtract-and-inverts will fire.
-For other states, the ancilla does not change and so the subtract-and-inverse happens either no times or two times (undoing itself) for a net effect of no effect.
-
-Note that, depending on whether or not the pivot is a compile-time constant or another qubit register, different subtraction constructions are used when reducing further.
-This distinction matters because the constant-offset construction uses $O(n \lg n)$ gates \cite{haner2016} instead of the $\Theta(n)$ gates used by enregistered addition \cite{takahashi2005}.
+For states that are less than the pivot, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled bi-flips will fire.
+For states equal to or larger than the pivot, the ancilla does not get toggled and so the bi-flips undo each other.
 
@@
     Controlled pivot-flip circuit with an enregistered pivot.
-    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+    The target register can be larger than the input register, but not smaller.
+    Uses $2$ dirty ancillae, $O(c + m \lg m)$ gates, and $O(c + m)$ depth where $c$ is the number of controls, $n$ is the size of the input register, $e$ is the extra size of the target register, and $m=n+e$ is the size of the target register.
   }
@@
 As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition can be implemented by three pivot-flips.
-To add $A$ into a register modulo $R$, perform pivot-flips with the pivot at $R-A$, then $R$, then $A$.
+To add $K$ into a register modulo $R$, perform pivot-flips with the pivot at $R-K$, then $R$, then $K$.
 (The pivot-flip at $R$ can be simplified because the values above $R$ don't matter.)
@@
     Controlled modular addition construction based on pivot-flips.
+    The arithmetic being performed on the input register temporarily transitions it from storing $x$ to $R-x$ for the first pivot flip.
+    Because $R$ is a compile-time constant, $R+1$ is also a compile-time constant and the $+R+1$ operation is a normal offset operation.
+    All arithmetic is two's complement, i.e. modulo $2^n$.
     Uses $2-c$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
@@
 We implement comparisons as in \cite{takahashi2005}, using an addition followed by a slightly smaller subtraction that clears all changes except the overflow signal into the target bit.
-For comparisons against another register this uses $1$ dirty ancilla (if controlled, otherwise no ancillae), $O(c + n)$ gates, and $O(c + n)$ depth.
-For comparisons against a constant, the number of gates increases to $O(c + n \lg n)$ and the dirty ancilla is always required.
+See figure \ref{fig:compare}.
 
-When dirty qubits are plentiful, the overflow-predicting construction from \cite{haner2016} with better constant factors can be used instead.
+For comparisons against another register this approach uses $1$ dirty ancilla (if controlled, otherwise no ancillae), $O(c + n)$ gates, and $O(c + n)$ depth.
+For comparisons against a constant, the number of gates increases to $O(c + n \lg n)$ and an extra dirty ancilla is required.
+
+When $n$ dirty qubits are available, the overflow-predicting construction from \cite{haner2016} with better constant factors can be used instead.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{assets/comparison-less-const.png}
+  \caption{\em
+    Controlled comparison.
+    Uses $2$ dirty ancillae, $O(c + n \lg n)$ gates, and $O(c + n)$ depth where $c$ is the number of controls and $n$ is the register size.
+  }
+  \label{fig:negate-mod}
+\end{figure}
 
@@
 We show an equivalent circuit in figure \ref{fig:inlineadder}.
+
 Because the construction uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
+(When the target register is smaller, we simply ignore the high bits of the input due to overflow wraparound.)
+This is a problem for us, because some of our constructions add constants into register (e.g. modular scale-add), and some others add inputs into larger target registers (e.g. comparison).
 
@@
 
-When the target register is larger than the input register, three increment/decrement tweaks to the same-register-size adder fix the insufficient-input-space problem without increasing asymptotic costs.
+When the target register is larger than the input register, the addition circuit in fig \ref{fig:inlineadder} can be modified without increasing the asymptotic cost.
+We do this by removing the surround CNOTs and adding three increment/decrement operations to the circuit, as follows.
 (Note that, to avoid cyclic dependencies, the increment and decrement constructions described in the next subsection will only use the same-register-size adder.)
-First, the carry signal is forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
-Second, to free the MSB of the input register for use as a dirty ancilla holding the carry signal, we add it into the target ahead of time with a controlled increment.
-Finally, when the MSB is on (instead of off) its role as carry signal causes an extra increment of the target.
-This extra increment is cancelled with a controlled decrement.
+
+First, because there are many target bits not reached by the time we have swept through the input, carry signals can't reach the high bit of the target register.
+We fix this by replacing innermost CNOT with a controlled increment.
+Second and third, because we're using the MSB of the input register to hold the carry signal, it causes an increment at the LSB instead of at the correct position.
+We undo the LSB increment with a controlled decrement, and use a controlled increment to apply the MSB's effect where it should have actually gone.
 See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
 
-To control addition gates, and offset gates, we do an uncontrolled addition followed by an uncontrolled subtraction bordered by NOT gates conditioned on all the controls.
-When the NOT gates don't fire, the subtraction cancels the addition.
-When the NOT gates do fire, the subtraction is inverted into an addition and $2 \cdot A$ is added to the target.
+To control addition gates, and offset gates, we use ``reverse-controlling".
+Reverse-controlling controls an operation by instead controlling simpler framing operations that, if preserve, reverse the effect of the framed operation.
+In particular, note that framing the target of a subtraction with NOT gates (i.e. applying a NOT gate to every target wire before and after the addition) reverses the subtraction into an addition.
+
+By following the framed-subtraction with an addition, we have created a controlled operation that almost does what is needed.
+When the controls are not satisfied, the NOT gates bordering the subtraction do not fire, the subtraction and addition cancel each other, and nothing happens.
+When the controls are satisfied, the NOT gates reverse the subtraction into an addition, that addition and the second addition both apply, and the input is added into the target twice.
+
 To add $A$ instead of $2 \cdot A$ into the target, we temporarily prepend a dirty bit onto the target register.
@@
 
-To improve from $n$ dirty bits to a single dirty bit, we break the increment operation into two halves.
-A high half that is incremented only if all of the bottom bits are on, and a low half that is unconditionally incremented.
-The low half can be incremented with the double-subtraction trick by borrowing the high half.
-But the double-subtraction trick doesn't work on the high half, because the low half can't be borrowed when it is being used as a control.
+To improve from $n$ dirty bits to the single dirty bit used by the circuit in figure \ref{fig:controlled-increment-odd}, we break the register into two halves.
+A high-half that is incremented only if all of the low bits are on, and a low-half that is unconditionally incremented.
 
-To work around not being able to operate on the borrowed bits while using them as a control, start with the double-subtraction trick but change the second subtraction to an addition and conditionally toggle the target bits instead of the input bits.
-When the condition isn't satisfied, the addition and subtraction will cancel each other.
-When the condition is satisfied, the addition will be inverted into a subtraction, and both the subtraction and addition-turned-subtraction will fire, subtracting the input register from the target register twice.
-However, because we are conditioning on the input bits all being on, we know the input must be -1.
-Therefore the target was incremented by 2.
-To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register.
+If the increment is uncontrolled, the low-half can be incremented with the double-subtraction trick by borrowing the high half.
+If there are controls, we combine the double-subtraction trick with the reverse-control technique we used to control addition.
+(Interestingly, combining the two fixes the double-effect problem that would have otherwise required us to prepend a dirty LSB onto the low-half.)
 
-Note that, to avoid a cyclic dependency, we must require that the additions and subtractions use input and target registers of the same size.
-So the construction described in the previous paragraph only works for odd-sized registers.
-For even-sized registers, we simply handle the LSB separately.
-Also, we support controls with another conditional-inversion trick.
-See figure \ref{fig:controlled-increment-odd}.
+The high-half is trickier to deal with; we want to borrow the low-half for the double-subtraction trick, but the low-half is being used as a control and so can't be borrowed.
+To work around not being able to operate on the borrowed low-half bits while using them as a control, we use reverse-controlling and some knowledge of what state the low-half must be in if operations it is controlling are firing.
+
+To setup the reverse-control, we add and subtract the low-half out of the high-half, but frame the addition with NOT gates controlled by all bits in the low half.
+When any of the bits in the low-half is off, the NOT gates don't fire and the addition and subtraction will cancel each other.
+When all of the bits in the low-half are on, i.e. when the low-half is storing the two's complement representation of -1, the NOT gates do fire.
+This reverses the addition into a subtraction, and the low-half (which is storing -1) is subtracted out of the high-half twice.
+Therefore the high-half was incremented by 2.
+To halve the +2 into a +1, we prepend a dirty least-significant-bit onto the target register.
+
+Recall that, earlier in the paper, we used increments and decrements to implement addition where the target was larger than the input.
+To avoid an expensive cyclic dependency in our increment construction, it must not use larger-target addition; we can only perform additions and subtractions with input and target registers of the same size.
+This means that the construction described so far in this section only works for odd-sized registers.
+To handle even-sized registers, simply treat the LSB as a control determining whether the odd-sized remainder of the register is incremented then toggle the LSB.
 
@@
 Bit permuting operations can usually be emulated by re-labelling qubits, so they are easy to overlook in circuits.
-But some of our circuit diagrams have used controlled bit rotations, requiring actual operations, so we provide the constructions in figures \ref{fig:bit-rotate} and \ref{fig:bit-reverse} for completeness.
+But some of our circuit diagrams have used controlled bit rotations which require actual gates.
+We provide the relevant constructions in figures \ref{fig:bit-rotate} and \ref{fig:bit-reverse}.
 
@@
   \caption{\em
-    An inline controlled bit order reversal.
+    Inline controlled bit order reversal on an odd-sized register of size $n=2k+1$, and on an even-sized register of size $n=2k+2$.
     Each XOR operation is a series of independent CNOTs (note that the inputs have opposite endian-ness to the outputs).
-    For reversals on an even number of bits, swap the center two bits beforehand then use one of them as the center workspace bit from this construction.
     Uses no ancillae, $O(c + n)$ gates, and $O(c + n)$ depth.
@@
 
-Several of our constructions have used CNOTs with many controls and many targets.
-To avoid paying the overhead of $c$ controls for every target, we use one big CNOT to toggle-control all the others.
-We then use the available dirty ancillae to reduce the single remaining CNOT with many controls into Toffoli gates.
+Several of our constructions have used CNOTs applied to wire bundles, with many controls and many targets.
+A naive approach to implementing these operations would be to apply a separate NOT, each controlled by every control, to every target.
+But this would use $O(n \cdot c)$ gates, which is not linear in the number of controls.
+
+To avoid paying the overhead of $c$ controls for every target, we toggle {\em one} target conditioned on all of the controls.
+We then use toggle-controlling to spread the toggling effect to all of the other targets.
+
+To efficiently reduce the single remaining CNOT with $c$ controls into constant-sized Toffoli gates, we take advantage of the many available dirty ancilla.
 See figures \ref{fig:multi-not} and \ref{fig:cnot-reduction}.
@@
   \caption{\em
-    Separating controlled not gates with many controls and many targets into single-target single-control not-gates and one single-target many-control not-gate.
-    Uses no ancillae, $O(c + n)$ gates, and $O(c + \lg n)$ depth.
+    Reducing a many-control many-target CNOT into one many-control single-target CNOT and many single-target single-control CNOTs.
+    Uses no ancillae, $O(c + n)$ gates, and $O(c + n)$ depth where $n$ is the number of targets and $c$ is the number of controls.
+    The depth can be reduced to $O(c + \lg n)$ by spreading the toggling effect more intelligently.
   }
@@
 In figure \ref{fig:dependencies} we show a dependency graph of the constructions discussed in this paper.
-The total cost of period-finding, using the full construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
+The total cost of the period-finding step in Shor's algorithm, using our full construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
 We perform $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
@@
 Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
-Pivot flips also require ancillae (two, actually), but the ancillae can be dirty and, in the context of period-finding, there's always qubits available to borrow whenever a pivot-flip is needed.
+Pivot flips also require ancillae (two, actually), but the ancillae can be dirty and, in the context of period-finding, there are always qubits available to borrow whenever a pivot-flip is needed.
 This improvement saves a qubit in the overall circuit, reducing the total number of qubits required for period finding (on modular exponentiation against an $n$-bit modulus) from $2n+2$ to $2n+1$.
@@
 Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
+
 Note that, for classical reversible computation, 1 dirty ancilla is the minimum possible for an incrementer.
-The parity of the state permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even, and odd-parity permutations can't be implemented by composing even-parity permutations.
-When quantum operations are available, the final ancilla can be cut without increasing the asymptotic cost (see figure \ref{fig:bootstrap-ancilla}).
+An increment operation on $n$ bits is equivalent to performing $2^n-1$ state-swaps that move the state $|0\rangle$ from the bottom of the state-space to the top one step at a time.
+Note that $2^n-1$ is odd for non-trivial $n$, and therefore the parity of the state permutation performed by an increment operation is odd.
+However, the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
+Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation (that involves every bit) with constant-sized operations.
+An unused bit (i.e. a dirty ancilla) must be present.
 
-We tested individual constructions in \href{https://github.com/Strilanc/Quirk}{Quirk} \cite{quirk2016} and the full construction down to Toffoli gates in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
+When quantum operations are available, an increment can be performed without any ancilla.
+The asymptotic cost is the same, but the constant factors are worse.
+See figure \ref{fig:bootstrap-ancilla} for the construction.
+
+We used \href{https://github.com/Strilanc/Quirk}{Quirk} \cite{quirk2016} to find, explore \cite{victor2013}, check, and refine individual constructions.
+The full construction down to Toffoli gates was tested in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
 The source code for the ProjectQ implementation, and an issue tracker for submitting errata, can be found online at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
@@
 Although the asymptotic costs of our constructions match previous work, the constant factors are significantly worse.
-For example, our equivalent of a controlled 32-qubit multiplication uses roughly 1.3 million Toffoli gates (the specific number depends on the factor, the modulus, whether or not a number of optimizations are used, and whether or not those optimizations apply to a given case).
+For example, our equivalent of a controlled 32-qubit multiplication uses roughly 1.3 million Toffoli gates (the specific number depends on the factor to multiply by, the modulus, whether or not a number of optimizations are used, and whether or not those optimizations apply to a given case).
 This cost is over an order of magnitude worse than previous work \cite{haner2016}.
@@
 All the other constructions, for simpler arithmetic, can be done with two or fewer dirty ancillae.
-Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there's always more than enough dirty ancillae available to implement the simpler arithmetic inline.
-There's so much slack in the simpler arithmetic constructions that improving the number of qubits used by period-finding effectively means improving the very first step, reducing multiplication into some other operation, and nothing else.
+Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there are always more than enough dirty ancillae available to implement the simpler arithmetic inline.
+Because there is slack in the simpler arithmetic constructions, improving the overall number of qubits used by Shor's algorithm requires only that we improve the first step of reducing modular multiplication into some other operation.
 
-For example, consider a situation where we could cut multiplication operations into two pieces.
-If we knew the factorization $p \cdot q$ of the modulus $R$, with $p$ and $q$ each having size $\approx n/2$, then the Chinese remainder theorem guarantees we could uniquely represent $x \mod R$ as the pair $(x \mod p, x \mod q)$.
-Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, so there would be three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) for a total qubit count of $\approx 1.5n$.
+One way we could reduce modular multiplication into better operations is by knowing the factorization $p \cdot q$ of the modulus $R$, with $p$ and $q$ each having size $\approx n/2$.
+The Chinese remainder theorem guarantees that $x \mod R$ could be uniquely represented as the pair pf half-sized values $(x \mod p, x \mod q)$.
+Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) and a total qubit count of $\approx 1.5n$.
 Despite the large reduction in the number of ancillae available to the underlying arithmetic operations, each operation would still have more than enough ancillae and contribute no additional ancillae to the overall circuit.
 (Note: because the initial value $x_0 = 1$ is trivial, and the final value is measured or discarded, it's not necessary to implement circuits that translate between the Chinese-remainder representation and the usual 2s-complement representation.)
-Of course, since we use period-finding to compute the factorization of $R$, using the factorization of $R$ to optimize period-finding would be paradoxical.
+Of course, since we use Shor's algorithm to compute the factorization of $R$, using the factorization of $R$ to optimize Shor's algorithm would be paradoxical.
 
@@
 The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.
-We consider this to be a step towards a period-finding construction that uses only $n + O(1)$ qubits total, which could reduce the expected lifetime of current cryptographic keys by years (depending on the efficiency of the construction and how quickly quantum computers will gain error-corrected capacity).
+We consider this to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
+By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the expected time until current cryptographic keys can be broken by months or years (depending on how quickly computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
+
+We also described how to perform various arithmetic operations using only dirty ancilla, and the techniques we used to find the constructions (e.g.
+reverse controlling, toggle controlling, pivot flipping, and exploration by directly manipulating circuit diagrams).
 

### 2017-05-28 doc/paper.tex
Cleanup pass

@@
 One part of a circuit can borrow dirty bits from another part of the same circuit, so circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancillae are too far away to be acquired quickly.
-When attempting to reduce the number of bits or qubits required by a circuit, converting clean ancillae to dirty ancillae is a useful intermediate goal.
-In this paper, we reduce the number of qubits required to find the period of exponentiation, modulo an $n$-bit modulus, from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits and $n-1$ dirty qubits.
+When attempting to reduce the number of bits or qubits required by a circuit, replacing constructions that use clean ancillae with ones that use dirty ancillae is a useful intermediate goal.
+
+In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
 We do so without increasing the asymptotic costs, though the constant factors of our constructions are worse.
@@
 Our paper is structured as follows.
-Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions will follow.
-In section \ref{sec:construct} we describe the circuit constructions we use to reduce period finding into constant-sized gates, while tracking the number of required dirty ancillae.
+Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions and circuit diagrams will follow.
+In section \ref{sec:construct} we describe all the circuit constructions we use to reduce the period finding step of Shor's algorithm into constant-sized gates, while tracking the number of required dirty ancillae.
 Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
@@
 all the non-modular arithmetic constructions we discus actually perform arithmetic modulo $2^n$, where $n$ is the size of the target register).
+
 All diagrams order qubits from least significant bit (LSB) at the top to most significant bit (MSB) at the bottom.
@@
 All our circuit diagrams annotate operations with the number of clean and/or dirty ancillae they need.
-For example, a dashed line from an operation down to a triangle inscribed with "3 dirty" means that the operation needs 3 dirty ancilla.
+For example, a dashed line from an operation down to a triangle inscribed with ``3 dirty" means that the operation needs 3 dirty ancilla.
 If there are 3 unused wires that the operation can borrow, then the triangle will be green and a note of "(satisfied)" will be written underneath.
@@
 Nearly all the circuit constructions we present are classical reversible circuits (i.e.
-they do not use any quantum operations), so tricks such as the bootstrapping shown in figure \ref{fig:bootstrap-ancilla} could cut the ancilla count of several of the constructions.
-However, none of these improvements decrease the ancillae required at a crucial bottleneck, and so they do not reduce the number of qubits required by our construction of Shor's algorithm.
+they do not use any quantum operations), so tricks such as the ancilla bootstrap shown in figure \ref{fig:bootstrap-ancilla} could cut the ancilla count in several places.
+However, none of these improvements decrease the ancillae required at any crucial bottleneck, and so they do not reduce the number of qubits required by our overall construction of Shor's algorithm.
 
@@
 
-In the case of operations parametrized by compile-time constants, diagrams will use the letters "K" and/or "R".
-For example, a box showing "+K (mod R)" refers to the modular-offset operation $|x\rangle \rightarrow |x+K \pmod{R}\rangle$.
-We always use $R$ t to refer to the modulus that is factored by Shor's algorithm.
+In the case of operations parametrized by compile-time constants, diagrams will use the letters ``K" and/or ``R".
+For example, a box showing ``+K (mod R)" refers to the modular-offset operation $|x\rangle \rightarrow |x+K \pmod{R}\rangle$.
+We always use $R$ to to refer to the modulus that is factored by Shor's algorithm.
 In several constructions we assume that $R$ is odd.
@@
 
-Because adding a constant into a register is more expensive than adding one register into another, we avoid ambiguity between the two by always referring to an operation that adds a constant as an "offset operation".
+Because adding a constant into a register is more expensive than adding one register into another, we avoid ambiguity between the two operations by always referring to the operation that adds a constant as an ``offset operation".
 We will refer to the operation that adds one register into another as just ``addition" or, if needed for clarity, ``enregistered addition".
@@
 When a single-qubit gate is applied to a wire bundle, the gate applies separately to every wire.
-For example, when a not gate controlled by one wire bundle is applied to another wire bundle, every wire in the latter is toggled if and only if every wire in the former is on.
+For example, when a NOT gate controlled by one wire bundle is applied to another wire bundle, every wire in the latter is toggled when and only when every wire in the former is on.
 
-For completeness, even when the period finding reduction doesn't require a controlled version of an operation, we nevertheless provide a controllable construction that scale linearly with the number of controls.
-For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted (a decrement is a reversed increment, a multiply-add is a reversed multiply-subtract, a modular division is a reversed modular multiplication, a modular halving is a reversed modular double, and so forth).
+For completeness, even when the period finding reduction doesn't require a controlled version of an operation, we nevertheless provide a controllable construction that scales linearly with the number of controls.
+
+For brevity, we take for granted that each circuit's inverse operation is decomposed into the same operations as the original operation but run in reverse order and with each sub-operation inverted.
+A decrement is a reversed increment, a multiply-add is a reversed multiply-subtract, a modular division is a reversed modular multiplication, a modular halving is a reversed modular double, and so forth.
 
@@
 
-\subsection{Period Finding}
+\subsection{Period Finding for Modular Exponentiation}
 
@@
 The randomly chosen $B$ must be co-prime to $R$.
-When a $B$ is accidentally chosen that is not co-prime to $R$, the lucky victim can instead factor $R$ by recursively factoring $R_1 = \gcd(B, R)$ and $R_2 = R / \gcd(B, R)$.
+When a $B$ that is not co-prime to $R$ is accidentally chosen, the lucky victim can instead factor $R$ by recursively factoring $R_1 = \gcd(B, R)$ and $R_2 = R / \gcd(B, R)$.
 
@@
 In the case where $K$ has no multiplicative inverse modulo $R$, this construction will not work (it would define an invalid irreversible operation).
-However, that would mean $K$ is a factor of $R$; a case that the very lucky user can check for and handle classically before bothering to factor $R$ with a quantum computer.
+However, that would mean $K$ is a factor of $R$; a case that can be checked for and handled classically before bothering to factor $R$ with a quantum computer.
 
@@
 
-To multiply a register by 2 modulo an odd $R$, we note that the permutation this operation performs is to take the values from $[(R+1)/2, R)$, the top half, and interleave them between the values from $[0, (R+1)/2)$, the bottom half.
+To multiply a register by 2 modulo an odd $R$, we note that the permutation this operation performs is to take the values from $[(R+1)/2, R)$, the top-half of the valid range, and interleave them between the values from $[0, (R+1)/2)$, the bottom-half of the valid range.
 
-We can perform this interleaving by moving the top half up until it starts at $2^{n-1}$, i.e. aligns with the MSB boundary.
+We can perform this interleaving by moving the valid-top-half up until it starts at $2^{n-1}$, i.e. aligns with the MSB boundary.
 A left-rotate of the register bits then moves the MSB to the LSB, converting the alignment into interleaving.
@@
 Because it's expensive to perform an offset that affects values above $R/2$ while leaving values below $R/2$ alone, the circuit shown in figure \ref{fig:modular-double} uses a different alignment strategy.
-It starts by offset-ing the whole range down by $\lceil R/2 \rceil$, which wraps the bottom half of the input around to the top of the register's range.
-This offset also aligns what was originally the top half against the bottom of the register range.
-Then the circuit uses a controlled offset to move just the top half of the register up by $\lceil R/2 \rceil$, re-wrapping what was originally the bottom-half past the top and around to the middle, aligning with the MSB boundary.
-The circuit then toggles the MSB, fixing the fact that, although the top half is aligned with 0 and the bottom half is aligned with $2^{n-1}$, that's exactly the reverse of what we want.
+It starts by offset-ing the whole range down by $\lceil R/2 \rceil$, which wraps the valid-bottom-half of the input around to the top of the register's range.
+This offset also aligns what was originally the valid-top-half against the bottom of the register range.
+Then the circuit uses a controlled offset to move just the top half of the register (including the invalid range above $R$) up by $\lceil R/2 \rceil$, re-wrapping what was originally the valid-bottom-half past the top and around to the middle, aligning it with the MSB boundary.
+The circuit then toggles the MSB, fixing the fact that, although the valid-top-half is aligned with 0 and the valid-bottom-half is aligned with $2^{n-1}$, that's exactly the reverse of what we want.
 With the desired alignment achieved, the circuit interleaves the two halves with a left-rotate.
@@
 For states equal to or larger than the pivot, the ancilla does not get toggled and so the bi-flips undo each other.
+Therefore only the range below the pivot is flipped.
 
@@
 To add $K$ into a register modulo $R$, perform pivot-flips with the pivot at $R-K$, then $R$, then $K$.
-(The pivot-flip at $R$ can be simplified because the values above $R$ don't matter.)
 See figure \ref{fig:controlled-modular-add} for the circuit.
+(An optimization we don't show is that, because the values above $R$ don't matter, the pivot-flip at $R$ can be replaced by a bi-flip.)
 
-Interestingly, modular addition can use its own controls as dirty bits that would otherwise be required.
+Interestingly, controlled modular addition can borrow its own controls as dirty bits.
 For modular offset (i.e. adding a compile-time constant into a register), the two dirty bits are required whether or not the operation is controlled.
@@
   }
-  \label{fig:negate-mod}
+  \label{fig:compare}
 \end{figure}
@@
 When the target register is larger than the input register, the addition circuit in fig \ref{fig:inlineadder} can be modified without increasing the asymptotic cost.
-We do this by removing the surround CNOTs and adding three increment/decrement operations to the circuit, as follows.
+We do this by removing the surrounding CNOTs and adding three increment/decrement operations to the circuit, as follows.
 (Note that, to avoid cyclic dependencies, the increment and decrement constructions described in the next subsection will only use the same-register-size adder.)
 
-First, because there are many target bits not reached by the time we have swept through the input, carry signals can't reach the high bit of the target register.
-We fix this by replacing innermost CNOT with a controlled increment.
-Second and third, because we're using the MSB of the input register to hold the carry signal, it causes an increment at the LSB instead of at the correct position.
+First, because there are many target bits not reached by the time we have swept through the entire input register, carry signals aren't reaching the high bit of the target register.
+We fix this by replacing innermost CNOT, the carry-propagating CNOT, with a controlled increment.
+Second and third, because we're using the MSB of the input register as the carry signal, it causes an increment at the LSB instead of at the correct position.
 We undo the LSB increment with a controlled decrement, and use a controlled increment to apply the MSB's effect where it should have actually gone.
@@
 
-To control addition gates, and offset gates, we use ``reverse-controlling".
-Reverse-controlling controls an operation by instead controlling simpler framing operations that, if preserve, reverse the effect of the framed operation.
-In particular, note that framing the target of a subtraction with NOT gates (i.e. applying a NOT gate to every target wire before and after the addition) reverses the subtraction into an addition.
+To control addition gates, and offset gates, we use ``commutator controlling": finding operations $G$ and $H$ such that their group-theory commutator $[G, H] = G \cdot H \cdot G^\dagger \cdot H^\dagger$ equals a desired operation $U$ and either $G$ or $H$ can be efficiently controlled.
+In particular, we focus on the case where $G^2 = U$ and $H$ satisfies $H \cdot G \cdot H^\dagger = G^\dagger$ (i.e. framing $H$ with $G$ inverts the effect of $H$).
 
-By following the framed-subtraction with an addition, we have created a controlled operation that almost does what is needed.
-When the controls are not satisfied, the NOT gates bordering the subtraction do not fire, the subtraction and addition cancel each other, and nothing happens.
-When the controls are satisfied, the NOT gates reverse the subtraction into an addition, that addition and the second addition both apply, and the input is added into the target twice.
-
-To add $A$ instead of $2 \cdot A$ into the target, we temporarily prepend a dirty bit onto the target register.
-See figure \ref{fig:controlled-addition}.
+In the case of addition, we have the property $\lnot (\lnot x + K) = \lnot (-x - 1 + K) = -(-x - 1 + K) - 1 = x-K$.
+So a valid inverting operation $H$ for addition is a NOT gate applied to every target wire; $H=X^{\otimes n}$.
+If we then consider $G=\text{S}_{-K}$, where $S$ is the offset-by-$K$ operation, we find that $[G, H] = \text{S}_{2K}$.
+We can control this constructed $\text{S}_{2K}$ operation by adding controls only to the multi-not operation $H$, which is cheap to do.
+To add $K$ instead of $2K$, we temporarily prepend a dirty LSB onto the target register.
+We show this construction in figure \ref{fig:controlled-addition}.
 
@@
       Offset circuit from \cite{haner2016}.
+      $e$ is either 0 or 1.
       Uses $1$ dirty ancilla, $O(n \lg n)$ gates, and $O(n)$ depth (by overlapping the recursive cases).
@@
     Reducing controlled addition to uncontrolled addition.
-  	Uses $O(c + n)$ gates, $O(c + n)$ depth, and a dirty ancilla.
-  	When applied to offset gates, uses $O(c + n \lg n)$ gates, $O(n + c)$ depth, and two dirty ancillae.
+    Uses $O(c + m)$ gates, $O(c + m)$ depth, and a dirty ancilla.
+    When applied to offset gates, uses $O(c + m \lg m)$ gates, $O(m + c)$ depth, and two dirty ancillae.
+    $c$ is the number of controls, $n$ is the size of the input, and $m > n$ is the size of the target register.
   }
@@
 If the increment is uncontrolled, the low-half can be incremented with the double-subtraction trick by borrowing the high half.
-If there are controls, we combine the double-subtraction trick with the reverse-control technique we used to control addition.
-(Interestingly, combining the two fixes the double-effect problem that would have otherwise required us to prepend a dirty LSB onto the low-half.)
+When there are controls, we instead increment the low-half using a commutator-control construction where one operation subtracts the ancilla register out of the target and the other applies a NOT gate to every input wire and target wire.
+This works because $\lnot (\lnot T + \lnot K) - K = \lnot (-T - 1 - K - 1) - K = (T+2+K) - 1 - K = T + 1$.
 
-The high-half is trickier to deal with; we want to borrow the low-half for the double-subtraction trick, but the low-half is being used as a control and so can't be borrowed.
-To work around not being able to operate on the borrowed low-half bits while using them as a control, we use reverse-controlling and some knowledge of what state the low-half must be in if operations it is controlling are firing.
+The high-half is trickier to deal with.
+We want to borrow the low-half for the double-subtraction trick, but the low-half is being used as a control and so can't be borrowed.
+To work around not being able to operate on the borrowed low-half bits while using them as a control, we use more commutator control tricks and some knowledge of what state the low-half must be in if operations it is controlling are firing.
 
-To setup the reverse-control, we add and subtract the low-half out of the high-half, but frame the addition with NOT gates controlled by all bits in the low half.
+We add and subtract the low-half out of the high-half, but frame the addition with NOT gates controlled by all bits in the low half.
 When any of the bits in the low-half is off, the NOT gates don't fire and the addition and subtraction will cancel each other.
 When all of the bits in the low-half are on, i.e. when the low-half is storing the two's complement representation of -1, the NOT gates do fire.
-This reverses the addition into a subtraction, and the low-half (which is storing -1) is subtracted out of the high-half twice.
+This inverts the addition into a subtraction, and the low-half (which is storing -1) is subtracted out of the high-half twice.
 Therefore the high-half was incremented by 2.
-To halve the +2 into a +1, we prepend a dirty least-significant-bit onto the target register.
+To halve the +2 into a +1, we prepend a dirty LSB onto the target register.
 
 Recall that, earlier in the paper, we used increments and decrements to implement addition where the target was larger than the input.
-To avoid an expensive cyclic dependency in our increment construction, it must not use larger-target addition; we can only perform additions and subtractions with input and target registers of the same size.
-This means that the construction described so far in this section only works for odd-sized registers.
-To handle even-sized registers, simply treat the LSB as a control determining whether the odd-sized remainder of the register is incremented then toggle the LSB.
+To avoid an expensive cyclic dependency in our increment construction, the additions we use must not be larger-target additions.
+(Also, when the two registers are not the same size, they define arithmetic modulo different powers of 2.
+This breaks properties such as $T + \lnot K = T - K - 1$.)
+
+Because we can only perform additions and subtractions with input and target registers of the same size, the construction described so far in this section only works for odd-sized registers.
+To handle even-sized registers, we decompose the increment into a controlled-increment and a NOT gate.
+We use the LSB as a control determining whether the (odd-sized) rest of the register is incremented, and then toggle the LSB.
 
@@
 
-In this paper we presented Toffoli-based arithmetic circuits that reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
-The constructions also reduce the number of {\em clean} qubits required by $n$, so the number of clean qubits required is now $n + O(1)$.
-We consider this to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
-By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the expected time until current cryptographic keys can be broken by months or years (depending on how quickly computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
+In this paper we described how to perform various arithmetic operations using only dirty ancilla.
+We also described various techniques we used to find the constructions: pivot flipping, toggle controlling, commutator controlling, and exploration by directly manipulating circuit diagrams.
+We showed how these constructions reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
+The constructions also reduce the number of {\em clean} qubits required by $n$.
 
-We also described how to perform various arithmetic operations using only dirty ancilla, and the techniques we used to find the constructions (e.g.
-reverse controlling, toggle controlling, pivot flipping, and exploration by directly manipulating circuit diagrams).
+The number of clean qubits required to perform Shor's algorithm is now $n + O(1)$.
+We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
+By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
 

### 2017-05-28 doc/paper.tex
Table

@@
 \usepackage{hyperref}
+\usepackage{pbox}
 \title{Factoring with $n+2$ clean qubits and $n-1$ dirty qubits}
@@
 
-In figure \ref{fig:dependencies} we show a dependency graph of the constructions discussed in this paper.
-The total cost of the period-finding step in Shor's algorithm, using our full construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
-We perform $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
+\begin{figure*}
+  \centering
+  \makebox[\linewidth]{
+    \begin{tabular}{ l | c c c c c l }
+       & Year & Depth & Gates & Clean Qubits & Total Qubits \\
+      \hline
+      Shor \cite{Shor1999} & 1994 & $\Theta(n M(n))$ & $\Theta(n M(n))$ & $\Theta(n)$ & $\Theta(n)$ \\
+      Beckman et al. \cite{beckman1996} & 1996 & $\Theta(n^3)$ & $\Theta(n^3)$ & $5n + 1$ & $5n + 1$ \\
+      Veldral et al. \cite{vedral1996} & 1996 & $\Theta(n^3)$ & $\Theta(n^3)$ & $4n + 3$ & $4n + 3$ \\
+      Beauregard \cite{beauregard2003} & 2003 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+3$ & $2n+3$ \\
+      Takahashi et al. \cite{takahashi2006} & 2006 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+2$ & $2n+2$ \\
+      (ours) & 2017 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $n+2$ & $2n+1$ \\
+    \end{tabular}
+  }
+  \caption{\emph{•}
+    Space-efficient constructions of Shor's algorithm over time.
+    The table only includes papers that improved the number of qubits required to perform the algorithm and gave explicit circuit constructions.
+    $M(n)$ is the classical time-complexity of multiplication, which is known to be asymptotically at most $n \cdot (\lg n) \cdot 2^{O(\lg^* n)}$ \cite{furer2007}.
+  }
+  \label{fig:table-over-time}
+\end{figure*}
 
-Our main improvements over previous constructions are 1) the use of pivot-flips for modular addition, 2) the use of bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
+Recall that figure \ref{fig:dependencies} shows a dependency graph of the constructions discussed in this paper.
+Our asymptotic costs are dominated by performing $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
+The total cost of the period-finding step in Shor's algorithm, using our construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
+
+In figure \ref{fig:table-over-time} we show how the number of qubits needed for Shor's algorithm has improved over time.
+The previous improvement, from $2n+3$ to $2n+2$, occurred over a decade ago.
+
+Our main improvements over previous arithmetic constructions are 1) the use of pivot-flips for modular addition, 2) the use of bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
 
 Previous modular addition constructions worked by temporarily storing an is-wraparound-needed comparison in a clean ancilla \cite{takahashi2006, haner2016}.
-Pivot flips also require ancillae (two, actually), but the ancillae can be dirty and, in the context of period-finding, there are always qubits available to borrow whenever a pivot-flip is needed.
-This improvement saves a qubit in the overall circuit, reducing the total number of qubits required for period finding (on modular exponentiation against an $n$-bit modulus) from $2n+2$ to $2n+1$.
+Pivot flips also require ancillae, but the ancillae can be dirty and, in the context of Shor's algorithm, there are always qubits available to borrow whenever a pivot-flip is needed.
+This improvement ends up saving a qubit, reducing the total number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
 
@@
 We use an an extra multiply-accumulate and negation to prevent the ancilla register from leaking into the work register, and add a clean-up multiplication informed by the work register at the end of the circuit to undo any effects on the ancilla register.
-This improvement allows $n-1$ of the qubits in the ancilla register to be dirty (only the MSB must be zero; to ensure the register's value is in range).
+This improvement allows $n-1$ of the qubits in the ancilla register to be dirty.
+Only the MSB has to stay clean, to ensure the register's value is less than the modulus being factored.
 
 Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancillae \cite{draper2000, barenco1995}.
-Our classical incrementer constructions use $O(n)$ gates and a single dirty ancilla.
+Our classical incrementer construction uses $O(n)$ gates and a single dirty ancilla.
 
 Note that, for classical reversible computation, 1 dirty ancilla is the minimum possible for an incrementer.
-An increment operation on $n$ bits is equivalent to performing $2^n-1$ state-swaps that move the state $|0\rangle$ from the bottom of the state-space to the top one step at a time.
+An increment operation on $n$ bits is equivalent to performing a state-permutation that uses $2^n-1$ swaps to move the state $|0\rangle$ from the bottom of the state-space to the top of the state-space one step at a time.
 Note that $2^n-1$ is odd for non-trivial $n$, and therefore the parity of the state permutation performed by an increment operation is odd.
 However, the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
-Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation (that involves every bit) with constant-sized operations.
-An unused bit (i.e. a dirty ancilla) must be present.
+Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation that covers every bit with constant-sized operations.
+An uncovered bit (i.e. a dirty ancilla) must be present.
 
-When quantum operations are available, an increment can be performed without any ancilla.
+When quantum operations are available, the parity barrier can be bypassed and incrementing can be performed without any ancilla.
 The asymptotic cost is the same, but the constant factors are worse.

### 2017-05-28 doc/paper.tex
Polish pass

@@
 \usepackage{hyperref}
-\usepackage{pbox}
 \title{Factoring with $n+2$ clean qubits and $n-1$ dirty qubits}
@@
 When a result would exceed the size of a register, it wraps (i.e.
-all the non-modular arithmetic constructions we discus actually perform arithmetic modulo $2^n$, where $n$ is the size of the target register).
+all the non-modular arithmetic constructions we discuss actually perform arithmetic modulo $2^n$, where $n$ is the size of the target register).
 
@@
 For example, a dashed line from an operation down to a triangle inscribed with ``3 dirty" means that the operation needs 3 dirty ancilla.
-If there are 3 unused wires that the operation can borrow, then the triangle will be green and a note of "(satisfied)" will be written underneath.
+If there are 3 unused wires that the operation can borrow, then the triangle will be green and a note of ``(satisfied)" will be written underneath.
 If there are not enough unused wires shown in the diagram, the triangle will be yellow.
@@
 The ``A" symbol refers to the value of the input register, in the computational basis, at the time of the operation.
-Diagrams will often use a sequence of operations each with their own ``Input A" specific to that operation.
-This re-use of ``A" does not indicate any relation between those inputs.
+Diagrams will often use a sequence of operations each with their own ``Input A" box specific to that operation.
+This re-use of ``A" does not indicate any relation between those input values.
 The use of ``A" is merely a convention for indicating how the input-indicating box and the effect-indicating box are supposed to combine to form a single operation.
@@
 For example, a box showing ``+K (mod R)" refers to the modular-offset operation $|x\rangle \rightarrow |x+K \pmod{R}\rangle$.
-We always use $R$ to to refer to the modulus that is factored by Shor's algorithm.
+We use $R$ to refer to the modulus in modular-arithmetic operations, and to the value that is factored by Shor's algorithm.
 In several constructions we assume that $R$ is odd.
-Removing factors of 2 from a factoring problem is trivial, so this assumption does not decrease the generality of the construction.
+Removing factors of 2 from a factoring problem is trivial, so this assumption does not decrease the generality of the overall construction.
 
@@
 
-We perform modular multiplication with multiply-accumulates and an ancilla register as in \cite{beauregard2003}.
+We perform modular multiplication with modular scaled-addition operations and an ancilla register as in \cite{beauregard2003}.
 However, to allow our ancilla register to be mostly dirty, we extend the operation so that it has a well defined effect on the second register: multiplying by the inverse factor.
@@
 
-Because our modular circuit constructions would have undefined behavior if any register has a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
+Because our modular circuit constructions would have undefined behavior if any register had a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
 We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e.
@@
 
-As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-accumulate operations.
-However, to allow the second register to be dirty, we use an additional scaled-accumulate as well as a negation operation.
+As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-addition operations.
+However, to allow the second register to be dirty, we use an additional scale-add as well as a negation operation.
 See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
@@
   \caption{\em
-    Controlled modular bimultiplication, with a constant multiplier $K$, using three modular scaled-accumulates and a swap.
+    Controlled modular bimultiplication, with a constant multiplier $K$, using three modular scaled-additions and a swap.
     $K$ must have a multiplicative inverse modulo $R$.
@@
 
-\subsection{Modular Scaled-Accumulation}
+\subsection{Modular Scaled-Addition}
 
-To perform scale-accumulate operations, we use a shift-and-add approach similar to \cite{beauregard2003}.
+To perform modular scale-add operations, we use a shift-and-add approach similar to \cite{beauregard2003}.
 We right-shift (i.e. divide by $2 {\pmod R}$) the target register $n-1$ times, then begin iteratively left-shifting (i.e. multiplying by $2 {\pmod R}$) and adding $K$ into the target.
@@
   \caption{\em
-    Reducing a controlled modular scale-accumulate, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
+    Reducing a controlled modular scale-add, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
     Requires the modulus $R$ to be odd.
@@
 A bi-flip at $K$ preserves $x<K$ for all $x$.
-This allows the toggling of the control qubit, that determines whether a toggled-controlled bi-flip will happen, to be based on a comparison of the pivot against the target register we are operating on.
+This allows the toggling of the control qubit, that determines whether a toggle-controlled bi-flip will happen, to be based on a comparison of the pivot against the target register we are operating on.
 Even at points where the target may or may not have been bi-flipped.
@@
 
-Because the construction uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
+Because the construction in figure \ref{fig:inlineadder} uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
 (When the target register is smaller, we simply ignore the high bits of the input due to overflow wraparound.)
@@
 
-When the target register is larger than the input register, the addition circuit in fig \ref{fig:inlineadder} can be modified without increasing the asymptotic cost.
+When the target register is larger than the input register, the addition circuit in figure \ref{fig:inlineadder} can be modified without increasing the asymptotic cost.
 We do this by removing the surrounding CNOTs and adding three increment/decrement operations to the circuit, as follows.
@@
 First, because there are many target bits not reached by the time we have swept through the entire input register, carry signals aren't reaching the high bit of the target register.
-We fix this by replacing innermost CNOT, the carry-propagating CNOT, with a controlled increment.
+We fix this by replacing the innermost CNOT, the carry-propagating CNOT, with a controlled increment.
 Second and third, because we're using the MSB of the input register as the carry signal, it causes an increment at the LSB instead of at the correct position.
@@
 To control addition gates, and offset gates, we use ``commutator controlling": finding operations $G$ and $H$ such that their group-theory commutator $[G, H] = G \cdot H \cdot G^\dagger \cdot H^\dagger$ equals a desired operation $U$ and either $G$ or $H$ can be efficiently controlled.
-In particular, we focus on the case where $G^2 = U$ and $H$ satisfies $H \cdot G \cdot H^\dagger = G^\dagger$ (i.e. framing $H$ with $G$ inverts the effect of $H$).
+In particular, we focus on the case where $G^2 = U$ and $H$ satisfies $H \cdot G \cdot H^\dagger = G^\dagger$ (i.e. framing $G$ with $H$ inverts the effect of $G$).
 
@@
 So a valid inverting operation $H$ for addition is a NOT gate applied to every target wire; $H=X^{\otimes n}$.
-If we then consider $G=\text{S}_{-K}$, where $S$ is the offset-by-$K$ operation, we find that $[G, H] = \text{S}_{2K}$.
+If we then consider $G=\text{S}_{\text{-} K}$, where $S$ is an offset-by-subscript operation, we find that $[G, H] = \text{S}_{2K}$.
 We can control this constructed $\text{S}_{2K}$ operation by adding controls only to the multi-not operation $H$, which is cheap to do.
@@
 
-If the increment is uncontrolled, the low-half can be incremented with the double-subtraction trick by borrowing the high half.
+If the increment is uncontrolled, the low-half can be incremented with the double-subtraction trick by borrowing the high-half.
 When there are controls, we instead increment the low-half using a commutator-control construction where one operation subtracts the ancilla register out of the target and the other applies a NOT gate to every input wire and target wire.
@@
 We add and subtract the low-half out of the high-half, but frame the addition with NOT gates controlled by all bits in the low half.
-When any of the bits in the low-half is off, the NOT gates don't fire and the addition and subtraction will cancel each other.
+When any of the bits in the low-half are off, the NOT gates don't fire and the addition and subtraction will cancel each other.
 When all of the bits in the low-half are on, i.e. when the low-half is storing the two's complement representation of -1, the NOT gates do fire.
@@
 Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}.
-We use an an extra multiply-accumulate and negation to prevent the ancilla register from leaking into the work register, and add a clean-up multiplication informed by the work register at the end of the circuit to undo any effects on the ancilla register.
+We use an extra scale-add and negation to prevent the ancilla register from leaking into the work register, and add a clean-up multiplication informed by the work register at the end of the circuit to undo any effects on the ancilla register.
 This improvement allows $n-1$ of the qubits in the ancilla register to be dirty.
-Only the MSB has to stay clean, to ensure the register's value is less than the modulus being factored.
+The MSB has to stay clean to ensure the register's value is less than the modulus being factored.
 
-Previous published incrementers (not counting an unpublished version of our construction \cite{gidney2015} cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancillae \cite{draper2000, barenco1995}.
+Previous published incrementers (not counting an unpublished version of our construction being cited by \cite{haner2016}) required either $O(n^2)$ gates or $\omega(1)$ ancillae \cite{draper2000, barenco1995}.
 Our classical incrementer construction uses $O(n)$ gates and a single dirty ancilla.
@@
 Note that, for classical reversible computation, 1 dirty ancilla is the minimum possible for an incrementer.
-An increment operation on $n$ bits is equivalent to performing a state-permutation that uses $2^n-1$ swaps to move the state $|0\rangle$ from the bottom of the state-space to the top of the state-space one step at a time.
+An increment operation on $n$ bits is equivalent to performing a state-permutation that uses $2^n-1$ swaps to sweep the state $|0\rangle$ from the bottom of the state-space to the top of the state-space one step at a time.
 Note that $2^n-1$ is odd for non-trivial $n$, and therefore the parity of the state permutation performed by an increment operation is odd.
 However, the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
-Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation that covers every bit with constant-sized operations.
+Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation, that covers every bit, with constant-sized operations.
 An uncovered bit (i.e. a dirty ancilla) must be present.
@@
 We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
-By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
+By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly quantum computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
 

### 2017-05-29 src/dirty_period_finding/decompositions/addition_rules.py
Switch addition_test to check_permutation_decomposition / decomposition_to_ascii

@@ -2,7 +2,6 @@
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule
-from projectq.meta import Control
 from projectq.ops import X
 
 from dirty_period_finding.extensions import (
@@ -18,6 +17,8 @@ from dirty_period_finding.gates import (
     Increment,
     Decrement,
     MultiNot,
+    Add,
+    Subtract,
 )
 
 
@@ -182,60 +183,62 @@ def do_addition(input_reg, target_reg, dirty, controls):
 
     # Remove controls with double-add-invert trick.
     expanded = [dirty] + target_reg
-    for _ in range(2):
-        do_addition_no_controls(input_reg, expanded)
-        MultiNot & controls | expanded
-        MultiNot | expanded
+    Add | (input_reg, expanded)
+    MultiNot & controls | expanded
+    Subtract | (input_reg, expanded)
+    MultiNot & controls | expanded
 
 
-def do_addition_with_all_ops_controlled(eng, input_reg, target_reg, controls):
-    with Control(eng, controls):
-        do_addition_no_controls(input_reg, target_reg)
+# Trivial case: empty input or empty target. Do nothing.
+decompose_addition_no_op = DecompositionRule(
+    gate_class=AdditionGate,
+    gate_decomposer=lambda cmd: None,
+    gate_recognizer=lambda cmd:
+        len(cmd.qubits[0]) == 0 or len(cmd.qubits[1]) == 0)
+
+# Trivial case: single target bit. Just a controlled NOT.
+decompose_addition_single_target = DecompositionRule(
+    gate_class=AdditionGate,
+    gate_recognizer=
+        min_register_sizes(1, 1) &
+        max_register_sizes(float('inf'), 1),
+    gate_decomposer=lambda cmd:
+        X & cmd.control_qubits & cmd.qubits[0][0] | cmd.qubits[1][0])
+
+# Trivial case: single input bit. Controlled increment.
+# Minimum control requirement avoids cyclic decomp w.r.t. increment.
+decompose_addition_single_input = DecompositionRule(
+    gate_class=AdditionGate,
+    gate_recognizer=
+        min_controls(1) &
+        min_register_sizes(1, 1) &
+        max_register_sizes(1, float('inf')),
+    gate_decomposer=lambda cmd:
+        Increment & cmd.control_qubits & cmd.qubits[0] | cmd.qubits[1])
+
+# Additions without controls can be done inline.
[…]

### 2017-05-29 test/_test_util.py

@@ -252,7 +252,7 @@ def fuzz_permutation_circuit(register_sizes,
 def cover(n, cut=10):
     if n < cut:
         return range(n)
-    return random.randint(0, n - 1),
+    return list(range(2)) + [random.randint(2, n - 1) for _ in range(3)]
 
 
 def _inverse_decomposition_rule(gate, rule):

### 2017-05-29 test/decompositions/addition_test.py

@@ -2,15 +2,12 @@
 from __future__ import division
 from __future__ import unicode_literals
 
-import itertools
 import random
 
 from projectq import MainEngine
 from projectq.cengines import (DummyEngine,
                                DecompositionRuleSet)
-from projectq.ops import X
 from projectq.setups.decompositions import swap2cnot
-from projectq.types import Qureg
 
 from dirty_period_finding.decompositions import (
     multi_not_rules,
@@ -18,7 +15,11 @@ from dirty_period_finding.decompositions import (
     increment_rules
 )
 from dirty_period_finding.decompositions.addition_rules import (
-    do_addition_with_same_size_and_no_controls
+    decompose_addition_controlled,
+    decompose_addition_no_op,
+    decompose_addition_single_input,
+    decompose_addition_single_target,
+    decompose_addition_uncontrolled,
 )
 from dirty_period_finding.extensions import (
     LimitedCapabilityEngine,
@@ -26,112 +27,222 @@ from dirty_period_finding.extensions import (
 )
 from dirty_period_finding.gates import (
     Add,
-    Subtract,
-    MultiNot,
-    IncrementGate,
-    DecrementGate,
 )
-from .._test_util import fuzz_permutation_circuit, check_permutation_circuit
-
-
-def test_exact_commands_for_small_circuit():
-    backend = DummyEngine(save_commands=True)
-    eng = MainEngine(backend=backend, engine_list=[])
-    src = eng.allocate_qureg(2)
-    dst = eng.allocate_qureg(2)
-    backend.received_commands = []
-    do_addition_with_same_size_and_no_controls(src, dst)
-
-    a, c = src[0], src[1]
-    x, y = dst[0], dst[1]
-    assert backend.received_commands == [
-        (MultiNot & c).generate_command(Qureg([y, x, a])),
-        (X & c).generate_command(x),
-        (X & x & c).generate_command(a),
-        (X & x & a).generate_command(c),
-        (X & c).generate_command(y),
-        (X & x & a).generate_command(c),
-        (X & x & c).generate_command(a),
-        (X & a).generate_command(x),
-        (MultiNot & c).generate_command(Qureg([a, x, y])),
-    ]
+from .._test_util import (
+    check_permutation_decomposition,
+    cover,
+    decomposition_to_ascii)
 
 
 def test_toffoli_size_of_addition():
     backend = DummyEngine(save_commands=True)
     eng = MainEngine(backend=backend, engine_list=[
         AutoReplacerEx(DecompositionRuleSet(modules=[
[…]

### 2017-05-29 doc/paper.tex
Minor tweaks

@@
 So a valid inverting operation $H$ for addition is a NOT gate applied to every target wire; $H=X^{\otimes n}$.
-If we then consider $G=\text{S}_{\text{-} K}$, where $S$ is an offset-by-subscript operation, we find that $[G, H] = \text{S}_{2K}$.
+If we then consider $G=\text{S}_{\text{-} K}$, where $S$ is an offset-by-subscript operation, we find that $[G, H] = [\text{S}_{-K}, X^{\otimes n}] = \text{S}_{2K}$.
 We can control this constructed $\text{S}_{2K}$ operation by adding controls only to the multi-not operation $H$, which is cheap to do.
@@
 However, the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
-Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to perform an increment operation, that covers every bit, with constant-sized operations.
+Since odd-parity permutations can't be implemented by composing even-parity permutations, it is impossible to compose smaller operations into an increment operation that covers every bit.
 An uncovered bit (i.e. a dirty ancilla) must be present.
@@
 The full construction down to Toffoli gates was tested in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
-The source code for the ProjectQ implementation, and an issue tracker for submitting errata, can be found online at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
+The python source code for our test implementation, and an issue tracker for submitting errata, can be found online at \href{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}{https://github.com/Strilanc/PaperImpl-2017-DirtyPeriodFinding}.
 
@@
 Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancillae used to implement modular multiplication in terms of scaled modular addition.
-All the other constructions, for simpler arithmetic, can be done with two or fewer dirty ancillae.
-Furthermore, because scaled modular addition is reduced into operations that have a $\Theta(n)$ surplus of unused qubits, in context there are always more than enough dirty ancillae available to implement the simpler arithmetic inline.
+Except for multiplication, all our arithmetic constructions use two or fewer dirty ancillae.
+Since we reduce modular scale-addition into operations that have a $\Theta(n)$ surplus of unused qubits, in context there are always more than enough dirty ancillae available to implement the simpler arithmetic inline.
 Because there is slack in the simpler arithmetic constructions, improving the overall number of qubits used by Shor's algorithm requires only that we improve the first step of reducing modular multiplication into some other operation.
@@
 One way we could reduce modular multiplication into better operations is by knowing the factorization $p \cdot q$ of the modulus $R$, with $p$ and $q$ each having size $\approx n/2$.
-The Chinese remainder theorem guarantees that $x \mod R$ could be uniquely represented as the pair pf half-sized values $(x \mod p, x \mod q)$.
+The Chinese remainder theorem guarantees that $x \mod R$ could be uniquely represented as the pair of half-sized values $(x \mod p, x \mod q)$.
 Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) and a total qubit count of $\approx 1.5n$.
@@
 We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
-By allowing keys to fit into quantum computers with half as much space, such a construction could reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly quantum computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
+Such a construction could, by allowing factoring problems to fit into quantum computers with half as much space, reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly quantum computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
 

### 2017-05-29 src/dirty_period_finding/decompositions/increment_rules.py

@@ -37,7 +37,7 @@ def do_increment_with_no_controls_and_n_dirty(target_reg, dirty_reg):
        │    │        │   │   │   │
        ┤    ├       ─┤   ├───┤   ├───
        │    │        │   │   │   │
-       ┤ +1 ├       ─┤−=A├───┤−=A├───
+       ┤ +1 ├       ─┤ −A├───┤ −A├───
        │    │        │   │   │   │
        ┤    ├       ─┤   ├───┤   ├───
        │    │        │   │   │   │
@@ -80,10 +80,10 @@ def do_increment_with_1_dirty(target_reg, dirty_qubit, controls):
         c                 c
        ━/━━━━●━━━        ━/━━━━━━━━━━━━━━●━━━━━━━●━━━━━━━━━━●━━━━━━━●━━━━━
         n ┌──┴─┐          n        ┌───┐ │ ┌───┐ │    ┌───┐ │ ┌───┐ │
-       ━/━┥    ┝━        ━/━━━━━━━━┥ A ┝━●━┥ A ┝━●━━━━┥+=A┝━⊕━┥−=A┝━⊕━━━━━
+       ━/━┥    ┝━        ━/━━━━━━━━┥ A ┝━●━┥ A ┝━●━━━━┥ +A┝━⊕━┥ −A┝━⊕━━━━━
        n-1│ +1 │         n-1       ├───┤ │ ├───┤ │    ├───┤ │ ├───┤ │
        ━/━┥    ┝━        ━/━━━╲ ╱──┤   ├─⊕─┤   ├─⊕────┤   ├─⊕─┤   ├─⊕─╲ ╱━
-          └────┘    =          ╳   │−=A│ │ │+=A│ │    │inp│ │ │inp│ │  ╳
+          └────┘    =          ╳   │ −A│ │ │ +A│ │    │inp│ │ │inp│ │  ╳
        ──────────        ─────╱ ╲━━┥   ┝━⊕━┥   ┝━⊕━━━━┥ A ┝━⊕━┥ A ┝━⊕━╱ ╲─
                                    └───┘   └───┘      └───┘   └───┘
     Args:

### 2017-05-30 src/dirty_period_finding/__init__.py
License headers

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 import decompositions
 import extensions
 import gates

### 2017-05-30 src/dirty_period_finding/decompositions/__init__.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 import addition_rules
 import bootstrap_ancilla_rules
 import increment_rules

### 2017-05-30 src/dirty_period_finding/decompositions/increment_rules.py

@@ -1,5 +1,17 @@
 # -*- coding: utf-8 -*-
 
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from projectq.cengines import DecompositionRule
 from projectq.ops import X
 

### 2017-05-30 src/dirty_period_finding/decompositions/modular_double_rules.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.cengines import DecompositionRule

### 2017-05-30 src/dirty_period_finding/decompositions/offset_rules.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 import math

### 2017-05-30 src/dirty_period_finding/extensions/__init__.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from . import _monkey_patching
 from ._basic_gate_ex import (
     BasicGateEx,

### 2017-05-30 src/dirty_period_finding/extensions/_basic_gate_ex.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 import projectq.ops

### 2017-05-30 src/dirty_period_finding/extensions/_cached_auto_replacer.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.cengines import (

### 2017-05-30 src/dirty_period_finding/extensions/_classical_simulator.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 """
 A simulator that only permits classical operations, for faster/easier testing.
 """

### 2017-05-30 src/dirty_period_finding/extensions/_command_ex.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.ops import Command

### 2017-05-30 src/dirty_period_finding/extensions/_command_predicates.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.types import WeakQubitRef

### 2017-05-30 src/dirty_period_finding/extensions/_limited_capability_engine.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.cengines import BasicEngine

### 2017-05-30 src/dirty_period_finding/extensions/_monkey_patching.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 import weakref
@@ -178,6 +192,24 @@ def simulator_receive(self, command_list):
             self.send([cmd])
 
 
+original_handle = Simulator._handle
+_x_list = [[0, 1], [1, 0]]
+
+
+def sim_handle(self, cmd):
+    if cmd.gate.__class__ is XGate:
+        target = [cmd._qubits[0][0].id]
+        controls = [qb.id for qb in cmd.control_qubits]
+        self._simulator.apply_controlled_gate(_x_list, target, controls)
+        if not self._gate_fusion:
+            self._simulator.run()
+    elif cmd.gate.__class__ is FlushGate:
+        self._simulator.run()
+    else:
+        original_handle(self, cmd)
+
+
+# Well well well, isn't this a terrifying little block of code.
 Command.__str__ = command_str
 Command.__repr__ = command_str
 Qureg.__str__ = qureg_str
@@ -194,22 +226,4 @@ XGate.__and__ = gate_and
 ZGate.__and__ = gate_and
 XGate.__eq__ = lambda self, other: other.__class__ is XGate
 XGate.__hash__ = lambda self: hash((XGate,))
-
-original_handle = Simulator._handle
-
-
-_x_list = [[0, 1], [1, 0]]
-
-
-def sim_handle(self, cmd):
-    if cmd.gate.__class__ is XGate:
-        target = [cmd._qubits[0][0].id]
-        controls = [qb.id for qb in cmd.control_qubits]
-        self._simulator.apply_controlled_gate(_x_list, target, controls)
-        if not self._gate_fusion:
-            self._simulator.run()
-    elif cmd.gate.__class__ is FlushGate:
-        self._simulator.run()
-    else:
-        original_handle(self, cmd)
 Simulator._handle = sim_handle

### 2017-05-30 src/dirty_period_finding/extensions/_permutation_simulator.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 import numpy as np

### 2017-05-30 src/dirty_period_finding/extensions/_to_ascii.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from collections import defaultdict

### 2017-05-30 src/dirty_period_finding/find_period.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import print_function
 from __future__ import unicode_literals
 

### 2017-05-30 src/dirty_period_finding/gates/__init__.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from ._addition import Add, Subtract, AdditionGate, SubtractionGate
 from ._comparison import (
     PredictOffsetOverflowGate,

### 2017-05-30 src/dirty_period_finding/gates/_comparison.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from dirty_period_finding.extensions import (

### 2017-05-30 src/dirty_period_finding/gates/_modular_addition.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from dirty_period_finding.extensions import BasicMathGateEx

### 2017-05-30 src/dirty_period_finding/gates/_modular_double.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from ._modular_bimultiplication import multiplicative_inverse

### 2017-05-30 src/dirty_period_finding/gates/_phase_gradient.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from fractions import Fraction

### 2017-05-30 src/dirty_period_finding/gates/_pivot_flip.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from dirty_period_finding.extensions import BasicMathGateEx, SelfInverseGateEx

### 2017-05-30 src/dirty_period_finding/gates/_rotate_bits.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.ops import NotMergeable

### 2017-05-30 src/dirty_period_finding/gates/_vector_phaser.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 import abc
@@ -9,11 +22,7 @@ import numpy as np
 import projectq.ops
 from projectq.ops import NotMergeable
 
-from dirty_period_finding.extensions import (
-    BasicGateEx,
-    BasicMathGateEx,
-    GateWithCurriedControls,
-)
+from dirty_period_finding.extensions import BasicGateEx
 
 
 def _exp_pi_i(f):

### 2017-05-30 test/__init__.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 import os
 import sys
 

### 2017-05-30 test/_test_util.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import division
 from __future__ import print_function
 from __future__ import unicode_literals

### 2017-05-30 test/decompositions/modular_negate_test.py

@@ -1,25 +1,30 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import division
 from __future__ import unicode_literals
-
-import fractions
 
 from projectq import MainEngine
 from projectq.cengines import DummyEngine, DecompositionRuleSet
 from projectq.setups.decompositions import swap2cnot
 
 from dirty_period_finding.decompositions import (
-    modular_bimultiplication_rules,
     multi_not_rules,
     addition_rules,
     increment_rules,
     pivot_flip_rules,
-    modular_scaled_addition_rules,
-    modular_addition_rules,
     offset_rules,
-    modular_double_rules,
-    rotate_bits_rules,
-    reverse_bits_rules,
     comparison_rules,
     modular_negate_rules,
 )

### 2017-05-30 test/decompositions/offset_test.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import division
 from __future__ import unicode_literals
 

### 2017-05-30 test/extensions/__init__.py

@@ -1,0 +1,13 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.

### 2017-05-30 test/extensions/limited_capability_engine_test.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 import pytest
 from projectq import MainEngine
 from projectq.cengines import DummyEngine

### 2017-05-30 test/extensions/permutation_simulator_test.py

@@ -1,3 +1,17 @@
+# -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from projectq import MainEngine
 from projectq.ops import BasicMathGate, X
 

### 2017-05-30 test/extensions/to_ascii_test.py

@@ -1,4 +1,17 @@
 # -*- coding: utf-8 -*-
+
+#   Licensed under the Apache License, Version 2.0 (the "License");
+#   you may not use this file except in compliance with the License.
+#   You may obtain a copy of the License at
+#
+#       http://www.apache.org/licenses/LICENSE-2.0
+#
+#   Unless required by applicable law or agreed to in writing, software
+#   distributed under the License is distributed on an "AS IS" BASIS,
+#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+#   See the License for the specific language governing permissions and
+#   limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.ops import X, Swap

### 2017-05-30 src/dirty_period_finding/__init__.py
Fix license headers

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 import decompositions
 import extensions

### 2017-05-30 src/dirty_period_finding/decompositions/__init__.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 import addition_rules
 import bootstrap_ancilla_rules

### 2017-05-30 src/dirty_period_finding/decompositions/increment_rules.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from projectq.cengines import DecompositionRule
 from projectq.ops import X

### 2017-05-30 src/dirty_period_finding/extensions/__init__.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from . import _monkey_patching
 from ._basic_gate_ex import (

### 2017-05-30 src/dirty_period_finding/extensions/_classical_simulator.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 """
 A simulator that only permits classical operations, for faster/easier testing.

### 2017-05-30 src/dirty_period_finding/find_period.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from __future__ import print_function
 from __future__ import unicode_literals

### 2017-05-30 src/dirty_period_finding/gates/__init__.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from ._addition import Add, Subtract, AdditionGate, SubtractionGate
 from ._comparison import (

### 2017-05-30 src/dirty_period_finding/gates/_modular_negate.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from __future__ import unicode_literals
 

### 2017-05-30 src/dirty_period_finding/gates/_multi_not.py

@@ -1,4 +1,19 @@
 # -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
 from __future__ import unicode_literals
 
 from projectq.ops import SelfInverseGate

### 2017-05-30 test/__init__.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 import os
 import sys

### 2017-05-30 test/_test_util.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from __future__ import division
 from __future__ import print_function

### 2017-05-30 test/decompositions/__init__.py

@@ -1,13 +1,15 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.

### 2017-05-30 test/decompositions/modular_negate_test.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from __future__ import division
 from __future__ import unicode_literals

### 2017-05-30 test/extensions/limited_capability_engine_test.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 import pytest
 from projectq import MainEngine

### 2017-05-30 test/extensions/permutation_simulator_test.py

@@ -1,16 +1,18 @@
 # -*- coding: utf-8 -*-
 
-#   Licensed under the Apache License, Version 2.0 (the "License");
-#   you may not use this file except in compliance with the License.
-#   You may obtain a copy of the License at
+# Copyright 2017 Google Inc.
 #
-#       http://www.apache.org/licenses/LICENSE-2.0
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
 #
-#   Unless required by applicable law or agreed to in writing, software
-#   distributed under the License is distributed on an "AS IS" BASIS,
-#   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-#   See the License for the specific language governing permissions and
-#   limitations under the License.
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
 
 from projectq import MainEngine
 from projectq.ops import BasicMathGate, X

### 2017-06-01 doc/paper.tex
License headers
Update document style - Remove emphasis from figure captions - Use autoref for figure/section references - Use quantumarticle style - Add affiliation / email - Blue links

@@
-\documentclass[twocolumn]{article}
+\documentclass[twocolumn,longbibliography]{quantumarticle}
 \usepackage{amsmath}
 \usepackage{graphicx}
-\usepackage{hyperref}
+\usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
+\hypersetup{colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue}
+
 \title{Factoring with $n+2$ clean qubits and $n-1$ dirty qubits}
 \author{Craig Gidney}
+\affiliation{Google, Santa Barbara, CA 93117, USA}
+\email{craiggidney@google.com}
+
+\def\sectionautorefname{Section}
 
@@
 \begin{abstract}
-\em
 We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae in an unknown state that must be preserved.
@@
 Our paper is structured as follows.
-Section \ref{sec:introduction} introduces and describes the conventions our circuit constructions and circuit diagrams will follow.
-In section \ref{sec:construct} we describe all the circuit constructions we use to reduce the period finding step of Shor's algorithm into constant-sized gates, while tracking the number of required dirty ancillae.
-Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
-Finally, section \ref{sec:conclusion} concludes with a discussion of possible future improvements.
+\autoref{sec:introduction} introduces and describes the conventions our circuit constructions and circuit diagrams will follow.
+In \autoref{sec:construct} we describe all the circuit constructions we use to reduce the period finding step of Shor's algorithm into constant-sized gates, while tracking the number of required dirty ancillae.
+Then, in \autoref{sec:costs}, we discuss the novelty and comparative costs of the presented circuit constructions.
+Finally, \autoref{sec:conclusion} concludes with a discussion of possible future improvements.
 
@@
 Nearly all the circuit constructions we present are classical reversible circuits (i.e.
-they do not use any quantum operations), so tricks such as the ancilla bootstrap shown in figure \ref{fig:bootstrap-ancilla} could cut the ancilla count in several places.
+they do not use any quantum operations), so tricks such as the ancilla bootstrap shown in \autoref{fig:bootstrap-ancilla} could cut the ancilla count in several places.
 However, none of these improvements decrease the ancillae required at any crucial bottleneck, and so they do not reduce the number of qubits required by our overall construction of Shor's algorithm.
@@
 Diagrams will often use a sequence of operations each with their own ``Input A" box specific to that operation.
-This re-use of ``A" does not indicate any relation between those input values.
+This reuse of ``A" does not indicate any relation between those input values.
 The use of ``A" is merely a convention for indicating how the input-indicating box and the effect-indicating box are supposed to combine to form a single operation.
@@
 Our reduction, from the period finding step of Shor's algorithm down to constant-sized gates, uses many circuit constructions.
-Figure \ref{fig:dependencies} shows an overview of the various operations we will use, and the path through them that we will take.
+\autoref{fig:dependencies} shows an overview of the various operations we will use, and the path through them that we will take.
 
@@
   \includegraphics[height=17.3cm]{assets/dependencies.png}
-  \caption{\em
+  \caption{
     Graph of the transitive reduction of dependencies between constructions in our paper.
@@
   }
-  \caption{\em
+  \caption{
 	High-level definition of the period finding circuit \cite{Shor1999} used by Shor's algorithm.
@@
 
-A high-level view of the circuit for period finding applied to modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in figure \ref{fig:period-finding}.
+A high-level view of the circuit for period finding applied to modular exponentiation, the core quantum subroutine of Shor's quantum factoring algorithm \cite{Shor1999}, is shown in \autoref{fig:period-finding}.
 
@@
 Only one of the qubits in the phase-estimation register needs to be present at a time, and so the phase register can be reduced to a single repeatedly-used qubit \cite{zalka1998, mosca1999, parker2000, beauregard2003}.
-Figure \ref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
+\autoref{fig:period-finding-solo-phase-qubit} shows a period-finding circuit with this property.
 It uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations classically parametrized by previous measurements, and (d) qubit resets.
@@
 We measure the work register to recover the fixup factor needed to restore the dirty register's original value.
-Figure \ref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
+\autoref{fig:period-finding-solo-phase-qubit-explicit-dirty-register} shows this construction.
 
@@
   }
-  \caption{\em
+  \caption{
 	Period finding with a single phase-estimation qubit \cite{beauregard2003}.
@@
   }
-  \caption{\em
+  \caption{
 	Period finding with a single phase-estimation qubit and paired inverse multiplications (``bimultiplications").
@@
 However, to allow the second register to be dirty, we use an additional scale-add as well as a negation operation.
-See figure \ref{fig:controlled-modular-multiply} for the circuit diagram.
+See \autoref{fig:controlled-modular-multiply} for the circuit diagram.
 
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-multiply.png}
-  \caption{\em
+  \caption{
     Controlled modular bimultiplication, with a constant multiplier $K$, using three modular scaled-additions and a swap.
@@
 (The more significant bits go first because their effects must be hit by more left-shifts.)
-See figure \ref{fig:controlled-modular-scale-accumulate}.
+See \autoref{fig:controlled-modular-scale-accumulate}.
 
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
-  \caption{\em
+  \caption{
     Reducing a controlled modular scale-add, with a constant multiplier $K$, into the modular equivalent of shift-and-add.
@@
 
-Because it's expensive to perform an offset that affects values above $R/2$ while leaving values below $R/2$ alone, the circuit shown in figure \ref{fig:modular-double} uses a different alignment strategy.
+Because it's expensive to perform an offset that affects values above $R/2$ while leaving values below $R/2$ alone, the circuit shown in \autoref{fig:modular-double} uses a different alignment strategy.
 It starts by offset-ing the whole range down by $\lceil R/2 \rceil$, which wraps the valid-bottom-half of the input around to the top of the register's range.
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-double.png}
-  \caption{\em
+  \caption{
     Controlled modular doubling.
@@
 
-See figures \ref{fig:controlled-pivot-flip} and \ref{fig:controlled-const-pivot-flip} for the circuit diagrams.
+See \autoref{fig:controlled-pivot-flip} and \autoref{fig:controlled-const-pivot-flip} for the circuit diagrams.
 For states that are less than the pivot, the comparison against the pivot keeps toggling the ancilla and exactly one of the controlled bi-flips will fire.
@@
   \includegraphics[width=\linewidth]{assets/controlled-pivot-flip.png}
-  \caption{\em
+  \caption{
     Controlled pivot-flip circuit with an enregistered pivot.
@@
   \includegraphics[width=\linewidth]{assets/controlled-const-pivot-flip.png}
-  \caption{\em
+  \caption{
     Controlled pivot-flip circuit with a constant pivot.
@@
 
-As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition can be implemented by three pivot-flips.
+As shown in \autoref{fig:mod-add-from-pivot-flip-bars}, a modular addition can be implemented by three pivot-flips.
 To add $K$ into a register modulo $R$, perform pivot-flips with the pivot at $R-K$, then $R$, then $K$.
-See figure \ref{fig:controlled-modular-add} for the circuit.
+See \autoref{fig:controlled-modular-add} for the circuit.
 (An optimization we don't show is that, because the values above $R$ don't matter, the pivot-flip at $R$ can be replaced by a bi-flip.)
@@
 For modular offset (i.e. adding a compile-time constant into a register), the two dirty bits are required whether or not the operation is controlled.
-See figure \ref{fig:controlled-modular-offset}.
+See \autoref{fig:controlled-modular-offset}.
 
@@
   \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
-  \caption{\em
+  \caption{
      Modular addition of $A \pmod{R}$ can be done with three pivot flips.
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
-  \caption{\em
+  \caption{
     Controlled modular addition construction based on pivot-flips.
@@
   \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
-  \caption{\em
+  \caption{
     Controlled modular offset construction based on pivot-flips.
@@
 We do so by temporarily moving $|0\rangle$ out the way with a decrement, pivot-flipping at $R-1$, then undoing the decrement.
-See figure \ref{fig:negate-mod}.
+See \autoref{fig:negate-mod}.
 
@@
   \includegraphics[width=\linewidth]{assets/negate-mod.png}
-  \caption{\em
+  \caption{
     Controlled modular negation.
@@
 We implement comparisons as in \cite{takahashi2005}, using an addition followed by a slightly smaller subtraction that clears all changes except the overflow signal into the target bit.
-See figure \ref{fig:compare}.
+See \autoref{fig:compare}.
 
@@
   \includegraphics[width=\linewidth]{assets/comparison-less-const.png}
-  \caption{\em
+  \caption{
     Controlled comparison.
@@
 \cite{takahashi2005} provides a reversible adder circuit that uses $O(n)$ gates, $O(n)$ depth, and no ancillae.
-We show an equivalent circuit in figure \ref{fig:inlineadder}.
+We show an equivalent circuit in \autoref{fig:inlineadder}.
 
-Because the construction in figure \ref{fig:inlineadder} uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
+Because the construction in \autoref{fig:inlineadder} uses the input register as workspace, it doesn't work when the input is a compile-time constant or when the target register is larger than the input register.
 (When the target register is smaller, we simply ignore the high bits of the input due to overflow wraparound.)
@@
   \includegraphics[width=\linewidth]{assets/inline-adder.png}
-  \caption{\em Adder for input and target registers of the same size.
+  \caption{ Adder for input and target registers of the same size.
   Requires no ancillae, and uses $O(N)$ gates and depth.
@@
 When the value to be added into a register is a compile-time constant (i.e. when applying an offset gate), we use the offset construction from \cite{haner2016}.
-Their offset circuit, shown in figure \ref{fig:offset}, uses $O(n \lg n)$ gates, $O(n)$ depth, and one dirty ancilla.
+Their offset circuit, shown in \autoref{fig:offset}, uses $O(n \lg n)$ gates, $O(n)$ depth, and one dirty ancilla.
 
-When the target register is larger than the input register, the addition circuit in figure \ref{fig:inlineadder} can be modified without increasing the asymptotic cost.
+When the target register is larger than the input register, the addition circuit in \autoref{fig:inlineadder} can be modified without increasing the asymptotic cost.
 We do this by removing the surrounding CNOTs and adding three increment/decrement operations to the circuit, as follows.
@@
 We undo the LSB increment with a controlled decrement, and use a controlled increment to apply the MSB's effect where it should have actually gone.
-See the circuit diagram in figure \ref{fig:inline-adder-into-large}.
+See the circuit diagram in \autoref{fig:inline-adder-into-large}.
 
@@
 To add $K$ instead of $2K$, we temporarily prepend a dirty LSB onto the target register.
-We show this construction in figure \ref{fig:controlled-addition}.
+We show this construction in \autoref{fig:controlled-addition}.
 
@@
   \includegraphics[width=\linewidth]{assets/offset.png}
-  \caption{\em
+  \caption{
       Offset circuit from \cite{haner2016}.
@@
   \includegraphics[width=\linewidth]{assets/inline-adder-into-large.png}
-  \caption{\em
+  \caption{
       Adder with target larger than source, using no ancillae.
@@
   \includegraphics[width=\linewidth]{assets/controlled-addition.png}
-  \caption{\em
+  \caption{
     Reducing controlled addition to uncontrolled addition.
@@
 A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
-When $n$ dirty bits are available, $x$ can come from a register defined by those $n$ arbitrary bits, as shown in figure \ref{fig:increment-many-dirty}.
+When $n$ dirty bits are available, $x$ can come from a register defined by those $n$ arbitrary bits, as shown in \autoref{fig:increment-many-dirty}.
 
@@
   \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
-  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancillae.}
+  \caption{ Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancillae.}
   \label{fig:increment-many-dirty}
@@
 
-To improve from $n$ dirty bits to the single dirty bit used by the circuit in figure \ref{fig:controlled-increment-odd}, we break the register into two halves.
+To improve from $n$ dirty bits to the single dirty bit used by the circuit in \autoref{fig:controlled-increment-odd}, we break the register into two halves.
 A high-half that is incremented only if all of the low bits are on, and a low-half that is unconditionally incremented.
@@
   \includegraphics[width=\linewidth]{assets/controlled-increment-odd.png}
-  \caption{\em
+  \caption{
     Odd-sized controlled increment.
@@
   \includegraphics[width=\linewidth]{assets/ancilla-bootstrap.png}
-  \caption{\em Bootstrapping a dirty ancilla out of an increment gate using quantum operations.
+  \caption{ Bootstrapping a dirty ancilla out of an increment gate using quantum operations.
   The $i^{\text{frac}}$ gate is a ``phase gradient" operation that phases each computational basis state $|v\rangle$ by an amount proportional to $v/2^d$, where $d$ is the size of the register.
@@
 But some of our circuit diagrams have used controlled bit rotations which require actual gates.
-We provide the relevant constructions in figures \ref{fig:bit-rotate} and \ref{fig:bit-reverse}.
+We provide the relevant constructions in \autoref{fig:bit-rotate} and \autoref{fig:bit-reverse}.
 
@@
   \includegraphics[width=\linewidth]{assets/controlled-bit-rotate.png}
-  \caption{\em
+  \caption{
     A controlled bit rotation / bit swap is three controlled bit-reverses.
@@
   \includegraphics[width=\linewidth]{assets/controlled-reverse.png}
-  \caption{\em
+  \caption{
     Inline controlled bit order reversal on an odd-sized register of size $n=2k+1$, and on an even-sized register of size $n=2k+2$.
@@
 To efficiently reduce the single remaining CNOT with $c$ controls into constant-sized Toffoli gates, we take advantage of the many available dirty ancilla.
-See figures \ref{fig:multi-not} and \ref{fig:cnot-reduction}.
+See \autoref{fig:multi-not} and \autoref{fig:cnot-reduction}.
 
@@
   \includegraphics[width=\linewidth]{assets/multi-cnot-to-single-cnots.png}
-  \caption{\em
+  \caption{
     Reducing a many-control many-target CNOT into one many-control single-target CNOT and many single-target single-control CNOTs.
@@
   \includegraphics[width=\linewidth]{assets/cnot-reduction.png}
-  \caption{\em
+  \caption{
     When $c-2$ dirty ancillae are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates.
@@
   }
-  \caption{\emph{•}
+  \caption{
     Space-efficient constructions of Shor's algorithm over time.
@@
 
-Recall that figure \ref{fig:dependencies} shows a dependency graph of the constructions discussed in this paper.
+Recall that \autoref{fig:dependencies} shows a dependency graph of the constructions discussed in this paper.
 Our asymptotic costs are dominated by performing $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
@@
 
-In figure \ref{fig:table-over-time} we show how the number of qubits needed for Shor's algorithm has improved over time.
+In \autoref{fig:table-over-time} we show how the number of qubits needed for Shor's algorithm has improved over time.
 The previous improvement, from $2n+3$ to $2n+2$, occurred over a decade ago.
@@
 The asymptotic cost is the same, but the constant factors are worse.
-See figure \ref{fig:bootstrap-ancilla} for the construction.
+See \autoref{fig:bootstrap-ancilla} for the construction.
 
@@
 The Chinese remainder theorem guarantees that $x \mod R$ could be uniquely represented as the pair of half-sized values $(x \mod p, x \mod q)$.
-Multiplications of the half-sized values could use and re-use the same half-sized clean ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) and a total qubit count of $\approx 1.5n$.
+Multiplications of the half-sized values could use and reuse the same half-sized clean ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) and a total qubit count of $\approx 1.5n$.
 Despite the large reduction in the number of ancillae available to the underlying arithmetic operations, each operation would still have more than enough ancillae and contribute no additional ancillae to the overall circuit.
@@
 
+\section{Acknowledgements}
+
+We thank Matthew Neeley and Dave Bacon for comments on earlier versions of this paper, from which it greatly benefited.
+
 \bibliographystyle{plain}

### 2017-06-01 doc/paper.tex
Final draft
Polishing for release - Fixed title linking to a journal search page by editing quantumarticle.cls (not in diff) - Fixed comparison table not mentioning rotation synthesis costs, and added haner2016 since while keeping space same it improved time - Fixed figure links going to the caption instead of the image. - Fixed increment/decrement mixup when describing the parity barrier - Fixed an extremely awkward and

@@
 \usepackage[pdfpagelabels,pdftex,bookmarks,breaklinks]{hyperref}
+\usepackage[all]{hypcap}
 \hypersetup{colorlinks,citecolor=blue,urlcolor=blue,linkcolor=blue}
@@
     A controlled bit rotation / bit swap is three controlled bit-reverses.
-    Uses no ancillae, $O(n+c)$ gates, and $O(n+c)$ depth.
+    Uses no ancillae, $O(c + n)$ gates, and $O(c + n)$ depth.
   }
@@
       Veldral et al. \cite{vedral1996} & 1996 & $\Theta(n^3)$ & $\Theta(n^3)$ & $4n + 3$ & $4n + 3$ \\
-      Beauregard \cite{beauregard2003} & 2003 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+3$ & $2n+3$ \\
-      Takahashi et al. \cite{takahashi2006} & 2006 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+2$ & $2n+2$ \\
+      Beauregard \cite{beauregard2003} & 2003 & $\Theta(n^3 \lg \frac{1}{\varepsilon})$ & $\Theta(n^3 \lg \frac{n}{\varepsilon} \lg \frac{1}{\varepsilon})$ & $2n+3$ & $2n+3$ \\
+      Takahashi et al. \cite{takahashi2006} & 2006 & $\Theta(n^3 \lg \frac{1}{\varepsilon})$ & $\Theta(n^3 \lg \frac{n}{\varepsilon} \lg \frac{1}{\varepsilon})$ & $2n+2$ & $2n+2$ \\
+      H\"{a}ner et al. \cite{haner2016} & 2016 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+2$ & $2n+2$ \\
       (ours) & 2017 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $n+2$ & $2n+1$ \\
@@
     Space-efficient constructions of Shor's algorithm over time.
-    The table only includes papers that improved the number of qubits required to perform the algorithm and gave explicit circuit constructions.
+    The table only includes papers that presented explicit circuit constructions improving the minimum number of qubits required to perform the algorithm, or else improved the time complexity at the minimum known size.
     $M(n)$ is the classical time-complexity of multiplication, which is known to be asymptotically at most $n \cdot (\lg n) \cdot 2^{O(\lg^* n)}$ \cite{furer2007}.
+    $\varepsilon$ is the maximum error when synthesizing the circuit out of a fixed set of universal gates, which is asymptotically relevant for algorithms that use Draper addition \cite{draper2000} (i.e.
+applying phase gradients in frequency space) instead of Toffoli-based addition/offset constructions.
   }
@@
 Recall that \autoref{fig:dependencies} shows a dependency graph of the constructions discussed in this paper.
-Our asymptotic costs are dominated by performing $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ gates \cite{haner2016} and $O(n)$ depth.
+Our asymptotic costs are dominated by performing $O(n)$ modular multiplications, each of which uses $O(n)$ modular additions and offsets, each of which uses $O(n \lg n)$ constant-sized classical gates \cite{haner2016} and $O(n)$ depth.
 The total cost of the period-finding step in Shor's algorithm, using our construction, is $O(n^3 \lg n)$ gates and $O(n^3)$ depth.
@@
 In \autoref{fig:table-over-time} we show how the number of qubits needed for Shor's algorithm has improved over time.
-The previous improvement, from $2n+3$ to $2n+2$, occurred over a decade ago.
+The previous space improvement, from $2n+3$ to $2n+2$, occurred over a decade ago.
 
@@
 Note that, for classical reversible computation, 1 dirty ancilla is the minimum possible for an incrementer.
-An increment operation on $n$ bits is equivalent to performing a state-permutation that uses $2^n-1$ swaps to sweep the state $|0\rangle$ from the bottom of the state-space to the top of the state-space one step at a time.
+An increment operation on $n$ bits is equivalent to performing a state-permutation that uses $2^n-1$ swaps to sweep the state $|2^n-1\rangle$ from the top of the state-space to the bottom of the state-space one step at a time.
 Note that $2^n-1$ is odd for non-trivial $n$, and therefore the parity of the state permutation performed by an increment operation is odd.
@@
 When quantum operations are available, the parity barrier can be bypassed and incrementing can be performed without any ancilla.
-The asymptotic cost is the same, but the constant factors are worse.
+The asymptotic cost is slightly worse due to the cost of synthesizing tiny rotations, and the constant factors hidden by the asymptotic notation are also worse.
 See \autoref{fig:bootstrap-ancilla} for the construction.
 
-We used \href{https://github.com/Strilanc/Quirk}{Quirk} \cite{quirk2016} to find, explore \cite{victor2013}, check, and refine individual constructions.
+We used \href{https://github.com/Strilanc/Quirk}{Quirk} \cite{quirk2016} to explore \cite{victor2013}, check, and refine individual constructions.
 The full construction down to Toffoli gates was tested in \href{https://github.com/ProjectQ-Framework/ProjectQ}{ProjectQ} \cite{projq2016}.
@@
 This cost is over an order of magnitude worse than previous work \cite{haner2016}.
-The reason for the increased cost is that using dirty ancillae results in more uncomputation work, in repeating some operations twice and conditionally to do them once unconditionally, and in less efficient constructions in general.
+The reason for the increased cost is that using dirty ancillae results in more uncomputation work, in repeating some operations twice conditionally to do them once unconditionally, and in less efficient constructions in general.
 These inefficiencies stack multiplicatively when one construction uses another as a subroutine.

### 2017-06-02 doc/paper.tex
Typo fix
that optimizations will preserve the behavior

@@
 This {\em includes input values}.
-For example, if our modular addition construction ($x \mathrel{{+}{=}} y \pmod{R}$) is applied to a $y \geq R$, then we don't guarantee that $y \mod R$ will be added into $x$ or that the operation will commute with other additions or that anything sensible at all will happen.
+For example, if our modular addition construction ($x \mathrel{{+}{=}} y \pmod{R}$) is applied to a $y \geq R$, then we don't guarantee that $y \mod R$ will be added into $x$, that the operation will commute with other additions, that optimizations won't change the behavior, or that anything sensible at all will happen.
 If the precondition $y < R$ is violated, what occurs is undefined behavior \cite{c11}.
@@
   \caption{ Adder for input and target registers of the same size.
-  Requires no ancillae, and uses $O(N)$ gates and depth.
+  Requires no ancillae, and uses $O(n)$ gates and depth.
   Based on \cite{van2004, takahashi2005}.}
@@
   \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
-  \caption{ Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth, size, and $n$ dirty ancillae.}
+  \caption{ Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(n)$ depth and size, and $n$ dirty ancillae.}
   \label{fig:increment-many-dirty}

### 2017-06-03 src/dirty_period_finding/count_gates.py
Add factor.py and count_gates.py commands - Replaces half-broken find_period.py

@@ -1,0 +1,71 @@
+# -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Prints gate counts for various sizes of the controlled bimultiply gate."""
+
+from __future__ import print_function
+from __future__ import unicode_literals
+
+import fractions
+import random
+
+from projectq import MainEngine
+from projectq.backends import ResourceCounter
+from projectq.cengines import DecompositionRuleSet, DummyEngine
+
+import dirty_period_finding.decompositions as decompositions
+from dirty_period_finding.extensions import (
+    LimitedCapabilityEngine,
+    AutoReplacerEx,
+)
+from dirty_period_finding.gates import ModularBimultiplicationGate
+
+
+def main():
+    x = 2
+    modulus = 2
+
+    for reg_size in range(2, 20):
+        reg_max_val = 1 << reg_size
+        while True:
+            modulus = random.randint(reg_max_val // 2 + 1, reg_max_val - 1)
+            x = random.randint(2, modulus - 1)
+            if modulus % 2 != 0 and fractions.gcd(x, modulus) == 1:
+                break
+        cnt = ResourceCounter()
+        eng = MainEngine(backend=DummyEngine(), engine_list=[
+            AutoReplacerEx(DecompositionRuleSet(modules=[decompositions])),
+            LimitedCapabilityEngine(
+                allow_toffoli=True,
+                allow_single_qubit_gates=True,
+                allow_classes=[]
+            ),
+            cnt,
+        ])
+        v1 = eng.allocate_qureg(reg_size)
+        v2 = eng.allocate_qureg(reg_size)
+        c = eng.allocate_qubit()
+        gate = ModularBimultiplicationGate(x, modulus)
+        gate & c | (v1, v2)
+        print()
+        print()
+        print("Register Size: {}".format(reg_size))
+        print("Gate count for controlled {}".format(gate))
[…]

### 2017-06-03 src/dirty_period_finding/decompositions/__init__.py

@@ -29,6 +29,7 @@ import pivot_flip_rules
 import comparison_rules
 import reverse_bits_rules
 import rotate_bits_rules
+from projectq.setups.decompositions import swap2cnot as _swap2cnot
 
 all_defined_decomposition_rules = [
     rule
@@ -48,6 +49,7 @@ all_defined_decomposition_rules = [
         comparison_rules,
         reverse_bits_rules,
         rotate_bits_rules,
+        _swap2cnot,
     ]
     for rule in module.all_defined_decomposition_rules
 ]

### 2017-06-03 src/dirty_period_finding/factor.py

@@ -1,0 +1,203 @@
+# -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""Factors a number given as a command-line argument."""
+
+from __future__ import print_function
+from __future__ import unicode_literals
+
+import fractions
+import math
+import random
+import sys
+from fractions import Fraction
+
+from projectq import MainEngine
+from projectq.backends import Simulator
+from projectq.cengines import DecompositionRuleSet
+from projectq.ops import X, Z
+from projectq.types import Qureg
+
+import dirty_period_finding.decompositions as decompositions
+from dirty_period_finding.extensions import (
+    H,
+    LimitedCapabilityEngine,
+    AutoReplacerEx,
+)
+from dirty_period_finding.gates import ModularBimultiplicationGate
+
+
+# The whole point is to break things down into toffoli gates, but simulating
+# all of them makes things ridiculously slow. So... it's up to you, reader.
+DECOMPOSE_INTO_TOFFOLIS_AND_GO_VERY_VERY_SLOW = False
+
+
+def shor_find_period(base,
+                     modulus,
+                     precision,
+                     phase_qubit,
+                     work_qureg,
+                     ancilla_qureg):
+    """
+    Args:
+        base (int):
+        modulus (int):
+        precision (int):
+            The number of iterations.
+        phase_qubit (Qureg):
+            A clean zero-initialized qubit.
+        work_qureg (Qureg):
+            A clean zero-initialized register with lg(modulus) bits.
+        ancilla_qureg (Qureg): A semi-dirty register.
+            A mostly-dirty register with lg(modulus) bits.
+            Only the most significant bit must be clean and zero-initialised.
+
+    Returns:
+        int:
[…]

### 2017-06-03 src/dirty_period_finding/find_period.py

@@ -1,205 +1,0 @@
-# -*- coding: utf-8 -*-
-
-# Copyright 2017 Google Inc.
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#     http://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
-
-from __future__ import print_function
-from __future__ import unicode_literals
-
-import fractions
-import math
-import random
-from fractions import Fraction
-
-from projectq import MainEngine
-from projectq.backends import Simulator, ResourceCounter
-from projectq.cengines import DecompositionRuleSet, DummyEngine
-from projectq.ops import X, Z
-from projectq.setups.decompositions import swap2cnot
-from projectq.types import Qureg
-
-from dirty_period_finding.decompositions import *
-from dirty_period_finding.extensions import (
-    H,
-    LimitedCapabilityEngine,
-    AutoReplacerEx,
-)
-from dirty_period_finding.gates import *
-
-
-def sample_period(base,
-                  modulus,
-                  precision,
-                  phase_qubit,
-                  work_qureg,
-                  ancilla_qureg):
-    """
-    Args:
-        base (int):
-        modulus (int):
-        precision (int):
-            The number of iterations.
-        phase_qubit (Qureg):
-            A clean zero-initialized qubit.
-        work_qureg (Qureg):
-            A clean zero-initialized register with lg(modulus) bits.
-        ancilla_qureg (Qureg): A semi-dirty register.
-            A mostly-dirty register with lg(modulus) bits.
-            Only the most significant bit must be clean and zero-initialised.
-
-    Returns:
-        int:
-            A number related to the period.
-    """
-
-    # Incremental phase estimation.
-    X | work_qureg[0]
-    frac = Fraction(0, 1)
-    for i in range(precision):
-        rev_i = precision - i - 1
-        op = ModularBimultiplicationGate(pow(base, 1 << rev_i, modulus),
-                                         modulus)
-
[…]

### 2017-06-03 src/dirty_period_finding/decompositions/__init__.py
Added negate and scale gates - Monkey-patched a math function onto projectq's X gate - Added min parameter to cover

@@ -14,26 +14,32 @@
 # See the License for the specific language governing permissions and
 # limitations under the License.
 
-import addition_rules
-import bootstrap_ancilla_rules
-import increment_rules
-import modular_addition_rules
-import modular_bimultiplication_rules
-import modular_double_rules
-import modular_negate_rules
-import modular_scaled_addition_rules
-import multi_not_rules
-import offset_rules
-import phase_gradient_rules
-import pivot_flip_rules
-import comparison_rules
-import reverse_bits_rules
-import rotate_bits_rules
+from __future__ import absolute_import
+
 from projectq.setups.decompositions import swap2cnot as _swap2cnot
+
+from . import addition_rules
+from . import bootstrap_ancilla_rules
+from . import comparison_rules
+from . import increment_rules
+from . import modular_addition_rules
+from . import modular_bimultiplication_rules
+from . import modular_double_rules
+from . import modular_negate_rules
+from . import modular_scaled_addition_rules
+from . import multi_not_rules
+from . import negate_rules
+from . import offset_rules
+from . import phase_gradient_rules
+from . import pivot_flip_rules
+from . import reverse_bits_rules
+from . import rotate_bits_rules
+from . import scale_rules
 
 all_defined_decomposition_rules = [
     rule
     for module in [
+        _swap2cnot,
         addition_rules,
         bootstrap_ancilla_rules,
         increment_rules,
@@ -43,13 +49,14 @@ all_defined_decomposition_rules = [
         modular_negate_rules,
         modular_scaled_addition_rules,
         multi_not_rules,
+        negate_rules,
         offset_rules,
         phase_gradient_rules,
         pivot_flip_rules,
         comparison_rules,
         reverse_bits_rules,
         rotate_bits_rules,
-        _swap2cnot,
+        scale_rules,
     ]
     for rule in module.all_defined_decomposition_rules
 ]

### 2017-06-03 src/dirty_period_finding/decompositions/negate_rules.py

@@ -1,0 +1,45 @@
+# -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+from __future__ import unicode_literals
+
+from projectq.cengines import DecompositionRule
+
+from dirty_period_finding.gates import MultiNot, Increment, NegateGate
+
+
+def do_negate(target_reg, controls):
+    """
+    Args:
+        target_reg (projectq.types.Qureg):
+            The register to negate.
+        controls (list[Qubit]):
+            Control qubits.
+    """
+    MultiNot & controls | target_reg
+    Increment & controls | target_reg
+
+
+decompose_negate = DecompositionRule(
+    gate_class=NegateGate,
+    gate_decomposer=lambda cmd: do_negate(
+        target_reg=cmd.qubits[0],
+        controls=cmd.control_qubits))
+
+
+all_defined_decomposition_rules = [
+    decompose_negate,
+]

### 2017-06-03 src/dirty_period_finding/extensions/__init__.py

@@ -45,3 +45,4 @@ from ._command_predicates import (
 from ._limited_capability_engine import LimitedCapabilityEngine
 from ._permutation_simulator import PermutationSimulator
 from ._to_ascii import commands_to_ascii_circuit
+from ._util import multiplicative_inverse, extended_gcd

### 2017-06-03 src/dirty_period_finding/extensions/_monkey_patching.py

@@ -229,3 +229,4 @@ ZGate.__and__ = gate_and
 XGate.__eq__ = lambda self, other: other.__class__ is XGate
 XGate.__hash__ = lambda self: hash((XGate,))
 Simulator._handle = sim_handle
+XGate.get_math_function = lambda self, _: lambda x: (x[0] ^ 1,)

### 2017-06-03 src/dirty_period_finding/extensions/_util.py

@@ -1,0 +1,11 @@
+def extended_gcd(a, b):
+    if a == 0:
+        return b, 0, 1
+    g, y, x = extended_gcd(b % a, a)
+    return g, x - (b // a) * y, y
+
+
+def multiplicative_inverse(a, m):
+    a %= m
+    g, x, y = extended_gcd(a, m)
+    return None if g != 1 else x % m

### 2017-06-03 src/dirty_period_finding/gates/_modular_bimultiplication.py

@@ -18,7 +18,10 @@ from __future__ import unicode_literals
 
 from projectq.ops import NotMergeable
 
-from dirty_period_finding.extensions import BasicMathGateEx
+from dirty_period_finding.extensions import (
+    BasicMathGateEx,
+    multiplicative_inverse,
+)
 
 
 class ModularBimultiplicationGate(BasicMathGateEx):
@@ -70,16 +73,3 @@ class ModularBimultiplicationGate(BasicMathGateEx):
             '×{} (mod {})'.format(self.factor, self.modulus),
             '×{} (mod {})'.format(self.inverse_factor, self.modulus)
         ]
-
-
-def extended_gcd(a, b):
-    if a == 0:
-        return b, 0, 1
-    g, y, x = extended_gcd(b % a, a)
-    return g, x - (b // a) * y, y
-
-
-def multiplicative_inverse(a, m):
-    a %= m
-    g, x, y = extended_gcd(a, m)
-    return None if g != 1 else x % m

### 2017-06-03 src/dirty_period_finding/gates/_modular_double.py

@@ -16,8 +16,10 @@
 
 from __future__ import unicode_literals
 
-from ._modular_bimultiplication import multiplicative_inverse
-from dirty_period_finding.extensions import BasicMathGateEx
+from dirty_period_finding.extensions import (
+    BasicMathGateEx,
+    multiplicative_inverse,
+)
 
 
 class ModularDoubleGate(BasicMathGateEx):

### 2017-06-03 src/dirty_period_finding/gates/_scale.py

@@ -1,0 +1,76 @@
+# -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+from __future__ import unicode_literals
+
+from projectq.ops import NotMergeable
+
+from dirty_period_finding.extensions import (
+    BasicSizedMathGateEx,
+    multiplicative_inverse,
+)
+
+
+class ScaleGate(BasicSizedMathGateEx):
+    def __init__(self, factor=1, inverse_factor=1):
+        if factor & 1 == 0 or inverse_factor & 1 == 0:
+            raise ValueError("Can't reversibly scale by even factors.")
+        BasicSizedMathGateEx.__init__(self)
+        self.factor = factor
+        self.inverse_factor = inverse_factor
+
+    def net_factor_for_size(self, size):
+        mask = ~(~0 << size)
+        inv = multiplicative_inverse(self.inverse_factor, 1 << size)
+        return (self.factor * inv) & mask
+
+    def do_operation(self, sizes, args):
+        n = sizes[0]
+        mask = ~(~0 << n)
+        return (args[0] * self.net_factor_for_size(n)) & mask,
+
+    def get_inverse(self):
+        return ScaleGate(self.inverse_factor, self.factor)
+
+    def get_merged(self, other):
+        if not isinstance(other, ScaleGate):
+            raise NotMergeable()
+        return ScaleGate(self.factor * other.factor,
+                         self.inverse_factor * other.inverse_factor)
+
+    def __repr__(self):
+        return 'ScaleGate({}, {})'.format(self.factor, self.inverse_factor)
+
+    def __str__(self):
+        return repr(self)
+
+    def __eq__(self, other):
+        return (isinstance(other, ScaleGate) and
+                self.factor == other.factor and
+                self.inverse_factor == other.inverse_factor)
+
+    def __hash__(self):
+        return hash((ScaleGate, self.factor, self.inverse_factor))
+
[…]

### 2017-06-03 test/_test_util.py

@@ -264,7 +264,9 @@ def fuzz_permutation_circuit(register_sizes,
     assert outputs == actual_outputs
 
 
-def cover(n, cut=10):
+def cover(n, cut=10, min=0):
+    if min > 0:
+        return [e + min for e in cover(n - min, cut=cut + min)]
     if n < cut:
         return range(n)
     return list(range(2)) + [random.randint(2, n - 1) for _ in range(3)]

### 2017-06-03 test/decompositions/modular_addition_test.py

@@ -116,7 +116,7 @@ def test_toffoli_size_of_modular_offset():
 
 
 def test_decompose_modular_addition():
-    for register_size in range(1, 100):
+    for register_size in cover(100, min=1):
         for control_size in range(3):
             for h_modulus in cover((1 << register_size) - 1):
                 modulus = h_modulus + 1
@@ -130,7 +130,7 @@ def test_decompose_modular_addition():
 
 
 def test_decompose_modular_offset():
-    for register_size in range(1, 100):
+    for register_size in cover(100, min=1):
         for control_size in range(3):
             for h_modulus in cover(1 << register_size):
                 modulus = h_modulus + 1

### 2017-06-03 test/decompositions/negate_test.py

@@ -1,0 +1,102 @@
+# -*- coding: utf-8 -*-
+
+# Copyright 2017 Google Inc.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+from __future__ import division
+from __future__ import unicode_literals
+
+from projectq import MainEngine
+from projectq.cengines import (DummyEngine,
+                               DecompositionRuleSet)
+from projectq.setups.decompositions import swap2cnot
+
+from dirty_period_finding.decompositions import (
+    addition_rules,
+    increment_rules,
+    multi_not_rules,
+    negate_rules,
+)
+from dirty_period_finding.decompositions.negate_rules import (
+    decompose_negate,
+)
+from dirty_period_finding.extensions import (
+    LimitedCapabilityEngine,
+    AutoReplacerEx,
+)
+from dirty_period_finding.gates import (
+    Negate,
+)
+from .._test_util import (
+    check_permutation_decomposition,
+    cover,
+    decomposition_to_ascii,
+)
+
+
+def test_toffoli_size_of_negate():
+    backend = DummyEngine(save_commands=True)
+    eng = MainEngine(backend=backend, engine_list=[
+        AutoReplacerEx(DecompositionRuleSet(modules=[
+            addition_rules,
+            increment_rules,
+            multi_not_rules,
+            negate_rules,
+            swap2cnot,
+        ])),
+        LimitedCapabilityEngine(allow_toffoli=True),
+    ])
+    src = eng.allocate_qureg(100)
+    workspace = eng.allocate_qubit()
+    Negate | src
+
+    assert 2000 < len(backend.received_commands) < 4000
+    assert workspace is not None
+
+
+def test_decompose_addition_uncontrolled():
+    for register_size in cover(50):
+        for control_size in cover(3):
+            check_permutation_decomposition(
+                decomposition_rule=decompose_negate,
+                gate=Negate,
[…]

### 2017-06-03 test/decompositions/offset_test.py

@@ -22,6 +22,7 @@ from dirty_period_finding.decompositions.offset_rules import (
     decompose_into_range_increments,
     decompose_into_recursion,
     decompose_remove_controls,
+    decompose_decrease_size,
 )
 from dirty_period_finding.gates import OffsetGate
 from .._test_util import (
@@ -46,7 +47,7 @@ def test_estimate_cost_of_bitrange_offset():
 
 
 def test_decompose_into_range_increments():
-    for register_size in range(100):
+    for register_size in cover(100):
         for offset in cover(1 << register_size):
             if estimate_cost_of_bitrange_offset(offset, register_size) > 4:
                 continue
@@ -57,9 +58,20 @@ def test_decompose_into_range_increments():
                 register_sizes=[register_size])
 
 
-def test_decompose_into_recursion():
-    for register_size in range(100):
+def test_decompose_decrease_size():
+    for register_size in cover(100):
         for offset in cover(1 << register_size):
+            for control_size in cover(2):
+                check_permutation_decomposition(
+                    decomposition_rule=decompose_decrease_size,
+                    gate=OffsetGate(offset * 2),
+                    register_sizes=[register_size],
+                    control_size=control_size)
+
+
+def test_decompose_into_recursion():
+    for register_size in range(1, 10):
+        for offset in cover(1 << (register_size - 1)):
             check_permutation_decomposition(
                 decomposition_rule=decompose_into_recursion,
                 gate=OffsetGate(offset),
@@ -68,7 +80,7 @@ def test_decompose_into_recursion():
 
 
 def test_decompose_remove_controls():
-    for register_size in range(100):
+    for register_size in cover(100):
         for offset in cover(1 << register_size):
             for controls in range(1, 4):
                 check_permutation_decomposition(
@@ -77,6 +89,29 @@ def test_decompose_remove_controls():
                     register_sizes=[register_size],
                     workspace=1,
                     control_size=controls)
+
+
+def test_diagram_decompose_decrease_size():
+    text_diagram = decomposition_to_ascii(
+        gate=OffsetGate(0b110111000),
+        decomposition_rule=decompose_decrease_size,
+        register_sizes=[9])
+    print(text_diagram)
+    assert text_diagram == """
+    .-------.
+|0>-|       |-
+    |       |
+|0>-|       |-
+    |       |
+|0>-|       |-
+    |       |
+|0>-|       |-
+    |       |
[…]

### 2017-06-13 doc/paper.tex
Thinko fix
Polishing paper content - Added note about constant factors being important - Rewrote some of the conclusion - Added QFT inverting multiplication remarks to conclusion - Fixed dependency diagram saying "scale-accumulate" instead of "scale-add".

@@
 
+Our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
+Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
+Achieving small constant factors under realistic architectural constraints is an important practical problem, but not the focus of this paper.
+
 \section{Constructions} \label{sec:construct}
@@
 Because our modular circuit constructions would have undefined behavior if any register had a value equal to or larger than $R$, we need the ancilla register to contain a value less than $R$.
-We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e.
-clean) and that the registers be as small as possible (i.e.
-have size $n = \lceil \lg_2(R-1) \rceil$, because only values up to $R-1$ need to be stored).
+We ensure this by requiring that the ancilla register's MSB be $|0\rangle$ (i.e. clean) and that the registers be as small as possible (i.e. have size $n = \lceil \lg_2(R) \rceil$).
 
@@
 
-Of the $2n+1$ qubits our period-finding construction requires: $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancillae used to implement modular multiplication in terms of scaled modular addition.
+In this paper we described how to perform various arithmetic operations using only dirty ancilla.
+We also described various techniques we used to find the constructions: pivot flipping, toggle controlling, commutator controlling, and exploration by directly manipulating circuit diagrams.
+We showed how these constructions reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
+The constructions also reduce the number of {\em clean} qubits required by $n$.
+
+Of the $2n+1$ qubits our period-finding construction requires, $1$ is used for the phase estimation qubit, $n$ are used to store the work register, and $n$ are ancillae used to implement modular multiplication in terms of scaled modular addition.
 Except for multiplication, all our arithmetic constructions use two or fewer dirty ancillae.
@@
 
-In this paper we described how to perform various arithmetic operations using only dirty ancilla.
-We also described various techniques we used to find the constructions: pivot flipping, toggle controlling, commutator controlling, and exploration by directly manipulating circuit diagrams.
-We showed how these constructions reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
-The constructions also reduce the number of {\em clean} qubits required by $n$.
+A second way to reduce modular multiplication into (slightly) smaller operations could be with commutator control.
+Note that the modular Fourier transform inverts the effect of modular multiplication: $QFT_{R} \cdot (\times K) \cdot QFT_R^{-1} = (\times K^{-1}) \pmod{R}$.
+This has two useful high-level effects.
+First, by applying the bimultiplication gate used in this paper twice, but inverting the effect on the ancilla register for the second application, we can build a proper modular multiplication gate that only affects a single register.
+Second, we can use commutator control to move controls from the modular multiplication gate onto modular Fourier transform gates.
+Assuming controlled modular QFTs can be performed with fewer ancillae than controlled modular multiplication, we could save a dirty ancilla when performing the modular multiplication by borrowing the control qubit.
 
-The number of clean qubits required to perform Shor's algorithm is now $n + O(1)$.
+Possible future improvements aside, in this paper we showed that a number of clean qubits sufficient to perform Shor's algorithm is $n + O(1)$.
 We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.

### 2017-06-22 doc/paper.tex
Tweaking

@@
 A second way to reduce modular multiplication into (slightly) smaller operations could be with commutator control.
-Note that the modular Fourier transform inverts the effect of modular multiplication: $QFT_{R} \cdot (\times K) \cdot QFT_R^{-1} = (\times K^{-1}) \pmod{R}$.
+Note that the modular Fourier transform inverts the effect of modular multiplication: $QFT_{R} \cdot (\times K \,\text{mod}\, R) \cdot QFT_R^{-1} = (\times K^{-1} \,\text{mod}\, R)$.
 This has two useful high-level effects.
@@
 Second, we can use commutator control to move controls from the modular multiplication gate onto modular Fourier transform gates.
-Assuming controlled modular QFTs can be performed with fewer ancillae than controlled modular multiplication, we could save a dirty ancilla when performing the modular multiplication by borrowing the control qubit.
+Assuming controlled modular QFTs can be performed with fewer ancillae than controlled modular multiplication, we could save a qubit when performing the modular multiplication by borrowing the control qubit.
 

### 2017-06-23 doc/paper.tex
Addressing Marissa's minor comment and Austin's major comment.

@@
 \begin{abstract}
-We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae in an unknown state that must be preserved.
+We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e. qubits in an unknown state that must be preserved).
 We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
@@
 In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
-We do so without increasing the asymptotic costs, though the constant factors of our constructions are worse.
+We do so without increasing the asymptotic costs.
+However, our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
+Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
+Achieving low T gate counts under realistic architectural constraints is an important practical problem, but not the focus of this paper.
 
@@
 If the precondition $y < R$ is violated, what occurs is undefined behavior \cite{c11}.
-
-Our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
-Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
-Achieving small constant factors under realistic architectural constraints is an important practical problem, but not the focus of this paper.
 
@@
 
-Possible future improvements aside, in this paper we showed that a number of clean qubits sufficient to perform Shor's algorithm is $n + O(1)$.
-We consider this improvement to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
-Such a construction could, by allowing factoring problems to fit into quantum computers with half as much space, reduce the time until it's possible to break current cryptographic keys by months or years (depending on how quickly quantum computers gain capacity, the constant factors of the construction, and the trade-offs between space, error correction, and qubit coherence time).
+Possible future improvements aside, in this paper we showed that the number of clean qubits sufficient to perform Shor's algorithm is no more than $n + O(1)$.
+We consider this to be a step towards a construction for Shor's algorithm that uses only $n + O(1)$ qubits total.
+We do note, however, that these improvements are of more theoretical than practical interest.
+In practice, on error corrected quantum computers, the important cost to optimize is the number of T gates.
+In that sense, the main {\em practical} contribution of this paper is the improved increment circuit.
 
@@
 
-We thank Matthew Neeley and Dave Bacon for comments on earlier versions of this paper, from which it greatly benefited.
+We thank Matthew Neeley, Dave Bacon, and Austin Fowler for comments on earlier versions of this paper, from which it greatly benefited.
 

### 2017-06-23 doc/paper.tex
Being clearer about dirty-vs-clean

@@
 \begin{abstract}
-We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e. qubits in an unknown state that must be preserved).
+We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e. extra qubits in an unknown state that must be restored before the circuit ends).
 We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
@@
 An ancilla is an extra bit or qubit that is available for use by a circuit as temporary workspace.
-Ancillae may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
+Ancillae may be initialized to a known computational basis state (``clean bit"), or be given to the circuit in an unknown and potentially entangled state that must be restored before the circuit finishes (``dirty bit").
 Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, since any temporarily unused bit is a borrowable dirty bit.

### 2017-07-05 doc/paper.tex
Include customized cls
Cite barenco et al for the many-control not to dirty toffoli construction - Also add a note that it's possible to do asymptotically better on size - Also note in the abstract that dirty bits are a thing

@@
-\documentclass[twocolumn,longbibliography]{quantumarticle}
+\documentclass[twocolumn,longbibliography]{quantumarticle-customized}
 \usepackage{amsmath}
@@
 \begin{abstract}
-We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e. extra qubits in an unknown state that must be restored before the circuit ends).
+We present reversible classical circuits for performing various arithmetic operations aided by dirty ancillae (i.e.
+extra bits/qubits in an unknown state that must be restored before the circuit ends).
 We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
@@
   \caption{
-    When $c-2$ dirty ancillae are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates.
+    When $c-2$ dirty ancillae are available, a controlled-not with $c$ controls can be reduced into $4c - 8$ Toffoli gates \cite{barenco1995}.
+    It is possible to reduce the number of ancilla needed to just one, by using twice as many Toffoli gates, but that construction isn't needed in the context of period finding because there are always enough unused qubits to borrow.
   }

### 2018-01-19 doc/paper.tex
Addressing quantum reviewer comments - Zalka had previously achieved 1.5n; oh my

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{graphicx}
@@
 extra bits/qubits in an unknown state that must be restored before the circuit ends).
-We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $2n+2$ \cite{takahashi2006, haner2016} to $n+2$, and the total number of qubits needed from $2n+2$ to $2n+1$, without increasing the asymptotic size or depth of the circuit.
+We improve the number of clean qubits needed to factor an $n$-bit number with Shor's algorithm \cite{Shor1999} from $1.5n + O(1)$ \cite{zalka2006} to $n+2$, assisted by $n-1$ dirty qubits, without increasing the asymptotic size or depth of the circuit.
 \end{abstract}
@@
 
-In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
-We do so without increasing the asymptotic costs.
-However, our circuit constructions are not optimized to achieve good constant factors on circuit depth or gate count.
-Our constructions also ignore machine architecture; they assume all-to-all connectivity between qubits.
-Achieving low T gate counts under realistic architectural constraints is an important practical problem, but not the focus of this paper.
+It is important to note that, pragmatically speaking, it is far more important to, for example, achieve low T gate counts under plausible architectural constraints than to reduce the number of clean qubits required by a circuit.
+Our goal in this paper is not to come up with an implementation of Shor's algorithm optimized in the ways that matter for plausible future quantum computer architectures.
+Our goal is to explore the consequences of picking a metric, trying to optimize it, and seeing what ideas fall out.
+The circuit constructions we present will not be optimized to achieve good constant factors on circuit depth or gate count (though their asymptotics are fine).
+And they will ignore machine architecture; they assume all-to-all connectivity between qubits.
+And the {\em total} number of qubits we use ($2n+1$) is higher than previous work.
+But interesting ideas don't always come from thinking directly about a problem; there is much to be found by placing artificial goals and seeing where they lead.
+That exploration is our intent.
+In this paper, we reduce the number of clean qubits required to factor an $n$-bit number with Shor's algorithm from $1.5n + O(1)$ clean qubits \cite{zalka2006} to $n+2$ clean qubits assisted by $n-1$ dirty qubits.
+We do so without increasing the asymptotic circuit depth or size.
 
@@
 As in \cite{beauregard2003}, we perform controlled modular multiplication with a second register and modular scaled-addition operations.
-However, to allow the second register to be dirty, we use an additional scale-add as well as a negation operation.
+However, to allow the second register to be dirty, we use an additional scale-add as well as a negation operation as in \cite{zalka2006}.
 See \autoref{fig:controlled-modular-multiply} for the circuit diagram.
@@
       Takahashi et al. \cite{takahashi2006} & 2006 & $\Theta(n^3 \lg \frac{1}{\varepsilon})$ & $\Theta(n^3 \lg \frac{n}{\varepsilon} \lg \frac{1}{\varepsilon})$ & $2n+2$ & $2n+2$ \\
+      Zalka \cite{zalka2006} & 2006 & $\Theta(n^3 \lg \frac{1}{\varepsilon})$ & $\Theta(n^3 \lg \frac{n}{\varepsilon} \lg \frac{1}{\varepsilon})$ & $1.5n+O(1)$ & $1.5n+O(1)$ \\
       H\"{a}ner et al. \cite{haner2016} & 2016 & $\Theta(n^3)$ & $\Theta(n^3 \lg n)$ & $2n+2$ & $2n+2$ \\
@@
     Space-efficient constructions of Shor's algorithm over time.
-    The table only includes papers that presented explicit circuit constructions improving the minimum number of qubits required to perform the algorithm, or else improved the time complexity at the minimum known size.
+    The table only includes papers that presented explicit circuit constructions and improved the minimum number of qubits required to perform the algorithm (or else the time complexity at a slightly larger qubit count).
     $M(n)$ is the classical time-complexity of multiplication, which is known to be asymptotically at most $n \cdot (\lg n) \cdot 2^{O(\lg^* n)}$ \cite{furer2007}.
@@
 In \autoref{fig:table-over-time} we show how the number of qubits needed for Shor's algorithm has improved over time.
-The previous space improvement, from $2n+3$ to $2n+2$, occurred over a decade ago.
-
-Our main improvements over previous arithmetic constructions are 1) the use of pivot-flips for modular addition, 2) the use of bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
+Our main improvements over previous arithmetic constructions are 1) the use of pivot-flips for modular addition, 2) the use of dirty bimultiplication for modular multiplication, and 3) the $O(n)$ incrementer requiring only a single dirty ancilla.
 
@@
 Pivot flips also require ancillae, but the ancillae can be dirty and, in the context of Shor's algorithm, there are always qubits available to borrow whenever a pivot-flip is needed.
-This improvement ends up saving a qubit, reducing the total number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
+This improvement ends up saving a qubit, reducing the total number of qubits we would have needed from $2n+2$ as in \cite{haner2016} to $2n+1$.
 
-Previous modular multiplication constructions required a clean ancilla register \cite{haner2016}.
-We use an extra scale-add and negation to prevent the ancilla register from leaking into the work register, and add a clean-up multiplication informed by the work register at the end of the circuit to undo any effects on the ancilla register.
+Previous modular multiplication constructions did not work with a dirty ancilla register.
+They either required a clean ancilla register because they leaked the ancilla register into the work register \cite{haner2016}, or else a trashable ancilla register because there was no way to undo the damage being done by the multiplications \cite{zalka2006}.
+The leakage problem is fixed by adding an extra scale-add operation and a negation operation \cite{zalka2006}.
+We fixed the trashing problem by measuring the work register at the end of the circuit, and using its value to drive a clean-up multiplication that restores the ancilla register.
 This improvement allows $n-1$ of the qubits in the ancilla register to be dirty.
@@
 We also described various techniques we used to find the constructions: pivot flipping, toggle controlling, commutator controlling, and exploration by directly manipulating circuit diagrams.
-We showed how these constructions reduce the smallest known number of qubits required to perform Shor's algorithm from $2n+2$ to $2n+1$.
-The constructions also reduce the number of {\em clean} qubits required by $n$.
+We showed how these constructions reduce the number of clean qubits required to perform Shor's algorithm from $1.5n+O(1)$ to $n+2$.
 
@@
 The Chinese remainder theorem guarantees that $x \mod R$ could be uniquely represented as the pair of half-sized values $(x \mod p, x \mod q)$.
-Multiplications of the half-sized values could use and reuse the same half-sized clean ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one ancilla register) and a total qubit count of $\approx 1.5n$.
+Multiplications of the half-sized values could use and reuse the same half-sized ancilla register, for a total of three half-sized registers (one for $x \mod p$, one for $x \mod q$, and one hopefully-dirty ancilla register).
 Despite the large reduction in the number of ancillae available to the underlying arithmetic operations, each operation would still have more than enough ancillae and contribute no additional ancillae to the overall circuit.
@@
 Of course, since we use Shor's algorithm to compute the factorization of $R$, using the factorization of $R$ to optimize Shor's algorithm would be paradoxical.
+The $1.5n$ achieved in \cite{zalka2006} was done via this kind of method, except that instead of factoring $R$ Zalka factored the value to multiply by and the factoring was done modulo $R$ (by stopping a generalized gcd at a midway point) instead of in $\mathbb{Z}$.
 
@@
 In practice, on error corrected quantum computers, the important cost to optimize is the number of T gates.
-In that sense, the main {\em practical} contribution of this paper is the improved increment circuit.
+In that sense, the main {\em practical} contribution of this paper is the improved increment circuit and the demonstration of constructing circuits under tight space constraints.
 
